---
title: "R Notebook"
---


This script is a direct continuation of 1_Prepare_Datasets.Rmd.
Environment from there has to be loaded before progressing.


# 0. Prepare environment

Load necessary packages and import functions.

```{r}
# Data import
library(googledrive)
library(readxl)
#library(phyloseq)


# All data format modifications and filtering and plotting
library(tidyverse)
'%!in%' <- function(x,y)!('%in%'(x,y))
source("Code/Functions.R")
library(zoo)

# Statistical tools
library(vegan)

# Specific plotting tools
library(ggpubr)
library(paletteer)
library(viridis)
library(cowplot)
#library(ggnewscale)
source("Code/scales_and_themes.R")

# Make sure tidyverse is the latest loaded package
detach("package:tidyverse", unload=TRUE)
library(tidyverse)
```



# 1. Filter and normalize 18S data


```{r}
seq_depth_plot_18S <- function(data) {
  tmp <- data %>%
    group_by(Library_ID, Sample_type, MiSeq_library) %>% 
    summarise(Library_Size = sum(Abundance)) %>% 
    arrange(Library_Size)
  
  tmp$Index <- seq(nrow(tmp))
  tmp %>% 
    ggplot() +
    geom_point(aes(Index, Library_Size, colour = Sample_type, label = Library_ID))
}
```


```{r}
# Plot ranked sequencing depth with threshold line
All_18S_data_pr2 %>%
  filter(is.na(Fastq_ID) == FALSE) %>%
  filter(Sample_type != "Control") %>% 
  seq_depth_plot() +
  theme_cowplot(12) +
  theme(aspect.ratio = 1,
        strip.background =element_blank()) +
  geom_abline(intercept = 10000, linetype = 4)
ggsave("Figure_output/Sequencing_Depth_18SPR.pdf", width = 7, height = 4.5)



# Filter out libraries under threshold line
Good_18S_data_pr2 <- All_18S_data_pr2 %>% 
  group_by(Library_ID) %>%
  filter(Site_ID == "QU39") %>% 
  filter(sum(Abundance)>10000) %>% 
  ungroup()

print("Median 18S library size:")
Good_18S_data_pr2 %>% 
  group_by(Library_ID) %>% 
  summarise(SeqDepth = sum(Abundance)) %>%
  pull(SeqDepth) %>% median

print("Number of final 18S libraries:")
Good_18S_data_pr2 %>% 
  pull(Library_ID) %>% unique %>%  length

```


```{r}
pr2_18S_filtered_df <- Good_18S_data_pr2 %>%
  filter(Sample_type == "DNA",
         Project_name == "QU39",
         Kingdom == "Eukaryota") %>%
  mutate(ASV = paste(MiSeq_library, ASV)) %>% 
  rarefy_sequence_depth(sample = Library_ID, taxa = ASV, abundance = Abundance) %>% 
  mutate(Phylum = Division)
```



# 2. Filter and normalize COI data

In this section, we have a look at sequenced negative controls for the 18S and the COI libraries, and remove samples that have sequence depth lower than the negative controls.\
Here we also produce statistics for annotation levels and read success.\
\
Plotting sequencing depth and removing low libraries for COI:


```{r}
seq_depth_plot <- function(data) {
  tmp <- data %>%
    mutate(Sample_Type = ifelse(
      is.na(Line_out_depth),
      "Control",
      "True_Sample")) %>% 
    group_by(Library_ID, Sample_Type, MiSeq_library) %>% 
    summarise(Library_Size = sum(Abundance)) %>% 
    arrange(Library_Size)
  
  tmp$Index <- seq(nrow(tmp))
  tmp %>% 
    ggplot() +
    geom_point(aes(Index, Library_Size, colour = Sample_Type, label = Library_ID))
}
```


```{r}
# Plot ranked sequencing depth with threshold line
All_COI_data %>% 
  seq_depth_plot() +
  theme_cowplot(12) +
  theme(aspect.ratio = 1,
        strip.background =element_blank()) +
  geom_abline(intercept = 10000, linetype = 4)
ggsave("Figure_output/Sequencing_Depth_COI.pdf", width = 7, height = 4.5)




# Filter out libraries under threshold line
Good_COI_data <- All_COI_data %>% 
  group_by(Library_ID) %>%
  filter(Site_ID == "QU39") %>% 
  filter(sum(Abundance)>10000) %>% 
  ungroup()

print("Median COI library size:")
Good_COI_data %>% 
  group_by(Library_ID) %>% 
  summarise(SeqDepth = sum(Abundance)) %>%
  pull(SeqDepth) %>% median

print("Number of final COI libraries:")
Good_COI_data %>% 
  pull(Library_ID) %>% unique %>%  length


#### NOTE: Taxonomic correction of the subgenus Ditrichocorycaeus, not yet accepted.
Good_COI_data <- Good_COI_data %>%
  mutate(Genus = ifelse(Genus == "Ditrichocorycaeus", "Corycaeus", Genus),
         Species = ifelse(Species == "Ditrichocorycaeus anglicus","Corycaeus anglicus", Species))
```

```{r}
# Get list of Straig of Georgia zooplankton taxa. 
Taxa_MIC <- df_count_2017 %>%
  filter(Abundance > 0) %>%
  group_by(Genus, Tax1) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  mutate(Species = ifelse(Tax1 == "*sp.", paste(Genus, "sp.", sep = " "), paste(Genus, Tax1, sep = " "))) %>%
  mutate(Level = ifelse(Tax1 == "*sp.", "Genus", "Species")) %>% 
  pull(Genus) %>% unique()

COI_filtered_df <- Good_COI_data %>%
   filter(Project_name == "QU39",
         year(Sample_date) %in% c(2017, 2018, 2019)) %>%
    filter(Sample_type == "eDNA",
         Kingdom == "Animalia") %>%
  filter(Genus %in% Taxa_MIC) %>% 
  filter(is.na(Abundance) == FALSE) %>%
  mutate(ASV = paste(ASV, MiSeq_run)) %>% 
  group_by(Library_ID, Hakai_ID, Line_out_depth, Sample_date, Project_name,
           ASV, Species, Genus, Family, Order, Class, Phylum) %>% 
  summarise(Abundance = sum(Abundance))
```



# 3. Filter and normalize 16S data

Force 16S data to same metadata standard as 18S and COI

```{r}
colnames(All_16S_data)
```


```{r}
rarefied_16S_df <- All_16S_data |>
  transmute(
    Library_ID = paste(Sample_ID, "_16S"),
    Hakai_ID = Sample_ID,
    Line_out_depth = Depth,
    Sample_date = Date,
    Project_name = Project,
    ASV, Species, Genus, Family, Order, Class, Phylum, Abundance
  ) |> 
  rarefy_sequence_depth(sample = Library_ID, taxa = ASV, abundance = Abundance)


rarefied_16S_df
```







# 3. Stapwise aggregation of low taxa
```{r}
get_taxa <- function(data, Taxa, IQR_cutoff = 0.01) {
  
  filt_dat <- data %>%
  filter(!str_ends({{Taxa}}, "@")) %>%
  filter(!str_ends({{Taxa}}, "X")) %>% 
  filter(Line_out_depth %in% c(0, 5, 30, 100)) %>%
  filter(Abundance > 0) %>%
  index_RA(Library_ID, {{Taxa}}, Abundance)


p <- filt_dat %>%
  group_by({{Taxa}}) %>%
  mutate(IQR = IQR(RA_Index)) %>%
  filter(IQR(RA_Index) > IQR_cutoff) %>% # Calculate IQR for each species
  ungroup() %>%
  mutate(Tax_1 = fct_reorder({{Taxa}}, IQR, .desc = TRUE)) %>%  # Reorder by IQR
  ggplot(aes(Tax_1, RA_Index)) +
  geom_boxplot() +
  labs(x = as_name(ensym(Taxa)), y = "eDNA-index") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

print(p)

abundant_taxa <- filt_dat %>%
  group_by({{Taxa}}) %>%
  filter(IQR(RA_Index) > IQR_cutoff) %>% # Adjust 0.5 to your desired IQR threshold
  pull({{Taxa}}) %>% unique()

return(abundant_taxa)
  
}
# Test
# get_taxa(tmp, Species, IQR_cutoff = 0.01)
```

### 18S
```{r}
pr2_genus <- get_taxa(pr2_18S_filtered_df, Genus, IQR_cutoff = 0.01)
pr2_Genus <- pr2_18S_filtered_df %>% 
  filter(Genus %in% pr2_genus) %>% 
  group_by(Library_ID, Taxa = Genus) %>% 
  summarise(Abundance = sum(Abundance),
            Level = "Genus",
            Hakai_ID, Sample_date, Line_out_depth,
            Phylum, Class, Order, Family, Genus)


pr2_class <- pr2_18S_filtered_df %>% 
  filter(Genus %!in% pr2_genus,
         #Family %!in%  pr2_family,
         #Order %!in%  pr2_order
         ) %>% 
  get_taxa(Class, IQR_cutoff = 0.05)
pr2_Class <- pr2_18S_filtered_df %>% 
  filter(Class %in% pr2_class) %>% 
  group_by(Library_ID, Taxa = Class) %>% 
  summarise(Abundance = sum(Abundance),
            Level = "Class",
            Hakai_ID, Sample_date, Line_out_depth,
            Phylum, Class)


pr2_18S_glomm <- bind_rows(pr2_Genus,
                           #pr2_Family,
                           #pr2_Order,
                           pr2_Class)



Seq <- pr2_18S_glomm %>% 
  group_by(Library_ID, Taxa) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  index_RA(Library_ID, Taxa, Abundance)


Samp <- pr2_18S_glomm %>% 
  group_by(Library_ID, Hakai_ID, Sample_date, Line_out_depth) %>% 
  summarise()

Tax <- pr2_18S_glomm %>% 
  group_by(Taxa, Phylum, Class, Order, Family, Genus) %>% 
  summarise()

pr2_18S_glomm_indexed <- Seq %>% 
  left_join(Samp, by = "Library_ID") %>% 
  left_join(Tax, by = "Taxa")

```

### COI
```{r}
COI_species <- get_taxa(COI_filtered_df, Species, IQR_cutoff = 0.01)
COI_Species <- COI_filtered_df %>%
  filter(Species %in% COI_species) %>%
  group_by(Library_ID, Taxa = Species) %>% 
  summarise(Abundance = sum(Abundance),
            Level = "Species",
            Hakai_ID, Sample_date, Line_out_depth,
            Phylum, Class, Order, Family, Genus, Species)


COI_genus <- COI_filtered_df %>%
  filter(Species %!in% COI_species) %>% 
  get_taxa(Genus, IQR_cutoff = 0.00)
COI_Genus <- COI_filtered_df %>% 
  filter(Genus %in% COI_genus) %>% 
  group_by(Library_ID, Taxa = Genus) %>% 
  summarise(Abundance = sum(Abundance),
            Level = "Genus",
            Hakai_ID, Sample_date, Line_out_depth,
            Phylum, Class, Order, Family, Genus)


COI_glomm <- bind_rows(COI_Species,
                           COI_Genus)

Seq <- COI_glomm %>% 
  group_by(Library_ID, Taxa) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  index_RA(Library_ID, Taxa, Abundance)


Samp <- COI_glomm %>% 
  group_by(Library_ID, Hakai_ID, Sample_date, Line_out_depth) %>% 
  summarise()

Tax <- COI_glomm %>% 
  group_by(Taxa, Phylum, Class, Order, Family, Genus) %>% 
  summarise()

COI_glomm_indexed <- Seq %>% 
  left_join(Samp, by = "Library_ID") %>% 
  left_join(Tax, by = "Taxa")

```

## 12S 


```{r}
species_12S <- Good_12S_QU39 %>% 
  get_taxa(Species, IQR_cutoff = 0.001)

Species_12S <- Good_12S_QU39 %>%
  filter(Species %in% species_12S) %>%
  group_by(Library_ID, Taxa = Species) %>% 
  summarise(Abundance = sum(Abundance),
            Level = "Species",
            Hakai_ID, Sample_date, Line_out_depth,
            Class, Genus, Species)


Seq <- Species_12S %>% 
  group_by(Library_ID, Taxa) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  index_RA(Library_ID, Taxa, Abundance)


Samp <- Species_12S %>% 
  group_by(Library_ID, Hakai_ID, Sample_date, Line_out_depth) %>% 
  summarise()

Tax <- Species_12S %>% 
  group_by(Taxa, Class, Genus) %>% 
  summarise()

glomm_indexed_12S <- Seq %>% 
  left_join(Samp, by = "Library_ID") %>% 
  left_join(Tax, by = "Taxa")

glomm_indexed_12S

```



```{r}

COI_genus <- COI_filtered_df %>%
  filter(Species %!in% COI_species) %>% 
  get_taxa(Genus, IQR_cutoff = 0.00)
COI_Genus <- COI_filtered_df %>% 
  filter(Genus %in% COI_genus) %>% 
  group_by(Library_ID, Taxa = Genus) %>% 
  summarise(Abundance = sum(Abundance),
            Level = "Genus",
            Hakai_ID, Sample_date, Line_out_depth,
            Phylum, Class, Order, Family, Genus)


COI_glomm <- bind_rows(COI_Species,
                           COI_Genus)

Seq <- COI_glomm %>% 
  group_by(Library_ID, Taxa) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  index_RA(Library_ID, Taxa, Abundance)


Samp <- COI_glomm %>% 
  group_by(Library_ID, Hakai_ID, Sample_date, Line_out_depth) %>% 
  summarise()

Tax <- COI_glomm %>% 
  group_by(Taxa, Phylum, Class, Order, Family, Genus) %>% 
  summarise()

COI_glomm_indexed <- Seq %>% 
  left_join(Samp, by = "Library_ID") %>% 
  left_join(Tax, by = "Taxa")
```



## 16S

```{r}
genus_16S <- get_taxa(rarefied_16S_df, Genus, IQR_cutoff = 0.05)
Genus_16S <- rarefied_16S_df %>% 
  filter(Genus %in% genus_16S) %>% 
  group_by(Library_ID, Taxa = Genus) %>% 
  summarise(Abundance = sum(Abundance),
            Level = "Genus",
            Hakai_ID, Sample_date, Line_out_depth,
            Phylum, Class, Order, Family, Genus)


class_16S <- rarefied_16S_df %>% 
  filter(Genus %!in% genus_16S,
         #Family %!in%  pr2_family,
         #Order %!in%  pr2_order
         ) %>% 
  get_taxa(Class, IQR_cutoff = 0.05)
Class_16S <- rarefied_16S_df %>% 
  filter(Class %in% class_16S) %>% 
  group_by(Library_ID, Taxa = Class) %>% 
  summarise(Abundance = sum(Abundance),
            Level = "Class",
            Hakai_ID, Sample_date, Line_out_depth,
            Phylum, Class)


glomm_16S <- bind_rows(Genus_16S,
                           Class_16S)



Seq <- glomm_16S %>% 
  group_by(Library_ID, Taxa) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  index_RA(Library_ID, Taxa, Abundance)


Samp <- glomm_16S %>% 
  group_by(Library_ID, Hakai_ID, Sample_date, Line_out_depth) %>% 
  summarise()

Tax <- glomm_16S %>% 
  group_by(Taxa, Phylum, Class, Order, Family, Genus) %>% 
  summarise()

glomm_indexed_16S <- Seq %>% 
  left_join(Samp, by = "Library_ID") %>% 
  left_join(Tax, by = "Taxa")


  
```


```{r}
rarefied_16S_df |> 
  group_by(Genus, Order) |> 
  summarise(Abundance = sum(Abundance)) |>
  arrange(-Abundance)
```








Most abundant bacteria:

```{r}
abundant_Orders <-  |>
  #filter(Phylum == "Cyanobacteria") |> 
  
  
  #RA
  group_by(Sample_ID, Date, Depth) |> 
  mutate(RA = Abundance / sum(Abundance)) |> 
  group_by(Order, Sample_ID, Date, Depth) |>
  summarise(RA = sum(RA, na.rm = TRUE)) |> 
  group_by(Order) |> 
  summarise(RA = mean(RA)) |> 
  arrange(-RA) |> 
  filter(RA > 0.01) |> 
  pull(Order)
  
  Good_16S_QU39
  
  abundant_Orders
  
```






```{r}
COI_glomm_indexed
glomm_indexed_12S
pr2_18S_glomm_indexed
glomm_indexed_16S
```




# 4. Environmental data

Get list of samples
```{r}
samples_COI <-  COI_glomm_indexed %>% 
  pull(Hakai_ID) %>% 
  unique()

samples_18S <-  pr2_18S_glomm_indexed %>% 
  pull(Hakai_ID) %>% 
  unique()

all_samples <- c(samples_COI, samples_18S) %>% unique()
```


Extract Hakai Microbial metadata for samples
```{r}
QMIC <- read_excel("./Data_import/2024-02-09_170138_HakaiData_microbial.xlsx",
                   sheet = "Hakai Data") %>% 
  filter(hakai_id %in% all_samples) %>% 
  arrange(event_pk)

QMIC

niskin <- QMIC %>% 
  mutate(NISKIN = paste(event_pk, line_out_depth, sep = "_")) %>% 
  select(hakai_id,date, line_out_depth, NISKIN ) %>% 
  mutate(date = as_date(date))
niskin
```
Chlorophyll bottle data
```{r}
CHL <- read_excel("./Data_import/2023-12-21_112549_HakaiData_chlorophyll_15-20.xlsx", sheet = "Hakai Data") %>% 
  mutate(NISKIN = paste(event_pk, line_out_depth, sep = "_")) %>% 
  filter(NISKIN %in% niskin$NISKIN) %>% 
  #select(NISKIN, filter_type, chla, phaeo) %>%
  group_by(NISKIN, filter_type) %>% 
  summarise(chla = mean(chla),
            phaeo = mean(phaeo)) %>%
  pivot_wider(names_from = filter_type, values_from = c(chla, phaeo))
CHL
```
Nutrient bottle data
```{r}
NUT <- read_excel("./Data_import/2023-12-21_113016_HakaiData_nutrients.xlsx", sheet = "Hakai Data") %>% 
  mutate(NISKIN = paste(event_pk, line_out_depth, sep = "_")) %>% 
  filter(NISKIN %in% niskin$NISKIN) %>% 
  select(NISKIN, no2_no3_um, po4, sio2) %>% 
  group_by(NISKIN) %>% 
  summarise(no2_no3_um = mean(no2_no3_um),
            po4 = mean(po4),
            sio2 = mean(po4))
NUT
```

CTD Bottle data
```{r}
# Identify the correct CTD drops (Can be improved)
drops <- read_excel("./Data_import/CTD.xlsx", sheet = "Drops") %>% 
  mutate(date = as_date(`Start time`)) %>% 
  filter(date %in% niskin$date,
         `Target depth (m)` > 240) %>% 
  group_by(date) %>% 
  summarise(CastPK = min(`Cast PK`))

# Load and modify CTD data, filter by drop.
CTD <- read_excel("./Data_import/CTD.xlsx", sheet = "Data") %>% 
  transmute(CastPK = `Cast PK`,
            Distance_from_station = `Distance from station`,
            Station,
            Time = `Measurement time`,
            Date = as_date(Time), 
            Depth = `Depth (m)`,
            Conductivity = `Conductivity (mS/cm)`,
            Temperature = `Temperature (deg C)`,
            Pressure = `Pressure (dbar)`,
            PAR = `PAR (umol m-2 s-1)`,
            Chlorophyll = `Fluorometry Chlorophyll (ug/L)`,
            Turbidity = `Turbidity (FTU)`,
            O2 = `Dissolved O2 (mL/L)`,
            Salinity = `Salinity (PSU)`) %>% 
  filter(CastPK %in% drops$CastPK)

# Function for extracting depth data.
closest <- function(dat, m) {
  dat %>% 
  group_by(Date) %>% 
  slice(which.min(abs(Depth - m))) %>% 
  mutate(Depth = m)
}

# Extract data for 0, 5, 30, 100 and 260 m.
ctd_bottle <- bind_rows(
  closest(CTD, 0),
  closest(CTD, 5),
  closest(CTD, 30),
  closest(CTD, 100),
  closest(CTD, 260)) %>%
  select(date = Date, line_out_depth = Depth, Temperature, Salinity, PAR, Turbidity, O2)

ctd_bottle


CTD

```


Merge all datasets to bottle file:
```{r}
bottle_file <- niskin %>% 
  left_join(CHL, by = "NISKIN") %>% 
  left_join(NUT, by = "NISKIN") %>% 
  left_join(ctd_bottle, by = c("date", "line_out_depth"))


```



# 5. Save COI and 18S data

```{r}
saveRDS(pr2_18S_glomm_indexed, "Data_output/Data_merge/pr2_18S_glomm_indexed.rds")
saveRDS(COI_glomm_indexed, "Data_output/Data_merge/COI_glomm_indexed.rds")
saveRDS(glomm_indexed_12S, "Data_output/Data_merge/12S_glomm_indexed.rds")
saveRDS(glomm_indexed_16S, "Data_output/Data_merge/16S_glomm_indexed.rds")
saveRDS(bottle_file, "Data_output/Data_merge/bottle_file.rds")
```




