---
title: "R Notebook"
---

# 0. Prepare environment

Load necessary packages and import functions.

```{r}
# Data import
library(googledrive)
library(readxl)
#library(phyloseq)


# All data format modifications and filtering and plotting
library(tidyverse)
'%!in%' <- function(x,y)!('%in%'(x,y))
source("Code/Functions.R")
library(zoo)

# Statistical tools
library(vegan)

# Specific plotting tools
library(ggpubr)
library(paletteer)
library(viridis)
library(cowplot)
#library(ggnewscale)
source("Code/scales_and_themes.R")

# Make sure tidyverse is the latest loaded package
detach("package:tidyverse", unload=TRUE)
library(tidyverse)
```



# 1. Filter and normalize 18S data


```{r}
seq_depth_plot_18S <- function(data) {
  tmp <- data %>%
    group_by(Library_ID, Sample_type, MiSeq_library) %>% 
    summarise(Library_Size = sum(Abundance)) %>% 
    arrange(Library_Size)
  
  tmp$Index <- seq(nrow(tmp))
  tmp %>% 
    ggplot() +
    geom_point(aes(Index, Library_Size, colour = Sample_type, label = Library_ID))
}
```


```{r}
# Plot ranked sequencing depth with threshold line
All_18S_data_pr2 %>%
  filter(is.na(Fastq_ID) == FALSE) %>%
  filter(Sample_type != "Control") %>% 
  seq_depth_plot() +
  theme_cowplot(12) +
  theme(aspect.ratio = 1,
        strip.background =element_blank()) +
  geom_abline(intercept = 10000, linetype = 4)
ggsave("Figure_output/Sequencing_Depth_18SPR.pdf", width = 7, height = 4.5)

All_18S_data_pr2 %>%
  group_by(Library_ID, Sample_type) %>% 
  summarise()

All_18S_data_pr2
All_18S_data_pr2 %>% 
  pull(Sample_type) %>% unique


# Filter out libraries under threshold line
Good_18S_data_pr2 <- All_18S_data_pr2 %>% 
  group_by(Library_ID) %>%
  filter(Site_ID == "QU39") %>% 
  filter(sum(Abundance)>10000) %>% 
  ungroup()

print("Median 18S library size:")
Good_18S_data_pr2 %>% 
  group_by(Library_ID) %>% 
  summarise(SeqDepth = sum(Abundance)) %>%
  pull(SeqDepth) %>% median

print("Number of final 18S libraries:")
Good_18S_data_pr2 %>% 
  pull(Library_ID) %>% unique %>%  length

```


```{r}
pr2_18S_filtered_df <- Good_18S_data_pr2 %>%
  filter(Sample_type == "DNA",
         Project_name == "QU39",
         Kingdom == "Eukaryota") %>%
  mutate(ASV = paste(MiSeq_library, ASV)) %>% 
  rarefy_sequence_depth(sample = Library_ID, taxa = ASV, abundance = Abundance) %>% 
  mutate(Phylum = Division)
```



# 2. Filter and normalize COI data

In this section, we have a look at sequenced negative controls for the 18S and the COI libraries, and remove samples that have sequence depth lower than the negative controls.\
Here we also produce statistics for annotation levels and read success.\
\
Plotting sequencing depth and removing low libraries for COI:


```{r}
seq_depth_plot <- function(data) {
  tmp <- data %>%
    mutate(Sample_Type = ifelse(
      is.na(Line_out_depth),
      "Control",
      "True_Sample")) %>% 
    group_by(Library_ID, Sample_Type, MiSeq_library) %>% 
    summarise(Library_Size = sum(Abundance)) %>% 
    arrange(Library_Size)
  
  tmp$Index <- seq(nrow(tmp))
  tmp %>% 
    ggplot() +
    geom_point(aes(Index, Library_Size, colour = Sample_Type, label = Library_ID))
}
```


```{r}
# Plot ranked sequencing depth with threshold line
All_COI_data %>% 
  seq_depth_plot() +
  theme_cowplot(12) +
  theme(aspect.ratio = 1,
        strip.background =element_blank()) +
  geom_abline(intercept = 10000, linetype = 4)
ggsave("Figure_output/Sequencing_Depth_COI.pdf", width = 7, height = 4.5)


# Filter out libraries under threshold line
Good_COI_data <- All_COI_data %>% 
  group_by(Library_ID) %>%
  filter(Site_ID == "QU39") %>% 
  filter(sum(Abundance)>10000) %>% 
  ungroup()

print("Median COI library size:")
Good_COI_data %>% 
  group_by(Library_ID) %>% 
  summarise(SeqDepth = sum(Abundance)) %>%
  pull(SeqDepth) %>% median

print("Number of final COI libraries:")
Good_COI_data %>% 
  pull(Library_ID) %>% unique %>%  length


#### NOTE: Taxonomic correction of the subgenus Ditrichocorycaeus, not yet accepted.
Good_COI_data <- Good_COI_data %>%
  mutate(Genus = ifelse(Genus == "Ditrichocorycaeus", "Corycaeus", Genus),
         Species = ifelse(Species == "Ditrichocorycaeus anglicus","Corycaeus anglicus", Species))
```

```{r}
# Get list of Straig of Georgia zooplankton taxa. 
Taxa_MIC <- df_count_2017 %>%
  filter(Abundance > 0) %>%
  group_by(Genus, Tax1) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  mutate(Species = ifelse(Tax1 == "*sp.", paste(Genus, "sp.", sep = " "), paste(Genus, Tax1, sep = " "))) %>%
  mutate(Level = ifelse(Tax1 == "*sp.", "Genus", "Species")) %>% 
  pull(Genus) %>% unique()

COI_filtered_df <- Good_COI_data %>%
   filter(Project_name == "QU39",
         year(Sample_date) %in% c(2017, 2018, 2019)) %>%
    filter(Sample_type == "eDNA",
         Kingdom == "Animalia") %>%
  filter(Genus %in% Taxa_MIC) %>% 
  filter(is.na(Abundance) == FALSE) %>%
  mutate(ASV = paste(ASV, MiSeq_run)) %>% 
  group_by(Library_ID, Hakai_ID, Line_out_depth, Sample_date, Project_name,
           ASV, Species, Genus, Family, Order, Class, Phylum) %>% 
  summarise(Abundance = sum(Abundance))
```


# 3. Stapwise aggregation of low taxa
```{r}
get_taxa <- function(data, Taxa, IQR_cutoff = 0.01) {
  
  filt_dat <- data %>%
  filter(!str_ends({{Taxa}}, "@")) %>%
  filter(!str_ends({{Taxa}}, "X")) %>% 
  filter(Line_out_depth %in% c(0, 5, 30, 100)) %>%
  filter(Abundance > 0) %>%
  index_RA(Library_ID, {{Taxa}}, Abundance)


p <- filt_dat %>%
  group_by({{Taxa}}) %>%
  mutate(IQR = IQR(RA_Index)) %>%
  filter(IQR(RA_Index) > IQR_cutoff) %>% # Calculate IQR for each species
  ungroup() %>%
  mutate(Tax_1 = fct_reorder({{Taxa}}, IQR, .desc = TRUE)) %>%  # Reorder by IQR
  ggplot(aes(Tax_1, RA_Index)) +
  geom_boxplot() +
  labs(x = as_name(ensym(Taxa)), y = "eDNA-index") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

print(p)

abundant_taxa <- filt_dat %>%
  group_by({{Taxa}}) %>%
  filter(IQR(RA_Index) > IQR_cutoff) %>% # Adjust 0.5 to your desired IQR threshold
  pull({{Taxa}}) %>% unique()

return(abundant_taxa)
  
}
# Test
# get_taxa(tmp, Species, IQR_cutoff = 0.01)
```

### 18S
```{r}
pr2_genus <- get_taxa(pr2_18S_filtered_df, Genus, IQR_cutoff = 0.05)
pr2_Genus <- pr2_18S_filtered_df %>% 
  filter(Genus %in% pr2_genus) %>% 
  group_by(Library_ID, Taxa = Genus) %>% 
  summarise(Abundance = sum(Abundance),
            Level = "Genus",
            Hakai_ID, Sample_date, Line_out_depth,
            Phylum, Class, Order, Family, Genus)


pr2_family <- pr2_18S_filtered_df %>% 
  filter(Genus %!in% pr2_genus) %>% 
  get_taxa(Family, IQR_cutoff = 0.05)
pr2_Family <- pr2_18S_filtered_df %>% 
  filter(Family %in% pr2_family) %>% 
  group_by(Library_ID, Taxa = Family) %>% 
  summarise(Abundance = sum(Abundance),
            Level = "Family",
            Hakai_ID, Sample_date, Line_out_depth,
            Phylum, Class, Order, Family)


pr2_order <- pr2_18S_filtered_df %>% 
  filter(Genus %!in% pr2_genus,
         Family %!in%  pr2_family) %>% 
  get_taxa(Order, IQR_cutoff = 0.05)
pr2_Order <- pr2_18S_filtered_df %>% 
  filter(Order %in% pr2_family) %>% 
  group_by(Library_ID, Taxa = Order) %>% 
  summarise(Abundance = sum(Abundance),
            Level = "Order",
            Hakai_ID, Sample_date, Line_out_depth,
            Phylum, Class, Order)

pr2_class <- pr2_18S_filtered_df %>% 
  filter(Genus %!in% pr2_genus,
         Family %!in%  pr2_family,
         Order %!in%  pr2_order) %>% 
  get_taxa(Class, IQR_cutoff = 0.05)
pr2_Class <- pr2_18S_filtered_df %>% 
  filter(Class %in% pr2_family) %>% 
  group_by(Library_ID, Taxa = Class) %>% 
  summarise(Abundance = sum(Abundance),
            Level = "Class",
            Hakai_ID, Sample_date, Line_out_depth,
            Phylum, Class)


pr2_18S_glomm <- bind_rows(pr2_Genus,
                           pr2_Family,
                           pr2_Order,
                           pr2_Class)



Seq <- pr2_18S_glomm %>% 
  group_by(Library_ID, Taxa) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  index_RA(Library_ID, Taxa, Abundance)


Samp <- pr2_18S_glomm %>% 
  group_by(Library_ID, Hakai_ID, Sample_date, Line_out_depth) %>% 
  summarise()

Tax <- pr2_18S_glomm %>% 
  group_by(Taxa, Phylum, Class, Order, Family, Genus) %>% 
  summarise()

pr2_18S_glomm_indexed <- Seq %>% 
  left_join(Samp, by = "Library_ID") %>% 
  left_join(Tax, by = "Taxa")

pr2_18S_glomm_indexed
```
### COI
```{r}
COI_species <- get_taxa(COI_filtered_df, Species, IQR_cutoff = 0.01)
COI_Species <- COI_filtered_df %>%
  filter(Species %in% COI_species) %>%
  group_by(Library_ID, Taxa = Species) %>% 
  summarise(Abundance = sum(Abundance),
            Level = "Species",
            Hakai_ID, Sample_date, Line_out_depth,
            Phylum, Class, Order, Family, Genus, Species)


COI_genus <- COI_filtered_df %>%
  filter(Species %!in% COI_species) %>% 
  get_taxa(Genus, IQR_cutoff = 0.00)
COI_Genus <- COI_filtered_df %>% 
  filter(Genus %in% COI_genus) %>% 
  group_by(Library_ID, Taxa = Genus) %>% 
  summarise(Abundance = sum(Abundance),
            Level = "Genus",
            Hakai_ID, Sample_date, Line_out_depth,
            Phylum, Class, Order, Family, Genus)


COI_glomm <- bind_rows(COI_Species,
                           COI_Genus)

Seq <- COI_glomm %>% 
  group_by(Library_ID, Taxa) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  index_RA(Library_ID, Taxa, Abundance)


Samp <- COI_glomm %>% 
  group_by(Library_ID, Hakai_ID, Sample_date, Line_out_depth) %>% 
  summarise()

Tax <- COI_glomm %>% 
  group_by(Taxa, Phylum, Class, Order, Family, Genus) %>% 
  summarise()

COI_glomm_indexed <- Seq %>% 
  left_join(Samp, by = "Library_ID") %>% 
  left_join(Tax, by = "Taxa")

```

# 4. Environmental data

Get list of samples
```{r}
samples_COI <-  COI_glomm_indexed %>% 
  pull(Hakai_ID) %>% 
  unique()

samples_18S <-  pr2_18S_glomm_indexed %>% 
  pull(Hakai_ID) %>% 
  unique()

all_samples <- c(samples_COI, samples_18S) %>% unique()
```


Extract Hakai Microbial metadata for samples
```{r}
QMIC <- read_excel("./Data_import/2024-02-09_170138_HakaiData_microbial.xlsx",
                   sheet = "Hakai Data") %>% 
  filter(hakai_id %in% all_samples) %>% 
  arrange(event_pk)

QMIC

niskin <- QMIC %>% 
  mutate(NISKIN = paste(event_pk, line_out_depth, sep = "_")) %>% 
  select(hakai_id,date, line_out_depth, NISKIN ) %>% 
  mutate(date = as_date(date))
niskin
```
Chlorophyll bottle data
```{r}
CHL <- read_excel("./Data_import/2023-12-21_112549_HakaiData_chlorophyll_15-20.xlsx", sheet = "Hakai Data") %>% 
  mutate(NISKIN = paste(event_pk, line_out_depth, sep = "_")) %>% 
  filter(NISKIN %in% niskin$NISKIN) %>% 
  #select(NISKIN, filter_type, chla, phaeo) %>%
  group_by(NISKIN, filter_type) %>% 
  summarise(chla = mean(chla),
            phaeo = mean(phaeo)) %>%
  pivot_wider(names_from = filter_type, values_from = c(chla, phaeo))
CHL
```
Nutrient bottle data
```{r}
NUT <- read_excel("./Data_import/2023-12-21_113016_HakaiData_nutrients.xlsx", sheet = "Hakai Data") %>% 
  mutate(NISKIN = paste(event_pk, line_out_depth, sep = "_")) %>% 
  filter(NISKIN %in% niskin$NISKIN) %>% 
  select(NISKIN, no2_no3_um, po4, sio2) %>% 
  group_by(NISKIN) %>% 
  summarise(no2_no3_um = mean(no2_no3_um),
            po4 = mean(po4),
            sio2 = mean(po4))
NUT
```

CTD Bottle data
```{r}
# Identify the correct CTD drops (Can be improved)
drops <- read_excel("./Data_import/CTD.xlsx", sheet = "Drops") %>% 
  mutate(date = as_date(`Start time`)) %>% 
  filter(date %in% niskin$date,
         `Target depth (m)` > 240) %>% 
  group_by(date) %>% 
  summarise(CastPK = min(`Cast PK`))

# Load and modify CTD data, filter by drop.
CTD <- read_excel("./Data_import/CTD.xlsx", sheet = "Data") %>% 
  transmute(CastPK = `Cast PK`,
            Distance_from_station = `Distance from station`,
            Station,
            Time = `Measurement time`,
            Date = as_date(Time), 
            Depth = `Depth (m)`,
            Conductivity = `Conductivity (mS/cm)`,
            Temperature = `Temperature (deg C)`,
            Pressure = `Pressure (dbar)`,
            PAR = `PAR (umol m-2 s-1)`,
            Chlorophyll = `Fluorometry Chlorophyll (ug/L)`,
            Turbidity = `Turbidity (FTU)`,
            O2 = `Dissolved O2 (mL/L)`,
            Salinity = `Salinity (PSU)`) %>% 
  filter(CastPK %in% drops$CastPK)

# Function for extracting depth data.
closest <- function(dat, m) {
  dat %>% 
  group_by(Date) %>% 
  slice(which.min(abs(Depth - m))) %>% 
  mutate(Depth = m)
}

# Extract data for 0, 5, 30, 100 and 260 m.
ctd_bottle <- bind_rows(
  closest(CTD, 0),
  closest(CTD, 5),
  closest(CTD, 30),
  closest(CTD, 100),
  closest(CTD, 260)) %>%
  select(date = Date, line_out_depth = Depth, Temperature, Salinity, PAR, Turbidity, O2)

ctd_bottle


CTD

```


Merge all datasets to bottle file:
```{r}
bottle_file <- niskin %>% 
  left_join(CHL, by = "NISKIN") %>% 
  left_join(NUT, by = "NISKIN") %>% 
  left_join(ctd_bottle, by = c("date", "line_out_depth"))


```



# 5. Save COI and 18S data

```{r}
saveRDS(pr2_18S_glomm_indexed, "Data_output/Data_merge/pr2_18S_glomm_indexed.rds")
saveRDS(COI_glomm_indexed, "Data_output/Data_merge/COI_glomm_indexed.rds")
saveRDS(bottle_file, "Data_output/Data_merge/bottle_file.rds")
```





# Plotting functional groups


```{r}
COI_glomm_indexed %>% 
  pull(Taxa) %>% unique()
```



```{r}
# Phytoplankton - primary producers, including various autotrophic groups
phytoplankton <- c("Aureococcus", "Bathycoccus", "Biecheleria", "Cerataulina", 
                   "Chaetoceros", "Chrysophyceae", "Coscinodiscus", "Dicrateria",
                   "Dictyocha", "Ditylum", "Gymnodinium", "Guinardia", 
                   "Hemiselmis", "Heterocapsa", "Karlodinium", "Micromonas", 
                   "Minidiscus", "Navicula", "Noctilucophyceae", "Ostreococcus", 
                   "Phaeocystis", "Picochlorum", "Prorocentrum", "Pseudo-nitzschia", 
                   "Rhizosolenia", "Skeletonema", "Teleaulax", "Thalassiosira", 
                   "Triparma")

# Zooplankton - primarily heterotrophic protists and small animals
zooplankton <- c("Acantharea_3b", "Askenasia", "Choreotrichida", "Didiniidae",
                 "Laboea", "Leegaardiella", "Litostomatea", "Pelagostrobilidium", 
                 "Protoperidinium", "Rimostrombidium_A", "Strobilidiidae_A", 
                 "Strobilidiidae_I", "Strombidiida", "Strombidiidae", "Tintinnida", 
                 "Tintinnidae", "Tontoniidae_B")

# Heterotrophic Flagellates - often bacterivores or part of microbial loops
heterotrophic_flagellates <- c("Cryothecomonas", "Paraphysomonas", "Picomonas", 
                               "Telonemia-Group-1", "Telonemia-Group-2")

# Mixotrophs - capable of both autotrophic and heterotrophic feeding modes
mixotrophs <- c("Chrysochromulina", "Haptolina", "Haptophyta", "Haptophyta_Clade_HAP3", 
                "Prymnesiales", "Prymnesiophyceae", "Prymnesium")

# Parasitic Protists - species known primarily as parasites or pathogens
parasitic_protists <- c("Hematodinium", "Syndinium", "Thalassomyces")

# Microbial or fungal-like protists - not traditional phytoplankton or zooplankton, often decomposers
microbial_fungi <- c("Aplanochytrium", "Cryptomycota", "Fungi", "Labyrinthulomycetes", 
                     "Oomycota", "Pucciniomycotina", "Thraustochytriaceae")
```

```{r}
pr2_18S_glomm_indexed  %>% 
  filter(Taxa %in% phytoplankton,
         Line_out_depth == 0,
         year(Sample_date) != 2019,
         Level != "Division") %>% 
  ggplot() +
  geom_line(aes(Sample_date, RA_Index, colour = Division)) +
  facet_wrap(~Taxa)


pr2_18S_glomm_indexed  %>% 
  filter(Taxa %in% mixotrophs,
         Line_out_depth == 0,
         year(Sample_date) != 2019,
         Level != "Division") %>% 
  ggplot() +
  geom_line(aes(Sample_date, RA_Index, colour = Division)) +
  facet_wrap(~Taxa)



pr2_18S_glomm_indexed  %>% 
  filter(Taxa %in% c(zooplankton, heterotrophic_flagellates),
         Line_out_depth == 0,
         year(Sample_date) != 2019,
         Level != "Division") %>% 
  ggplot() +
  geom_line(aes(Sample_date, RA_Index, colour = Division)) +
  facet_wrap(~Taxa)

pr2_18S_glomm_indexed  %>% 
  filter(Taxa %in% parasitic_protists,
         Line_out_depth == 0,
         year(Sample_date) != 2019,
         Level != "Division") %>% 
  ggplot() +
  geom_line(aes(Sample_date, RA_Index, colour = Division)) +
  facet_wrap(~Taxa)

pr2_18S_glomm_indexed  %>% 
  filter(Taxa %in% microbial_fungi,
         Line_out_depth == 0,
         year(Sample_date) != 2019,
         Level != "Division") %>% 
  ggplot() +
  geom_line(aes(Sample_date, RA_Index, colour = Division)) +
  facet_wrap(~Taxa)
```





## Plot some importan PP genera
```{r}
tmp1 <- tmp %>% 
  filter(Genus %in% c(
    "Thalassiosira",
    "Skeletonema",
    "Chaetoceros",
    "Ditylum",
    "Ceratium",
    "Protoperidinium",
    "Alexandrium",
    "Micromonas",
    "Bathycoccus",
    "Rhodomonas",
    "Phaeocystis",
    "Pseudo-nitzschia"
  )) %>%
  filter(Line_out_depth == 0,
         year(Sample_date) %in% c(2016, 2017, 2018)) %>%
  index_RA(Library_ID, Genus, abundance = Abundance,
           Sample_date)

tmp1  %>% 
  ggplot() +
  geom_line(aes(Sample_date, RA_Index)) +
  facet_wrap(~Genus)

  mutate(Date = factor(Sample_date)) %>% 
  ggplot() +
  #geom_bar(aes(Date, RA_Index, fill = Genus), stat = "identity", position = "fill") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

plotly::ggplotly(p)

tmp
  
```





```{r}
tmp <- Indexed_18S_pr2 <- Good_18S_data_pr2 %>%
    filter(Sample_type == "DNA",
         Project_name == "QU39") %>% 
  index_RA(sample = Library_ID, taxa = Class, abundance = Abundance,
           Sample_date, Site_ID, Project_name, Line_out_depth, Sample_type)

tmp




Indexed_18S_pr2 %>% 
  pull(Class) %>%  unique()
```


```{r}
Indexed_18S_pr2 %>% 
  filter(Class == "Bacillariophyta") %>% 
      mutate(Depth = factor(Line_out_depth, levels = c(
        250, 100, 30, 5, 0 ))) %>% 
      ggplot() +
      geom_point(aes(Sample_date, Depth, fill = RA_Index))
      
```



```{r}
pr2_18S_glomm_indexed
saveRDS(pr2_18S_glomm_indexed, file = "Data_output/pr2_18S_glomm_indexed.rds")
```


# Same but for COI

```{r}
# Get list of Straig of Georgia zooplankton taxa. 
Taxa_MIC <- df_count_2017 %>%
  filter(Abundance > 0) %>%
  group_by(Genus, Tax1) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  mutate(Species = ifelse(Tax1 == "*sp.", paste(Genus, "sp.", sep = " "), paste(Genus, Tax1, sep = " "))) %>%
  mutate(Level = ifelse(Tax1 == "*sp.", "Genus", "Species")) %>% 
  pull(Genus) %>% unique()

COI_filtered_df <- Good_COI_data %>%
   filter(Project_name == "QU39",
         year(Sample_date) %in% c(2017)) %>%
    filter(Sample_type == "eDNA",
         Kingdom == "Animalia") %>%
  filter(Genus %in% Taxa_MIC) %>% 
  filter(is.na(Abundance) == FALSE) %>%
  mutate(ASV = paste(ASV, MiSeq_run)) %>% 
  group_by(Library_ID, Hakai_ID, Line_out_depth, Sample_date, Project_name,
           ASV, Species, Genus, Family, Order, Class, Phylum) %>% 
  summarise(Abundance = sum(Abundance))
```

```{r}
get_taxa <- function(data, Taxa, IQR_cutoff = 0.01) {
  
  filt_dat <- data %>%
  filter(!str_ends({{Taxa}}, "@")) %>%
  filter(!str_ends({{Taxa}}, "X")) %>% 
  filter(Line_out_depth %in% c(0, 5, 30, 100)) %>%
  filter(Abundance > 0) %>%
  index_RA(Library_ID, {{Taxa}}, Abundance)


p <- filt_dat %>%
  group_by(Species) %>%
  mutate(IQR = IQR(RA_Index)) %>%
  filter(IQR(RA_Index) > IQR_cutoff) %>% # Calculate IQR for each species
  ungroup() %>%
  mutate(Tax_1 = fct_reorder({{Taxa}}, IQR, .desc = TRUE)) %>%  # Reorder by IQR
  ggplot(aes(Tax_1, RA_Index)) +
  geom_boxplot() +
  labs(x = as_name(ensym(Taxa)), y = "eDNA-index") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


abundant_taxa <- filt_dat %>%
  group_by({{Taxa}}) %>%
  filter(IQR(RA_Index) > IQR_cutoff) %>% # Adjust 0.5 to your desired IQR threshold
  pull({{Taxa}}) %>% unique()

print(p)
return(abundant_taxa)
  
}

get_taxa(tmp, Species, IQR_cutoff = 0.01)
```


```{r}
Species_level <- tmp %>%
  filter(!str_ends(Species, "@")) %>%
  filter(!str_ends(Species, "X")) %>% 
  filter(Line_out_depth %in% c(0, 5, 30, 100)) %>%
  filter(Abundance > 0) %>%
  index_RA(Library_ID, Species, Abundance)


Species_level %>%
  group_by(Species) %>%
  mutate(IQR = IQR(RA_Index)) %>%
  filter(IQR(RA_Index) > 0.01) %>% # Calculate IQR for each species
  ungroup() %>%
  mutate(Species = fct_reorder(Species, IQR, .desc = TRUE)) %>%  # Reorder by IQR
  ggplot(aes(Species, RA_Index)) +
  geom_boxplot() +
  labs(x = "Species", y = "RA Index") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


abundant_species <- Species_level %>%
  group_by(Species) %>%
  filter(IQR(RA_Index) > 0.001) %>% # Adjust 0.5 to your desired IQR threshold
  pull(Species) %>% unique()

abundant_species
```


```{r}
Genus_level <- tmp %>%
  filter(!str_ends(Genus, "@")) %>%
  filter(!str_ends(Genus, "X")) %>% 
  filter(Line_out_depth %in% c(0, 5, 30, 100)) %>%
  filter(Species %!in% abundant_species) %>% 
  index_RA(Library_ID, Genus, Abundance)

Genus_level %>%
  group_by(Genus) %>%
  #filter(IQR(RA_Index) > 0.001) %>% # Adjust 0.5 to your desired IQR threshold
  ungroup() %>% 
  mutate(Genus = fct_reorder(Genus, RA_Index, .fun = ~ quantile(.x, 0.75), .desc = TRUE)) %>%
  ggplot() +
  geom_boxplot(aes(Genus, RA_Index)) +
  labs(x = "Genus", y = "RA Index") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


abundant_genera <- Genus_level %>%
  group_by(Genus) %>%
  #filter(IQR(RA_Index) > 0.01) %>% # Adjust 0.5 to your desired IQR threshold
  pull(Genus) %>% unique()

abundant_genera
```



```{r}
Species <- tmp %>% 
  filter(Species %in% abundant_species) %>% 
  group_by(Library_ID, Taxa = Species) %>% 
  summarise(Abundance = sum(Abundance),
            Level = "Species",
            Hakai_ID, Sample_date, Line_out_depth,
            Phylum, Class, Order, Family, Genus, Species)



Genus <- tmp %>% 
  filter(Genus %in% abundant_genera) %>% 
  group_by(Library_ID, Taxa = Genus) %>% 
  summarise(Abundance = sum(Abundance),
            Level = "Genus",
            Hakai_ID, Sample_date, Line_out_depth,
            Phylum, Class, Order, Family, Genus)



COI_glom <- bind_rows(Species, Genus)

COI_glom_indexed <- COI_glom %>%
  index_RA(Library_ID, Taxa, Abundance) %>% 
  left_join(select(COI_glom, !Abundance), by = c("Library_ID", "Taxa"))

COI_glom_indexed
```



