---
title: "R Notebook"
---


```{r}
# Data import

library(readxl)
#library(phyloseq)


# All data format modifications and filtering and plotting
library(tidyverse)
'%!in%' <- function(x,y)!('%in%'(x,y))
source("Code/Functions.R")
library(zoo)

# Statistical tools
library(vegan)

# Specific plotting tools
library(ggpubr)
library(paletteer)
library(viridis)
library(cowplot)
#library(ggnewscale)
source("Code/scales_and_themes.R")

# Make sure tidyverse is the latest loaded package
detach("package:tidyverse", unload=TRUE)
library(tidyverse)
```


```{r}
Dataset_COI <- readRDS("Data_output/Data_merge/COI_glomm_indexed.rds")
Dataset_18S <- readRDS("Data_output/Data_merge/pr2_18S_glomm_indexed.rds")
Dataset_Bottle <- readRDS("Data_output/Data_merge/bottle_file.rds")


```


```{r}
# Custom color scale for depths
depth_colors <- c("0" = "#add8e6",    # Light blue for 0 m
                  "5" = "#87ceeb",    # Sky blue for 5 m
                  "30" = "#4682b4",   # Steel blue for 30 m
                  "100" = "#00008b")  # Dark blue for 100 m
```



# Interpolation

	1.	Fill Missing Combinations with Zeros: The complete(Sample_ID, Sample_date, Taxa, Sampling_Depth, fill = list(Relative_Abundance = 0)) line ensures that any combination without observed data is filled with 0 for Relative_Abundance.
	2.	Generate Weekly Dates: complete(Sample_date = seq.Date(min(Sample_date), max(Sample_date), by = "week")) generates weekly dates for each Taxa and Sampling_Depth combination.
	3.	Interpolate Missing Values: na.approx() fills in any NAs generated by the new weekly dates in Sample_date.


```{r}
library(tidyverse)
library(zoo)

# Complete the data frame with all combinations of Sample_ID, Sample_date, Taxa, and Sampling_Depth
complete_data <- Dataset_COI %>%
  filter(Line_out_depth %in% c(0,5,30,100)) %>%
  mutate(Line_out_depth = factor(Line_out_depth, levels = c("0", "5", "30", "100"))) %>% 
  mutate(Sample_date = as.Date(Sample_date)) %>% 
  complete(Sample_date, Taxa, Line_out_depth, fill = list(RA_Index = 0)) %>% 
  # Ensure dates are properly ordered after completion
  arrange(Sample_date, Taxa, Line_out_depth)

strict_weekly_dates <- seq.Date(from = min(complete_data$Sample_date),
                                to = max(complete_data$Sample_date),
                                by = "week")

smoothed_data <- complete_data %>%
  # Expand to include all combinations of Taxa, Depth, and weekly dates
  group_by(Taxa, Line_out_depth) %>%
  complete(Sample_date = strict_weekly_dates) %>%
  arrange(Sample_date) %>%
  # Interpolate missing values in RA_Index within each group
  mutate(RA_Index = ifelse(is.na(RA_Index),
                           na.approx(RA_Index, rule = 2, na.rm = FALSE),
                           RA_Index)) %>%
  # Apply 2-week rolling mean for smoothing (7-day intervals)
  mutate(Smoothed_Relative_Abundance = rollmean(RA_Index, k = 2, fill = NA, align = "center")) %>%
  ungroup() %>%
  # Filter to include only rows with Smoothed_Relative_Abundance calculated
  filter(!is.na(Smoothed_Relative_Abundance)) %>% 
  filter(Sample_date %in% strict_weekly_dates)

```



Plot Acartia example
```{r}
# View the final interpolated dataset
smoothed_data %>% 
  filter(Taxa == "Acartia longiremis") %>% 
  ggplot() +
  geom_line(aes(Sample_date, Smoothed_Relative_Abundance, colour = Line_out_depth)) +
  scale_color_manual(values = depth_colors) +
  labs(color = "Depth (m)") +
  theme_minimal() +
  theme(legend.position = "right")


```
# Average over sample depth?

```{r}
smoothed_data %>% 
  group_by(Sample_date, Taxa) %>% 
  summarise(Smoothed_Relative_Abundance = mean(Smoothed_Relative_Abundance)) %>% 
  filter(Taxa == "Acartia longiremis") %>% 
  ggplot() +
  geom_line(aes(Sample_date, Smoothed_Relative_Abundance)) +
  labs(color = "Depth (m)") +
  theme_minimal() +
  theme(legend.position = "right")
```






# TSL

```{r}
library(forecast)

# Example for one Taxa and one Sampling_Depth
# Subset the data for a specific Taxa and Depth
taxa_depth_data <- smoothed_data %>%
  filter(Taxa == "Acartia longiremis", Line_out_depth == "0") %>%
  arrange(Sample_date) %>% 
  mutate(Smoothed_Relative_Abundance = ifelse(is.na(Smoothed_Relative_Abundance) == TRUE, 0,
                                              Smoothed_Relative_Abundance))

# Convert to time series format for STL (assuming daily data)
taxa_depth_ts <- ts(taxa_depth_data$Smoothed_Relative_Abundance, 
                    start = c(2017, 8),  # Set start year and day based on data
                    frequency = 52)  # Daily frequency for a 3-year period
taxa_depth_ts
# Perform STL decomposition
stl_decomp <- stl(taxa_depth_ts, s.window = "periodic")

# Plot the STL decomposition
autoplot(stl_decomp)
```

```{r}

taxa_depth_data <- smoothed_data %>%
  filter(Taxa == "Acartia longiremis") %>%
  arrange(Sample_date) %>% 
  mutate(Smoothed_Relative_Abundance = ifelse(is.na(Smoothed_Relative_Abundance) == TRUE, 0,
                                              Smoothed_Relative_Abundance))

taxa_depth_data

depth = 30
  depth_data <- taxa_depth_data %>%
    filter(Line_out_depth == depth) %>%
    pull(Smoothed_Relative_Abundance) %>% 
    ts(start = c(2017, 8), frequency = 52)  # Assuming weekly data
  
  # Decompose the time series
  decomposed <- stl(depth_data, s.window = "periodic")
  
  # Plot the decomposition
  autoplot(decomposed) + 
    ggtitle(paste("STL Decomposition for Depth:", depth)) +
    labs(x = "Date", y = "Relative Abundance") +
    theme_minimal()


decomposed <- stl(depth_data, s.window = "periodic")
```

# Correlation matrix

```{r}

library(igraph)
library(ggraph)

# Assume data frame has Smoothed_Relative_Abundance, Sample_date, Taxa, Line_out_depth
seasonal_data <- smoothed_data %>%
  mutate(Smoothed_Relative_Abundance = ifelse(is.na(Smoothed_Relative_Abundance) == TRUE, 0,
                                              Smoothed_Relative_Abundance)) %>%
  filter(Line_out_depth == 0) %>% 
  select(Taxa, Sample_date, Smoothed_Relative_Abundance) %>% 

  # Extract seasonal data from STL time series, per taxa
  group_by(Taxa) %>%
  do({
    ts_data <- ts(.$Smoothed_Relative_Abundance, start = c(2017, 8), frequency = 52)  # Weekly frequency
    stl_decomposition <- stl(ts_data, s.window = "periodic")
    data.frame(Sample_date = .$Sample_date, Seasonal = stl_decomposition$time.series[, "seasonal"])
  }) %>%
  ungroup() %>%
  spread(key = Taxa, value = Seasonal) %>%
  drop_na()

seasonal_data

correlation_matrix <- cor(seasonal_data %>% select(-Sample_date), use = "pairwise.complete.obs")

# Convert the correlation matrix to a tidy data frame
correlation_data <- as.data.frame(as.table(correlation_matrix)) %>%
  rename(Taxa1 = Var1, Taxa2 = Var2, Correlation = Freq) %>%
  filter(Taxa1 != Taxa2) %>%        # Remove self-correlations
  filter(Correlation > 0.5)    # Filter for strong correlations

correlation_data %>% 
  ggplot() +
  geom_raster(aes(Taxa1, Taxa2, fill = Correlation)) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), name = "Correlation") +
  theme_minimal() +
  labs(x = "Taxa 2", y = "Taxa 1", title = "Correlation between Taxa") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)
```


# With lags

```{r}
# Assuming 'seasonal_data' is your dataframe containing the seasonal trends
# and that it has been processed to include only relevant columns

# Convert to time series format if not already done
# Replace 'Taxa1', 'Taxa2' with the actual column names of your seasonal data
taxa1_series <- ts(seasonal_data$`Clytia gregaria`, frequency = 52)
taxa2_series <- ts(seasonal_data$`Mitrocoma cellularia`, frequency = 52)

# Compute cross-correlation function
ccf_result <- ccf(taxa1_series, taxa2_series, lag.max = 4, plot = TRUE, na.action = na.pass)

# To extract significant lags
# You can define a threshold for significance based on your data
# Here, we consider absolute values greater than 0.5 as significant (adjust as needed)
significant_lags <- which(abs(ccf_result$acf) > 0.5)

# Print the significant lags and their corresponding values
significant_ccf_values <- ccf_result$acf[significant_lags]
significant_lags_df <- data.frame(Lag = ccf_result$lag[significant_lags],
                                   Correlation = significant_ccf_values)

print(significant_lags_df)
```
# CFF for all taxa 

```{r}
seasonal_data
```


```{r}
# Assuming 'seasonal_data' is your dataframe containing the seasonal trends
# Ensure your data frame is in the correct format
seasonal_data_long <- seasonal_data %>% 
  select(-Sample_date) %>% 
  pivot_longer(cols = everything(), names_to = "Taxa", values_to = "Relative_Abundance")

seasonal_data_long %>% 
  filter(Taxa == "Acartia longiremis")
```


```{r}
# Create a list to store the results
results_list <- list()

# Get unique taxa names
unique_taxa <- unique(seasonal_data_long$Taxa)

# Loop through all pairs of taxa
for (i in 1:(length(unique_taxa) - 1)) {
  for (j in (i + 1):length(unique_taxa)) {
    taxa1 <- unique_taxa[i]
    taxa2 <- unique_taxa[j]
    
    # Extract time series data for the two taxa
    taxa1_series <- seasonal_data_long %>% 
      filter(Taxa == taxa1) %>% 
      pull(Relative_Abundance) %>% 
      ts(frequency = 52)
    
    taxa2_series <- seasonal_data_long %>% 
      filter(Taxa == taxa2) %>% 
      pull(Relative_Abundance) %>% 
      ts(frequency = 52)
    
    # Compute cross-correlation function
    ccf_result <- ccf(taxa1_series, taxa2_series, lag.max = 4, plot = FALSE, na.action = na.pass)
    
    # Extract significant lags
    significant_lags <- which(ccf_result$acf > 0.6)
    significant_ccf_values <- ccf_result$acf[significant_lags]
    
    # Store results in a data frame
    if (length(significant_lags) > 0) {
      results_df <- data.frame(
        Taxa1 = taxa1,
        Taxa2 = taxa2,
        Lag = ccf_result$lag[significant_lags],
        Correlation = significant_ccf_values
      )
      results_list[[length(results_list) + 1]] <- results_df
    }
  }
}

# Combine all results into a single data frame
final_results <- bind_rows(results_list)

# View results
print(final_results)

final_results %>% 
  group_by(Taxa1, Taxa2) %>% 
  filter(Correlation == max(Correlation)) %>% 
  ggplot() +
  geom_raster(aes(Taxa1, Taxa2, fill = Correlation)) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), name = "Correlation") +
  theme_minimal() +
  labs(x = "Taxa 2", y = "Taxa 1", title = "Correlation between Taxa") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)
```


