---
title: "R Notebook"
---


```{r}
# Data import

library(readxl)
#library(phyloseq)


# All data format modifications and filtering and plotting
library(tidyverse)
'%!in%' <- function(x,y)!('%in%'(x,y))
source("Code/Functions.R")
library(zoo)

# Statistical tools
library(vegan)

# Specific plotting tools
library(ggpubr)
library(paletteer)
library(viridis)
library(cowplot)
#library(ggnewscale)
source("Code/scales_and_themes.R")

# Make sure tidyverse is the latest loaded package
detach("package:tidyverse", unload=TRUE)
library(tidyverse)


# Custom color scale for depths
depth_colors <- c("0" = "#add8e6",    # Light blue for 0 m
                  "5" = "#87ceeb",    # Sky blue for 5 m
                  "30" = "#4682b4",   # Steel blue for 30 m
                  "100" = "#00008b")  # Dark blue for 100 m
```


```{r}
Dataset_COI <- readRDS("Data_output/Data_merge/COI_glomm_indexed.rds")
Dataset_18S <- readRDS("Data_output/Data_merge/pr2_18S_glomm_indexed.rds")
Dataset_Bottle <- readRDS("Data_output/Data_merge/bottle_file.rds")

Dataset_18S_glom <- Dataset_18S %>%
    filter(Phylum != "Metazoa") #%>%
#    filter(Phylum %in% c(
#    "Chlorophyta",
#    "Cryptophyta",
#    "Dinoflagellata",
#    "Haptophyta",
#    "Ochrophyta"
#  )) %>% 
  #filter(is.na(Genus)==FALSE) %>% 
  group_by(Library_ID, Sample_date, Line_out_depth, Taxa = Class) %>%
  summarise(Abundance = sum(Abundance)) %>% 
  index_RA(Library_ID, Taxa, Abundance, Sample_date, Line_out_depth)
  
  
taxatab_18S <- Dataset_18S %>%
  group_by(Phylum, Class, Order, Family, Genus, Taxa) %>% 
  summarise()

Taxtab_X <- taxatab_18S %>% 
  group_by(PhylumX = Phylum, ClassX = Class, OrderX = Order,
           FamilyX = Family, GenusX = Genus, TaxaX = Taxa) %>% 
  summarise()
Taxtab_Y <- taxatab_18S %>% 
  group_by(PhylumY = Phylum, ClassY = Class, OrderY = Order,
           FamilyY = Family, GenusY = Genus, TaxaY = Taxa) %>% 
  summarise()

taxatab_18S
  
```



# 1. Interpolation

	1.	Fill Missing Combinations with Zeros: The complete(Sample_ID, Sample_date, Taxa, Sampling_Depth, fill = list(Relative_Abundance = 0)) line ensures that any combination without observed data is filled with 0 for Relative_Abundance.
	2.	Generate Weekly Dates: complete(Sample_date = seq.Date(min(Sample_date), max(Sample_date), by = "week")) generates weekly dates for each Taxa and Sampling_Depth combination.
	3.	Interpolate Missing Values: na.approx() fills in any NAs generated by the new weekly dates in Sample_date.


### COI
```{r}

start_date <- as.Date("2017-01-01")
end_date <- as.Date("2019-02-15")


# Complete the data frame with all combinations of Sample_ID, Sample_date, Taxa, and Sampling_Depth
complete_data <- Dataset_COI %>%
  filter(Line_out_depth %in% c(0,5,30,100)) %>%
  mutate(Line_out_depth = factor(Line_out_depth, levels = c("0", "5", "30", "100"))) %>% 
  mutate(Sample_date = as.Date(Sample_date)) %>% 
  filter(Sample_date >= start_date & Sample_date <= end_date) %>% 
  complete(Sample_date, Taxa, Line_out_depth, fill = list(Abundance = 0)) %>% 
  # Ensure dates are properly ordered after completion
  arrange(Sample_date, Taxa, Line_out_depth)


strict_weekly_dates <- seq.Date(from = min(complete_data$Sample_date),
                                to = max(complete_data$Sample_date),
                                by = "week")

smoothed_data <- complete_data %>%
  # Expand to include all combinations of Taxa, Depth, and weekly dates
  group_by(Taxa, Line_out_depth) %>%
  complete(Sample_date = strict_weekly_dates) %>%
  arrange(Sample_date) %>%
  # Interpolate missing values in RA_Index within each group
  mutate(Abundance = ifelse(is.na(Abundance),
                           na.approx(Abundance, rule = 2, na.rm = FALSE),
                           Abundance)) %>%
  # Apply 2-week rolling mean for smoothing (7-day intervals)
  mutate(Smoothed_Relative_Abundance = rollmean(Abundance, k = 2, fill = NA, align = "center")) %>%
  ungroup() %>%
  # Filter to include only rows with Smoothed_Relative_Abundance calculated
  filter(!is.na(Smoothed_Relative_Abundance)) %>% 
  filter(Sample_date %in% strict_weekly_dates)

Interpolated_COI_Data <- smoothed_data %>% 
  group_by(Taxa, Sample_date) %>% 
  summarise(Abundance = sum(Smoothed_Relative_Abundance)/4)
```

### 18S

```{r}


# Complete the data frame with all combinations of Sample_ID, Sample_date, Taxa, and Sampling_Depth
complete_data <- Dataset_18S_glom %>%
  filter(Line_out_depth %in% c(0,5,30,100)) %>%
  mutate(Line_out_depth = factor(Line_out_depth, levels = c("0", "5", "30", "100"))) %>%
  #filter(Phylum != "Metazoa") %>%
  #  filter(Phylum %in% c(
  #  "Chlorophyta",
  #  "Cryptophyta",
  #  "Dinoflagellata",
  #  "Haptophyta",
  #  "Ochrophyta"
  #)) %>% 
  #filter(is.na(Genus)==FALSE) %>% 
  mutate(Sample_date = as.Date(Sample_date)) %>% 
  complete(Sample_date, Taxa, Line_out_depth, fill = list(Abundance = 0)) %>%
  filter(Sample_date >= start_date & Sample_date <= end_date) %>% 
  # Ensure dates are properly ordered after completion
  arrange(Sample_date, Taxa, Line_out_depth)


# NOTE: Commented out to reuse the series for COI
#strict_weekly_dates <- seq.Date(from = min(complete_data$Sample_date),
#                                to = max(complete_data$Sample_date),
#                                by = "week")

smoothed_data <- complete_data %>%
  # Expand to include all combinations of Taxa, Depth, and weekly dates
  group_by(Taxa, Line_out_depth) %>%
  complete(Sample_date = strict_weekly_dates) %>%
  arrange(Sample_date) %>%
  # Interpolate missing values in RA_Index within each group
  mutate(Abundance = ifelse(is.na(Abundance),
                           na.approx(Abundance, rule = 2, na.rm = FALSE),
                           Abundance)) %>%
  # Apply 2-week rolling mean for smoothing (7-day intervals)
  mutate(Smoothed_Relative_Abundance = rollmean(Abundance, k = 4, fill = NA, align = "center")) %>%
  ungroup() %>%
  # Filter to include only rows with Smoothed_Relative_Abundance calculated
  filter(!is.na(Smoothed_Relative_Abundance)) %>% 
  filter(Sample_date %in% strict_weekly_dates)

Interpolated_18S_Data <- smoothed_data %>% 
  group_by(Taxa, Sample_date) %>% 
  summarise(Abundance = sum(Smoothed_Relative_Abundance)/4)
```



# 2. STL - Extract seasonal data
```{r}

### COI
seasonal_data_COI <- Interpolated_COI_Data %>%
  # Extract seasonal data from STL time series, per taxa
  group_by(Taxa) %>%
  do({
    ts_data <- ts(.$Abundance, start = c(2017, 8), frequency = 52)  # Weekly frequency
    stl_decomposition <- stl(ts_data, s.window = "periodic")
    data.frame(Sample_date = .$Sample_date, Seasonal = stl_decomposition$time.series[, "seasonal"])
  }) %>%
  ungroup() %>%
  spread(key = Taxa, value = Seasonal) %>%
  drop_na()


#### 18S
seasonal_data_18S <- Interpolated_18S_Data %>%
  # Extract seasonal data from STL time series, per taxa
  group_by(Taxa) %>%
  do({
    ts_data <- ts(.$Abundance, start = c(2017, 8), frequency = 52)  # Weekly frequency
    stl_decomposition <- stl(ts_data, s.window = "periodic")
    data.frame(Sample_date = .$Sample_date, Seasonal = stl_decomposition$time.series[, "seasonal"])
  }) %>%
  ungroup() %>%
  spread(key = Taxa, value = Seasonal) %>%
  drop_na()

### Combine
seasonal_data_all <-left_join(seasonal_data_18S, seasonal_data_COI, by = "Sample_date")
seasonal_data_all %>% 
  colnames()


```

# 3. Correlation
```{r}

correlation_matrix <- cor(seasonal_data_all %>% select(-Sample_date), use = "pairwise.complete.obs")

correlation_matrix

# Convert the correlation matrix to a tidy data frame
correlation_data <- as.data.frame(as.table(correlation_matrix)) %>%
  rename(Taxa1 = Var1, Taxa2 = Var2, Correlation = Freq) %>%
  filter(Taxa1 != Taxa2) %>%        # Remove self-correlations
  filter(Correlation > 0.5) %>%     # Filter for strong correlations
  filter(Taxa1 %in% Interpolated_COI_Data$Taxa,
         Taxa2 %in% Interpolated_18S_Data$Taxa)



correlation_data %>% 
  ggplot() +
  geom_raster(aes(Taxa1, Taxa2, fill = Correlation)) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), name = "Correlation") +
  theme_minimal() +
  labs(x = "Taxa 2", y = "Taxa 1", title = "Correlation between Taxa") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)
```

# Correlation without lag

```{r}
correlation_matrix_all <- cor(seasonal_data_all %>% select(-Sample_date), use = "pairwise.complete.obs")
```




```{r}
# Define vectors for filtering
phototrophic_plankton <- c(
  "Chlorarachniophyceae", "Chlorodendrophyceae", "Chlorophyceae", "Mamiellophyceae", 
  "Nephroselmidophyceae", "Pedinophyceae", "Prasino-Clade-VIII", "Pyramimonadophyceae", 
  "Trebouxiophyceae", "Ulvophyceae", "Cryptophyceae", "Dinophyceae", "Noctilucophyceae", 
  "Syndiniales", "Prymnesiophyceae", "Bacillariophyta", "Bolidophyceae", "Chrysophyceae", 
  "Dictyochophyceae", "Pelagophyceae", "Phaeophyceae", "Raphidophyceae", "Perkinsida", 
  "Prasinodermophyceae", "Florideophyceae", "Embryophyceae"
)

heterotrophic_protists <- c(
  "Ellobiopsidae", "Gregarinomorphea", "Apusomonadidae_Group-1", "Filosa", 
  "Filosa-Granofilosea", "Filosa-Imbricatea", "Filosa-Thecofilosea", "Phaeodarea", 
  "Choanoflagellatea", "CONTH_7", "CONTH_8", "CONThreeP", "Cyclotrichium_like_organism", 
  "Litostomatea", "Nassophorea", "Oligohymenophorea", "Phyllopharyngea", 
  "Prostomatea_1", "Spirotrichea", "Ascomycota", "Basidiomycota", "Chytridiomycota", 
  "Mucoromycota", "Haptophyta_Clade_HAP2", "Haptophyta_Clade_HAP3", "Haptophyta_Clade_HAP4", 
  "Haptophyta_Clade_HAP5", "Planomonadida", "Katablepharidaceae", "Tubulinea", "Filasterea", 
  "Ichthyosporea", "MOCH-1", "MOCH-3", "MOCH-4", "RAD-A", "RAD-B", "RAD-C", "Rhodelphea"
)

metazoan_zooplankton <- c(
  "Annelida", "Arthropoda", "Brachiopoda", "Bryozoa", "Cnidaria", "Craniata", 
  "Ctenophora", "Echinodermata", "Mollusca", "Nemertea", "Porifera", "Rotifera", 
  "Urochordata"
)

microbial_eukaryotes <- c(
  "Endomyxa", "Endomyxa-Phytomyxea", "Katablepharidaceae", "Hyphochytriomyceta", "Oomycota", 
  "Pirsonia_Clade", "Polycystinea", "Acantharea", "Labyrinthulomycetes", "MAST-10", 
  "MAST-12", "MAST-3", "MAST-4", "MAST-6", "MAST-7", "MAST-8", "MAST-9", "MAST-25", 
  "Stramenopiles_X-Group-2", "Stramenopiles_X-Group-6"
)

marine_viruses_symbionts <- c(
  "Picozoa_X", "Prasinodermophyceae", "Telonemia_X", "Syndiniales"
)

miscellaneous_unclassified <- c(
  "CONThreeP", "MOCH-4", "Rhodelphea", "Filasterea"
)
```


Zooplankton to Phytoplankton
```{r}
TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% Dataset_COI$Taxa,
                                         colnames(correlation_matrix_all) %in% TaxaX]
# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]

pdf(file = "Figure_output/Corr_Plots/Zooplankton_Phytoplankton.pdf", width = 20, height = 20)
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Phytoplankton"
)

dev.off()
```
Zooplankton to Heterotrophic 
```{r}
TaxaX <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% Dataset_COI$Taxa,
                                         colnames(correlation_matrix_all) %in% TaxaX]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]


pdf(file = "Figure_output/Corr_Plots/Zooplankton_Microplankton.pdf", width = 20, height = 20)
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Phytoplankton"
)

dev.off()
```
Heterotrophic to phytoplankton
```{r}
TaxaY <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% TaxaY,
                                         colnames(correlation_matrix_all) %in% TaxaX]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]

pdf(file = "Figure_output/Corr_Plots/Microplankton_Phytoplankton.pdf", width = 20, height = 20)
# Create the heatmap with a custom color scale
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Phytoplankton"
)

dev.off()
```



Phytoplankton
```{r}
TaxaY <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% TaxaX,
                                         colnames(correlation_matrix_all) %in% TaxaX]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]


pdf(file = "Figure_output/Corr_Plots/Phytoplankton.pdf", width = 20, height = 20)
# Create the heatmap with a custom color scale
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Phytoplankton"
)

dev.off()
```

Zooplankton
```{r}
TaxaY <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% Dataset_COI$Taxa,
                                         colnames(correlation_matrix_all) %in% Dataset_COI$Taxa]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]

pdf(file = "Figure_output/Corr_Plots/Zooplankton.pdf", width = 20, height = 20)
# Create the heatmap with a custom color scale
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Zooplankton"
)

dev.off()
```

Microplankton
```{r}
TaxaY <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% TaxaY,
                                         colnames(correlation_matrix_all) %in% TaxaY]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]


pdf(file = "Figure_output/Corr_Plots/Microplankton.pdf", width = 20, height = 20)
# Create the heatmap with a custom color scale
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Heterotrophic Protozooplankton"
)

dev.off()
```

```{r}

# Example data (10x10 matrix)
set.seed(123)
data_matrix <- matrix(runif(100, min = 0, max = 1), nrow = 10)

# Create the heatmap with a custom color scale
pheatmap(data_matrix,
         color = colorRampPalette(c("white", "red"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Heatmap with Custom Color Scale"
)
```







#### OLD

Phototrophic taxa only
```{r}


TaxaX <- taxatab_18S %>%
  #filter(is.na(Genus) == FALSE) %>% 
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

TaxaX


correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% TaxaX,
                                         colnames(correlation_matrix_all) %in% TaxaX]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]

correlation_matrix

heatmap(correlation_matrix)

hclust(as.dist(1 - correlation_matrix)) %>% plot()

# Convert the correlation matrix to a tidy data frame and reorder factors
correlation_data <- as.data.frame(as.table(correlation_matrix)) %>%
  rename(TaxaX = Var1, TaxaY = Var2, Correlation = Freq) %>%
  mutate(
    TaxaX = factor(TaxaX, levels = ordered_taxa),
    TaxaY = factor(TaxaY, levels = ordered_taxa)
  ) %>% 
  #filter(TaxaX != TaxaY) %>%        # Remove self-correlations
  filter(Correlation > 0.6)     # Filter for strong correlations



correlation_data %>%
  ggplot() +
  geom_raster(aes(TaxaX, TaxaY, fill = Correlation)) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
                       midpoint = 0.5, limit = c(0, 1), name = "Correlation") +
  theme_minimal() +
  labs(x = "Taxa X", y = "Taxa Y", title = "Correlation between Taxa") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)

taxatab_18S %>% 
  pull(Class) %>% unique()
```



Phytoplankton
```{r}
correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% c(phytoplankton, mixotrophs),
                                         colnames(correlation_matrix_all) %in% c(phytoplankton, mixotrophs)]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]

heatmap(correlation_matrix)
```

Zooplankton
```{r}
correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% zooplankton,
                                         colnames(correlation_matrix_all) %in% zooplankton]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]

heatmap(correlation_matrix)
```











Ochrophyta
```{r}
TaxaX <- taxatab_18S %>%
  #filter(is.na(Genus) == FALSE) %>% 
  filter(Phylum == "Dinoflagellata") %>% 
  pull(Taxa)

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% TaxaX,
                                         colnames(correlation_matrix_all) %in% TaxaX]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]

heatmap(correlation_matrix)

taxatab_18S
```









# 4. Correlation with lags



```{r}

```






```{r}
# Define the matrix (excluding date column)
species_matrix <- as.matrix(seasonal_data_all %>% select(-Sample_date))
species_matrix
# Define maximum lag
max_lag <- 3

# Prepare an empty data frame to store results
ccf_results <- data.frame(
  Species1 = character(),
  Species2 = character(),
  Max_Correlation = numeric(),
  Lag_At_Max_Correlation = integer()
)


# Loop through each pair of species
species_names <- colnames(species_matrix)
for (i in 1:(ncol(species_matrix) - 1)) {
  for (j in (i + 1):ncol(species_matrix)) {
    
    # Get time series data for the two species
    ts1 <- species_matrix[, 1]
    ts2 <- species_matrix[, 1]
    
    # Compute cross-correlation
    ccf_result <- ccf(ts1, ts2, lag.max = max_lag, plot = FALSE)
    
    # Find the lag with the highest absolute cross-correlation
    max_corr_idx <- which.max(abs(ccf_result$acf))
    max_corr <- ccf_result$acf[max_corr_idx]
    lag_at_max_corr <- ccf_result$lag[max_corr_idx]
    
    # Store the result
    ccf_results <- rbind(ccf_results, data.frame(
      Species1 = species_names[i],
      Species2 = species_names[j],
      Max_Correlation = max_corr,
      Lag_At_Max_Correlation = lag_at_max_corr
    ))
  }
}

ccf_results
```


### Zoop vs Microzoo

```{r}
corr_data <- ccf_results %>%
  dplyr::filter(Species2 %in% Interpolated_COI_Data$Taxa,
         Species1 %in% Interpolated_18S_Data$Taxa) %>% 
  filter(Max_Correlation > 0.4,
         Lag_At_Max_Correlation >= 0) %>% 
  left_join(taxatab_18S, by = c("Species1"))


corr_data %>% 
  ggplot() +
  geom_raster(aes(Species2, Species1, fill = Max_Correlation)) +
  geom_text(aes(Species2, Species1, label = Lag_At_Max_Correlation), size = 2) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
                       midpoint = 0.4, limit = c(0.4, 1), name = "Correlation") +
  theme_minimal(8) +
  labs(x = "Zooplankton", y = "Microeukaryotes", title = "Correlation between Taxa with 0-4 weeks lag") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)

ggsave("Figure_output/Correlation_plot.pdf")

```
### Phyto vs Phyto

```{r}
taxatab_18S <- Dataset_18S %>%
  group_by(Phylum, Species1 = Class) %>% 
  summarise()

corr_data <- ccf_results %>%
  dplyr::filter(Species2 %in% Interpolated_18S_Data$Taxa,
         Species1 %in% Interpolated_18S_Data$Taxa) %>% 
  filter(Max_Correlation > 0.4,
         Lag_At_Max_Correlation >= 0) %>% 
  left_join(taxatab_18S, by = c("Species1"))


corr_data %>% 
  ggplot() +
  geom_raster(aes(Species2, Species1, fill = Max_Correlation)) +
  geom_text(aes(Species2, Species1, label = Lag_At_Max_Correlation), size = 2) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
                       midpoint = 0.4, limit = c(0.4, 1), name = "Correlation") +
  theme_minimal(8) +
  labs(x = "Zooplankton", y = "Microeukaryotes", title = "Correlation between Taxa with 0-4 weeks lag") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)

ggsave("Figure_output/Correlation_plot.pdf")

```

```{r}
taxatab_18S <- Dataset_18S %>%
  group_by(Phylum, Species1 = Class) %>% 
  summarise()

corr_data <- ccf_results %>%
  dplyr::filter(Species2 %in% Interpolated_COI_Data$Taxa,
         Species1 %in% Interpolated_COI_Data$Taxa) %>% 
  filter(Max_Correlation > 0.4,
         Lag_At_Max_Correlation >= 0) %>% 
  left_join(taxatab_18S, by = c("Species1"))


corr_data %>% 
  ggplot() +
  geom_raster(aes(Species2, Species1, fill = Max_Correlation)) +
  geom_text(aes(Species2, Species1, label = Lag_At_Max_Correlation), size = 2) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
                       midpoint = 0.4, limit = c(0.4, 1), name = "Correlation") +
  theme_minimal(8) +
  labs(x = "Zooplankton", y = "Microeukaryotes", title = "Correlation between Taxa with 0-4 weeks lag") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)

ggsave("Figure_output/Correlation_plot.pdf")

```




# 5. Correlation by sample

```{r}

all_community_mat <- bind_rows(
  filter(Dataset_18S, Hakai_ID %in% Dataset_COI$Hakai_ID),
  filter(Dataset_COI, Hakai_ID %in% Dataset_COI$Hakai_ID)) %>% 
  group_by(Hakai_ID, Taxa) %>%
  summarize(RA = mean(RA)) %>% 
  pivot_wider(names_from = Taxa, values_from = RA) %>% 
  column_to_rownames("Hakai_ID") %>%
  mutate(across(everything(), ~ replace_na(.x, 0))) %>% 
  as.matrix()

all_community_mat

all_cor_matrix <- cor(all_community_mat, method = "pearson", use = "pairwise.complete.obs")

cor_df <- as.data.frame(all_cor_matrix) %>%
  rownames_to_column(var = "Taxa1") %>%
  pivot_longer(-Taxa1, names_to = "Taxa2", values_to = "Correlation") %>%
  filter(Taxa1 %!in% Dataset_COI$Taxa,
         Taxa2 %!in% Dataset_18S$Taxa)

cor_df %>%
  filter(Correlation > 0.4) %>% 
  ggplot() +
  geom_raster(aes(Taxa2, Taxa1, fill = Correlation)) +
  #geom_text(aes(Taxa2, Taxa1, label = Lag_At_Max_Correlation), size = 2) +
  #scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
  #                     midpoint = 0.4, limit = c(0.4, 1), name = "Correlation") +
  theme_minimal(8) +
  labs(x = "Zooplankton", y = "Microeukaryotes", title = "Correlation between Taxa with 0-4 weeks lag") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)
  
cor_df
```



