---
title: "R Notebook"
---

# 0. Preparations and data import

Read in libraries:
```{r}
# Data import
library(readxl)
#library(phyloseq)

# All data format modifications and filtering and plotting
library(tidyverse)
'%!in%' <- function(x,y)!('%in%'(x,y))
#source("Code/Functions.R")

# Statistical tools
library(vegan)
library(zoo)

# Specific plotting tools
#library(ggpubr)
#library(paletteer)
#library(viridis)
#library(cowplot)
#library(ggnewscale)
#source("Code/scales_and_themes.R")

# Make sure tidyverse is the latest loaded package
detach("package:tidyverse", unload=TRUE)
library(tidyverse)

# Network analysis
library(igraph)
library(tidygraph)
library(ggraph)
library(pheatmap)
```
```{r}
# Functional groups
functional_groups <- c(
  "Autotrophic_protist",
  "Mixotrophic_protist",
  "Heterotrophic_protist",
  "Omnivorous_zooplankton",
  "Carnivorous_zooplankton",
  "Pelagic_fish"
)

# Corresponding colors
colors <- c(
  "Autotrophic_protist" = "#66C2A5",
  "Mixotrophic_protist" = "#FC8D62",
  "Heterotrophic_protist" = "#FFD92F",
  "Omnivorous_zooplankton" = "#8DA0CB",
  "Carnivorous_zooplankton" = "#3288BD",
  "Pelagic_fish" = "#E78AC3"
)

Metrics_levels <- c(
  "In_Degree",
  "Out_Degree",
  "Degree_Centrality",
  "Closeness_Centrality",
  "Betweenness_Centrality",
  "Eigenvector",
  "Trophic_level",
  "Num_Paths"
)
```	


Read in Hakai Datasets
```{r}
# These datasets have been created in the "2_Merge_COI_18S_Env.Rmd" file

Dataset_COI <- readRDS("Data_output/Data_merge/COI_glomm_indexed.rds")
Dataset_18S <- readRDS("Data_output/Data_merge/pr2_18S_glomm_indexed.rds")
Dataset_12S <- readRDS("Data_output/Data_merge/12S_glomm_indexed.rds")
#Dataset_Bottle <- readRDS("Data_output/Data_merge/bottle_file.rds")
Functional_Traits <- read_excel("Data_import/Functional_Traits.xlsx")
```

Extract some taxonomy tables, used to construct trait dataset
```{r}
taxatab_18S <- Dataset_18S %>%
  group_by(Phylum, Class, Order, Family, Genus, Taxa) %>% 
  summarise()

taxatab_COI <- Dataset_COI %>%
  group_by(Phylum, Class, Order, Family, Genus, Taxa) %>% 
  summarise()

taxtab_12S <- Dataset_12S %>% 
  group_by(Class, Taxa) %>% 
  summarise() %>% 
  mutate(Phylum = "Chordata")

taxtab_12S %>% 
  write_csv("Data_output/taxtab_12S.csv")

taxatab_COI %>% 
  write_csv("Data_output/taxtab_COI.csv")

rm(taxatab_18S, taxatab_COI, taxtab_12S)
```



# 1. Time Series Interpolation

	1.	Fill Missing Combinations with Zeros: The complete(Sample_ID, Sample_date, Taxa, Sampling_Depth, fill = list(Relative_Abundance = 0)) line ensures that any combination without observed data is filled with 0 for Relative_Abundance.
	2.	Generate Weekly Dates: complete(Sample_date = seq.Date(min(Sample_date), max(Sample_date), by = "week")) generates weekly dates for each Taxa and Sampling_Depth combination.
	3.	Interpolate Missing Values: na.approx() fills in any NAs generated by the new weekly dates in Sample_date.

Interpolate COI
```{r}

start_date <- as.Date("2017-01-01")
end_date <- as.Date("2019-02-15")


# Complete the data frame with all combinations of Sample_ID, Sample_date, Taxa, and Sampling_Depth
complete_data <- Dataset_COI %>%
  filter(Line_out_depth %in% c(0,5,30,100)) %>%
  mutate(Line_out_depth = factor(Line_out_depth, levels = c("0", "5", "30", "100"))) %>% 
  mutate(Sample_date = as.Date(Sample_date)) %>% 
  filter(Sample_date >= start_date & Sample_date <= end_date) %>% 
  complete(Sample_date, Taxa, Line_out_depth, fill = list(Abundance = 0)) %>% 
  # Ensure dates are properly ordered after completion
  arrange(Sample_date, Taxa, Line_out_depth)


strict_weekly_dates <- seq.Date(from = min(complete_data$Sample_date),
                                to = max(complete_data$Sample_date),
                                by = "week")

smoothed_data <- complete_data %>%
  # Expand to include all combinations of Taxa, Depth, and weekly dates
  group_by(Taxa, Line_out_depth) %>%
  complete(Sample_date = strict_weekly_dates) %>%
  arrange(Sample_date) %>%
  # Interpolate missing values in RA_Index within each group
  mutate(Abundance = ifelse(is.na(Abundance),
                           na.approx(Abundance, rule = 2, na.rm = FALSE),
                           Abundance)) %>%
  # Apply 2-week rolling mean for smoothing (7-day intervals)
  mutate(Smoothed_Relative_Abundance = rollmean(Abundance, k = 2, fill = NA, align = "center")) %>%
  ungroup() %>%
  # Filter to include only rows with Smoothed_Relative_Abundance calculated
  filter(!is.na(Smoothed_Relative_Abundance)) %>% 
  filter(Sample_date %in% strict_weekly_dates)

Interpolated_COI_Data <- smoothed_data %>% 
  group_by(Taxa, Sample_date) %>% 
  summarise(Abundance = sum(Smoothed_Relative_Abundance)/4)
```

Interpolate 12S
```{r}

start_date <- as.Date("2017-01-01")
end_date <- as.Date("2019-02-15")


# Complete the data frame with all combinations of Sample_ID, Sample_date, Taxa, and Sampling_Depth
complete_data <- Dataset_12S %>%
  filter(Line_out_depth %in% c(0,5,30,100)) %>%
  mutate(Line_out_depth = factor(Line_out_depth, levels = c("0", "5", "30", "100"))) %>% 
  mutate(Sample_date = as.Date(Sample_date)) %>% 
  #filter(Sample_date >= start_date & Sample_date <= end_date) %>% 
  complete(Sample_date, Taxa, Line_out_depth, fill = list(Abundance = 0)) %>% 
  # Ensure dates are properly ordered after completion
  arrange(Sample_date, Taxa, Line_out_depth)


#strict_weekly_dates <- seq.Date(from = min(complete_data$Sample_date),
#                                to = max(complete_data$Sample_date),
#                                by = "week")

smoothed_data <- complete_data %>%
  # Expand to include all combinations of Taxa, Depth, and weekly dates
  group_by(Taxa, Line_out_depth) %>%
  complete(Sample_date = strict_weekly_dates) %>%
  arrange(Sample_date) %>%
  # Interpolate missing values in RA_Index within each group
  mutate(Abundance = ifelse(is.na(Abundance),
                           na.approx(Abundance, rule = 2, na.rm = FALSE),
                           Abundance)) %>%
  # Apply 2-week rolling mean for smoothing (7-day intervals)
  mutate(Smoothed_Relative_Abundance = rollmean(Abundance, k = 2, fill = NA, align = "center")) %>%
  ungroup() %>%
  # Filter to include only rows with Smoothed_Relative_Abundance calculated
  filter(!is.na(Smoothed_Relative_Abundance)) %>% 
  filter(Sample_date %in% strict_weekly_dates)

Interpolated_12S_Data <- smoothed_data %>% 
  group_by(Taxa, Sample_date) %>% 
  summarise(Abundance = sum(Smoothed_Relative_Abundance)/4)




```

Interpolate 18S: 
```{r}

# Complete the data frame with all combinations of Sample_ID, Sample_date, Taxa, and Sampling_Depth
complete_data <- Dataset_18S %>%
  filter(Line_out_depth %in% c(0,5,30,100)) %>%
  mutate(Line_out_depth = factor(Line_out_depth, levels = c("0", "5", "30", "100"))) %>%
  #filter(Phylum != "Metazoa") %>%
  #  filter(Phylum %in% c(
  #  "Chlorophyta",
  #  "Cryptophyta",
  #  "Dinoflagellata",
  #  "Haptophyta",
  #  "Ochrophyta"
  #)) %>% 
  #filter(is.na(Genus)==FALSE) %>% 
  mutate(Sample_date = as.Date(Sample_date)) %>% 
  complete(Sample_date, Taxa, Line_out_depth, fill = list(Abundance = 0)) %>%
  filter(Sample_date >= start_date & Sample_date <= end_date) %>%
  #filter(Sample_date <= end_date) %>%
  # Ensure dates are properly ordered after completion
  arrange(Sample_date, Taxa, Line_out_depth)


# NOTE: Commented out to reuse the series for COI
strict_weekly_dates <- seq.Date(from = min(complete_data$Sample_date),
                                to = max(complete_data$Sample_date),
                                by = "week")

smoothed_data <- complete_data %>%
  # Expand to include all combinations of Taxa, Depth, and weekly dates
  group_by(Taxa, Line_out_depth) %>%
  complete(Sample_date = strict_weekly_dates) %>%
  arrange(Sample_date) %>%
  # Interpolate missing values in RA_Index within each group
  mutate(Abundance = ifelse(is.na(Abundance),
                           na.approx(Abundance, rule = 2, na.rm = FALSE),
                           Abundance)) %>%
  # Apply 2-week rolling mean for smoothing (7-day intervals)
  mutate(Smoothed_Relative_Abundance = rollmean(Abundance, k = 4, fill = NA, align = "center")) %>%
  ungroup() %>%
  # Filter to include only rows with Smoothed_Relative_Abundance calculated
  filter(!is.na(Smoothed_Relative_Abundance)) %>% 
  filter(Sample_date %in% strict_weekly_dates)

Interpolated_18S_Data <- smoothed_data %>% 
  group_by(Taxa, Sample_date) %>% 
  summarise(Abundance = sum(Smoothed_Relative_Abundance)/4)
```

Check the limits of interpolated datasets:

```{r}
Interpolated_COI_Data$Sample_date %>% min()
Interpolated_COI_Data$Sample_date %>% max()

Interpolated_12S_Data$Sample_date %>% min()
Interpolated_12S_Data$Sample_date %>% max()

Interpolated_18S_Data$Sample_date %>% min()
Interpolated_18S_Data$Sample_date %>% max()
```


# 2. STL - Extract seasonal data
```{r}
### COI
seasonal_data_COI <- Interpolated_COI_Data %>%
  # Extract seasonal data from STL time series, per taxa
  group_by(Taxa) %>%
  do({
    ts_data <- ts(.$Abundance, start = c(2017, 08), frequency = 52)  # Weekly frequency
    stl_decomposition <- stl(ts_data, s.window = "periodic")
    data.frame(Sample_date = .$Sample_date, Seasonal = stl_decomposition$time.series[, "seasonal"])
  }) %>%
  ungroup() %>%
  spread(key = Taxa, value = Seasonal) %>%
  drop_na()


#### 18S
seasonal_data_18S <- Interpolated_18S_Data %>%
  # Extract seasonal data from STL time series, per taxa
  group_by(Taxa) %>%
  do({
    ts_data <- ts(.$Abundance, start = c(2017, 15), frequency = 52)  # Weekly frequency
    stl_decomposition <- stl(ts_data, s.window = "periodic")
    data.frame(Sample_date = .$Sample_date, Seasonal = stl_decomposition$time.series[, "seasonal"])
  }) %>%
  ungroup() %>%
  spread(key = Taxa, value = Seasonal) %>%
  drop_na()


#### 12S
seasonal_data_12S <- Interpolated_12S_Data %>%
  filter(Sample_date != "2017-01-08") %>% 
  # Extract seasonal data from STL time series, per taxa
  group_by(Taxa) %>%
  do({
    ts_data <- ts(.$Abundance, start = c(2017, 08), frequency = 52)  # Weekly frequency
    stl_decomposition <- stl(ts_data, s.window = "periodic")
    data.frame(Sample_date = .$Sample_date, Seasonal = stl_decomposition$time.series[, "seasonal"])
  }) %>%
  ungroup() %>%
  spread(key = Taxa, value = Seasonal) %>%
  drop_na()


### Combine
start_date <- as.Date("2017-01-15")
end_date <- as.Date("2019-01-13")

seasonal_data_all <-left_join(seasonal_data_18S, seasonal_data_COI, by = "Sample_date") %>%
  left_join(seasonal_data_12S, by = "Sample_date") %>% 
  filter(Sample_date >= start_date & Sample_date <= end_date)
seasonal_data_all

## seasonal_data_all will be used for correlation analysis later.
```

Test matrix of just TS data

```{r}
#weekly_data_all <- pivot_wider(Interpolated_12S_Data, names_from = Taxa, values_from = Abundance) %>% 
#  left_join(pivot_wider(Interpolated_18S_Data, names_from = Taxa, values_from = Abundance)) %>% 
#  left_join(pivot_wider(Interpolated_COI_Data, names_from = Taxa, values_from = Abundance)) %>% 
#  filter(Sample_date != "2017-01-08", 
#         Sample_date != "2019-01-20",
#         Sample_date != "2019-01-27",
#         Sample_date != "2019-02-03")

#weekly_data_all
```

Remove unwanted objects from interpolation and STL steps:
```{r}
# Remove Intermediate Variables
rm(complete_data, smoothed_data, end_date, start_date, strict_weekly_dates)
# Remove Original Data
rm(Dataset_12S, Dataset_18S, Dataset_COI)
# Interpolated Data
rm(Interpolated_12S_Data, Interpolated_18S_Data, Interpolated_COI_Data)
# STL Seasonal Data
rm(seasonal_data_COI, seasonal_data_18S, seasonal_data_12S, smoothed_data
```


# 3. Construct Trait Based Metaweb

Constructing a trait based metaweb using size and feeding assumptions:

Parasites and symbionts are excluded from the analysis, and so are fungi,
and streptophytes (land plants).

- Protozoan predators are non-metazoan eukaryotes with known heterotrophy.
- They can also be mixotrophs, but not purely autotrophs.
- They prey on any genus that atr in all size classes lower than themselves, either heterotrophic or phototrophic.

```{r}
# Protozoan Predators:
ppredator <- Functional_Traits %>% 
  filter(Genus != "NA",
         is.na(Parasitism),
         is.na(Symbiont),
         Heterotrophy == 1,
         Phylum %!in% c("Fungi"),
         !is.na(Sizeclass)) %>% 
  select(Predator = Taxa, Predator_size = Sizeclass)
## Prey
pprey <- Functional_Traits %>% 
  filter(Genus != "NA",
         is.na(Parasitism),
         is.na(Symbiont),
         Phylum %!in% c("Fungi"),
         !is.na(Sizeclass)) %>% 
  select(Prey = Taxa, Prey_size = Sizeclass)
  
PZP <- tidyr::crossing(ppredator, pprey) %>% 
  filter(Predator_size > Prey_size)
```

- carnivorous zooplankton, are based on trait data
- they prey on mesoplankton (or micro-meso protozoans),
but only those 1.5 size classes below, (or smaller) than themselves.

```{r}
# Carnivorous zooplankton
carnpred <- Functional_Traits %>% 
  filter(TrophicState == "carnivore") %>% 
  select(Predator = Taxa, Predator_size = Sizeclass)

## Prey
carnpreyd <- Functional_Traits %>% 
  filter(Genus != "NA",
         is.na(Parasitism),
         is.na(Symbiont),
         Phylum %!in% c("Fungi"),
         !is.na(Sizeclass)) %>% 
  filter(Sizeclass > 3) %>% 
  select(Prey = Taxa, Prey_size = Sizeclass)

CZP <- tidyr::crossing(carnpred, carnpreyd) %>% 
  filter((Predator_size - Prey_size) > 1)
```
- omnivorous zooplankton, are based on trait data
- they prey on any plankton in the nano range and up, 
but only those 1 size classes below, (or smaller) than themselves.
```{r}
# Omnivorous zooplankton
omnpred <- Functional_Traits %>% 
  filter(TrophicState == "omnivore") %>% 
  select(Predator = Taxa, Predator_size = Sizeclass)

## Prey
omnprey <- Functional_Traits %>% 
  filter(Genus != "NA",
         is.na(Parasitism),
         is.na(Symbiont),
         Phylum %!in% c("Fungi"),
         !is.na(Sizeclass)) %>% 
  filter(Sizeclass > 1) %>% 
  select(Prey = Taxa, Prey_size = Sizeclass)

OZP <- tidyr::crossing(omnpred, omnprey) %>% 
  filter((Predator_size - Prey_size) > 0.5)
```
- pelagic fish eat zooplankton...

```{r}
pelagic_fish <- c(
  "Alosa sapidissima", "Ammodytes personatus", "Anoplopoma fimbria", 
  "Boreogadus saida", "Clupea pallasii", "Diaphus theta", 
  "Engraulis mordax", "Gasterosteus aculeatus", "Mallotus villosus", 
  "Merluccius productus", "Oncorhynchus gorbuscha", "Oncorhynchus keta", 
  "Oncorhynchus kisutch", "Oncorhynchus nerka", "Oncorhynchus tshawytscha", 
  "Osmerus mordax", "Sardinops sagax", "Stenobrachius leucopsarus", 
  "Stenobrachius nannochir", "Thaleichthys pacificus", "Zaprora silenus")

fishpred <- Functional_Traits %>% 
  ungroup() %>%
  filter(Taxa %in% pelagic_fish) %>% 
  select(Predator = Taxa)

fishprey <- Functional_Traits %>% 
  filter(TrophicState %in% c("carnivore", "omnivore")) %>% 
  select(Prey = Taxa)

FISH <- tidyr::crossing(fishpred, fishprey)

```

```{r}
METAWEB <- bind_rows(PZP, CZP, OZP, FISH) %>% 
  select(Predator, Prey)

METAWEB
# METAWEB is a adjacency dataframe that will be used later to guide correlation analysis.
```

```{r}
rm(carnpred, carnpreyd, CZP, FISH, fishpred, fishprey, omnpred, omnprey,
   OZP, ppredator, pprey, PZP)
```


# 4. Correlation matrixes:


## a. Without lags
Auto-correlations with no lags of all involved species.
```{r}
correlation_matrix_all <- cor(seasonal_data_all %>%
                                select(-Sample_date), use = "pairwise.complete.obs")
```

## b. With lags
Testing only METAWEB interactions:

```{r}
pw_ccf <- function(mat, adj.df, max_lag = 3, min_corr = -1) {
  # Initialize results data frame
  results <- data.frame(Predator = character(), Prey = character(),
                        Max_Correlation = numeric(), Lag = integer(), stringsAsFactors = FALSE)
  
  # Loop through each predator-prey pair in adj.df
  for (i in seq_len(nrow(adj.df))) {
    pred <- adj.df$Predator[i]
    prey_taxa <- adj.df$Prey[i]
    
    # Extract time series for predator and prey as numeric vectors
    ts_pred <- mat[, pred]
    ts_prey <- mat[, prey_taxa]
    
    # Check if both species exist in mat
    if (!is.null(ts_pred) && !is.null(ts_prey)) {
      # Compute cross-correlation without plotting
      ccf_result <- ccf(ts_prey, ts_pred, lag.max = max_lag, plot = FALSE)
      
      # Check for valid cross-correlation values
      if (!is.null(ccf_result$acf) && length(ccf_result$acf) > 0 && !all(is.na(ccf_result$acf))) {
        max_corr_idx <- which.max(ccf_result$acf)
        max_corr <- ccf_result$acf[max_corr_idx]
        lag_at_max_corr <- ccf_result$lag[max_corr_idx]
        
        # Filter for significant correlations and relevant lags
        if (!is.na(max_corr) && lag_at_max_corr <= 0 && max_corr >= min_corr) {
          results <- rbind(results, data.frame(Predator = pred, Prey = prey_taxa,
                                               Max_Correlation = max_corr, Lag = lag_at_max_corr))
        }
      }
    } else {print("Warning, names of metaweb not in time series")}
  }
  
  return(results)
}
```


```{r}
Correlation_lags <- pw_ccf(seasonal_data_all, METAWEB, max_lag = 3, min_corr = -1)


# plot correlation heatmap:



pdf(file = "Figure_output/All_Metaweb_Correlations.pdf", width = 10, height = 10)
# Create the heatmap with a custom color scale
Correlation_lags %>%
  select(!Lag) %>%
  #filter(Max_Correlation > 0.5) %>% 
  # To matrix
  pivot_wider(names_from = "Predator", values_from = "Max_Correlation", values_fill = 0) %>% 
  column_to_rownames("Prey") %>% as.matrix() %>%
pheatmap(#color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         #breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "All Network Correlations",
         fontsize = 3.5
)
dev.off()


```



# 5. Prepare Network

1. Use correlation data to build a tidygraph.

```{r}
CORR_WEB <- Correlation_lags %>% 
  select(Prey, Predator, Max_Correlation, Lag) %>% 
  as_tbl_graph(, directed = TRUE)
```

2. Import Node data from trait dataset

```{r}
# Trait information per functional group
Autotrophic_protist <- Functional_Traits %>% 
  filter(Phototrophy == 1,
         is.na(Heterotrophy)) %>% 
  mutate(FG = "Autotrophic_protist")

Heterotrophic_protist <- Functional_Traits %>% 
  filter(Heterotrophy == 1,
         is.na(Phototrophy)) %>% 
  mutate(FG = "Heterotrophic_protist")

Mixotrophic_protist <- Functional_Traits %>% 
  filter(#Taxa %in% pull(CORR_WEB, name),
         Heterotrophy == 1,
         Phototrophy == 1) %>% 
  mutate(FG = "Mixotrophic_protist")

Omnivorous_zooplankton <- Functional_Traits %>% 
  filter(#Taxa %in% pull(CORR_WEB, name),
         TrophicState == "omnivore") %>% 
  mutate(FG = "Omnivorous_zooplankton")

Carnivorous_zooplankton <- Functional_Traits %>% 
  filter(#Taxa %in% pull(CORR_WEB, name),
         TrophicState == "carnivore") %>% 
  mutate(FG = "Carnivorous_zooplankton")

Pelagic_fish <- Functional_Traits %>% 
  filter(#Taxa %in% pull(CORR_WEB, name),
         TrophicState == "fish") %>% 
  mutate(FG = "Pelagic_fish")

# Information about abundance peaks from TS data
peaks <- seasonal_data_all %>%
  filter(year(Sample_date) == 2017) %>% 
  pivot_longer(!1, names_to = "Taxa", values_to = "signal") %>% 
  group_by(Taxa) %>%
  slice_max(signal, with_ties = FALSE) %>% 
  ungroup() %>% 
  arrange(Sample_date) %>% 
  #filter(Taxa %in% colnames(adj_matrix)) %>%
  select(Taxa, Date_max = Sample_date)

# Merge all Node data
Node_Data <- bind_rows(Autotrophic_protist,
          Heterotrophic_protist,
          Mixotrophic_protist,
          Omnivorous_zooplankton,
          Carnivorous_zooplankton,
          Pelagic_fish) %>% 
  left_join(peaks, by = "Taxa") %>%
  filter(Taxa %in% pull(CORR_WEB, name)) %>% 
  select(name = Taxa, FG, Sizeclass, SIzeRange, Date_max,
         Phylum, Class, Order, Family, Genus,
         Phototrophy, Heterotrophy, Mixotrophy, HAB)

rm(Autotrophic_protist, Heterotrophic_protist, Mixotrophic_protist,
   Omnivorous_zooplankton, Carnivorous_zooplankton, Pelagic_fish,
   peaks, pelagic_fish, Het_prot)
```

3. Add node data to network
```{r}
CORR_WEB_NODETRAITS <- CORR_WEB %>% 
  
  # Add node data
  activate(nodes) %>% 
  left_join(Node_Data, by = "name") %>%
  
  # Filter only positive strong correlations
  activate(edges) %>% 
  filter(Max_Correlation > 0.5) %>%
  
  # Order FGs
  activate(nodes) %>%
  mutate(FG = factor(FG, levels = functional_groups)) %>%
  # Remove nodes with no links:
  mutate(Total_Degree = centrality_degree(mode = "total")) %>% 
  filter(Total_Degree > 0) %>% 
  
  # Calculate Basig Node Metrics:
  mutate(
    In_Degree = centrality_degree(mode = "in"),
    Betweenness = centrality_betweenness(),
    Closeness = centrality_closeness(),
    Eigenvector = centrality_eigen(),
    Clustering = local_transitivity(),
    Coreness = node_coreness()
  )
```

Common Node-Specific Metrics in tidygraph
	•	Degree centrality: Number of edges connected to a node.
	•	Betweenness centrality: Measures the frequency a node is part of the shortest paths between other nodes.
	•	Closeness centrality: Inverse of the average distance to all other nodes.
	•	Eigenvector centrality: Measures influence based on the connections of a node’s neighbors.
	•	PageRank: Google’s algorithm for ranking webpages, adapted for graphs.
	•	Trophic position: For directed graphs, measures trophic level.
	
3. Add trophic level as node trait:

```{r}
# Identify Basal nodes:
basal_nodes <- CORR_WEB_NODETRAITS %>% 
  filter(Phototrophy == 1) %>% pull(name)

# List heterotrophic protists.
Het_prot <- CORR_WEB_NODETRAITS %>% 
  filter(FG == "Heterotrophic_protist") %>% 
  pull(name)


# 1. Calculate distances to basal nodes
distances_to_base_1 <- CORR_WEB_NODETRAITS %>%
  # 1.1. initial caclulation of distance to base
  as.igraph() %>%
  # Calculate distance
  distances(v= basal_nodes, mode = "out")

  # Summarize per node column
trophic_levels <- apply(distances_to_base_1, 2, function(x) {
  if (all(is.infinite(x))) {
    return(NA)
  } else {
    return(mean(x[!is.infinite(x)], na.rm = TRUE))
  }}) %>%
  
  # 1.2 Correct for Het_prot with no base, and iterate calculation 
  # Create a base known that is unknown, that het prots with no other source is linked to.

  enframe() %>%
  mutate(unknown_base =
           ifelse(name %in% Het_prot & is.na(value), 1, Inf)) %>% 
  select(name, unknown_base) %>% 
  pivot_wider(names_from = name, values_from = unknown_base) %>%
  mutate(basal_nodes = "unknown_base") %>% 
  column_to_rownames("basal_nodes") %>% 
  
  bind_rows(as.data.frame(distances_to_base_1)) %>% 
  as.matrix() %>% 
  apply(2, function(x) {
  if (all(is.infinite(x))) {
    return(NA)
  } else {
    return(mean(x[!is.infinite(x)], na.rm = TRUE))
  }}) %>%
  enframe(name = "name",
          value = "Trophic_Level")

CORR_WEB_NODETRAITS_TL <- CORR_WEB_NODETRAITS %>%
  left_join(trophic_levels, by = "name")
```
	
	
## Extra: Plot Seasonal Data
	
```{r}
seasonal_data_all %>% 
  pivot_longer(!1, names_to = "name", values_to = "RA") %>%
  group_by(name) %>% 
  mutate(RA = RA / max(RA)) %>% 
  filter(year(Sample_date) == 2017) %>%
  filter(name %in% Node_Data$name) %>% 
  left_join(Node_Data, by = "name") %>% 
  mutate(Month = month(Date_max)) %>%
  mutate(FG = factor(FG, levels = functional_groups)) %>%
  ggplot() +
  geom_line(aes(Sample_date, RA,  group = name, colour = FG), alpha = 0.9,
            linewidth = 0.3) +
  facet_wrap("Month", scales = "free", ncol = 3) +
  scale_color_manual(values = colors) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave("Figure_output/All_Seasonality.pdf", height = 7, width = 10)
```	

```{r}
rm(CORR_WEB, CORR_WEB_NODETRAITS)
rm(METAWEB, Correlation_lags, correlation_matrix_all, Node_Data, trophic_levels,
   distances_to_base_1, basal_nodes)
```

# 6. Food Web Visualizations:

## 6.1 Stress based layout 


Plot stress layout with Short Random Walk Communities
```{r}
CORR_WEB_NODETRAITS_TL %>%
  mutate(group = factor(group_walktrap())) %>%
  mutate(Community = factor(ifelse(group %in% c(1,2,3,4,5,6),
                        group, "Other"))) %>% 
  ggraph("stress") + 
  geom_edge_link(arrow = arrow(length = unit(3, 'mm')),
                 #aes(alpha = Max_Correlation)
                 alpha = 0.2) +
  geom_node_point(aes(colour = Community)) +
  #geom_node_text(aes(label = name), angle = 90, size = 2.5, vjust = 2) +
  theme_graph() 

ggsave("Figure_output/Network_Stress_Community.png", height = 7, width = 10)
```

Plot stress layout with Functional Groups
```{r}
CORR_WEB_NODETRAITS_TL %>%
  ggraph("stress") + 
  geom_edge_link(arrow = arrow(length = unit(3, 'mm')),
                 #aes(alpha = Max_Correlation)
                 alpha = 0.2) +
  geom_node_point(aes(colour = FG)) +
  #geom_node_text(aes(label = name), angle = 90, size = 2.5, vjust = 2) +
  theme_graph() +
  scale_color_manual(values = colors)

ggsave("Figure_output/Network_Stress_FG.png", height = 7, width = 10)
```

## 6.2 Stress based plotting 

Plot trophic level / seasonality with functional groups
```{r}
# Create a manual layout using layout_tbl_graph_manual
layout_manual <- CORR_WEB_NODETRAITS_TL %>%
  # Adjust TL 1 to include some variation for visibility
  mutate(TL = ifelse(Trophic_Level == 1,
                     Trophic_Level - (1/Sizeclass),
                     Trophic_Level)) %>% 
  as_tibble() %>%
  select(name, y_position = TL, x_position = Date_max) %>% 
  mutate(x_position = week(x_position))

CORR_WEB_NODETRAITS_TL %>%
  ggraph(x=layout_manual$x_position, y=layout_manual$y_position) + 
  geom_edge_link(arrow = arrow(length = unit(3, 'mm')),
                 #aes(alpha = Max_Correlation)
                 alpha = 0.2) +
  geom_node_point(aes(colour = FG), size = 2) +
  #geom_node_text(aes(label = name), angle = 90, size = 2.5, vjust = 2) +
  theme_graph() +
  scale_color_manual(values = colors)

ggsave("Figure_output/Network_All_MAX_TL.png", height = 7, width = 10)
```


```{r}
CORR_WEB_NODETRAITS_TL %>%
  mutate(group = factor(group_walktrap())) %>%
  mutate(Community = factor(ifelse(group %in% c(1,2,3,4,5,6),
                        group, "Other"))) %>% 
  ggraph(x=layout_manual$x_position, y=layout_manual$y_position) + 
  geom_edge_link(arrow = arrow(length = unit(3, 'mm')),
                 #aes(alpha = Max_Correlation)
                 alpha = 0.2) +
  geom_node_point(aes(colour = Community), size = 2) +
  #geom_node_text(aes(label = name), angle = 90, size = 2.5, vjust = 2) +
  theme_graph()

ggsave("Figure_output/Network_Trophic_Community.png", height = 7, width = 10)
```


## 6.3 Node level metrics

```{r}
CORR_WEB_NODETRAITS_TL %>% 
  as.tibble() %>% 
  select(name, FG, Date_max, In_Degree, Betweenness, Closeness, Eigenvector, Clustering, Trophic_Level) %>% 
  
  pivot_longer(!1:3, names_to = "Metric", values_to = "Value") %>% 
  group_by(FG, Metric) %>% 
  reframe(Value, Value_Norm = Value/sum(Value, na.rm = TRUE),
            name, Date_max) %>% 
  group_by(Metric, FG) %>% 
  filter(sum(Value, na.rm = TRUE) != 0) %>% 
  ggplot() +
  geom_point(aes(Date_max, Value, colour = FG)) +
  geom_smooth(aes(Date_max, Value, colour = FG), se = FALSE, linewidth = 0.5) +
  facet_wrap("Metric", scales = "free") +
  scale_color_manual(values = colors) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        #strip.text = element_text(face = "italic"),
        panel.background = element_rect(fill = '#F5F5F5', colour = 'darkgrey'),
        #legend.position="none",,
        aspect.ratio = 1/1.5)
```
