---
title: "R Notebook"
---


```{r}
# Data import

library(readxl)
#library(phyloseq)


# All data format modifications and filtering and plotting
library(tidyverse)
'%!in%' <- function(x,y)!('%in%'(x,y))
source("Code/Functions.R")
library(zoo)

# Statistical tools
library(vegan)

# Specific plotting tools
library(ggpubr)
library(paletteer)
library(viridis)
library(cowplot)
#library(ggnewscale)
source("Code/scales_and_themes.R")

# Make sure tidyverse is the latest loaded package
detach("package:tidyverse", unload=TRUE)
library(tidyverse)


# Custom color scale for depths
depth_colors <- c("0" = "#add8e6",    # Light blue for 0 m
                  "5" = "#87ceeb",    # Sky blue for 5 m
                  "30" = "#4682b4",   # Steel blue for 30 m
                  "100" = "#00008b")  # Dark blue for 100 m
```


```{r}
Dataset_COI <- readRDS("Data_output/Data_merge/COI_glomm_indexed.rds")
Dataset_18S <- readRDS("Data_output/Data_merge/pr2_18S_glomm_indexed.rds")
Dataset_12S <- readRDS("Data_output/Data_merge/12S_glomm_indexed.rds")
Dataset_Bottle <- readRDS("Data_output/Data_merge/bottle_file.rds")

Dataset_18S_glom <- Dataset_18S %>%
    filter(Phylum != "Metazoa") #%>%
#    filter(Phylum %in% c(
#    "Chlorophyta",
#    "Cryptophyta",
#    "Dinoflagellata",
#    "Haptophyta",
#    "Ochrophyta"
#  )) %>% 
  #filter(is.na(Genus)==FALSE) %>% 
  group_by(Library_ID, Sample_date, Line_out_depth, Taxa = Class) %>%
  summarise(Abundance = sum(Abundance)) %>% 
  index_RA(Library_ID, Taxa, Abundance, Sample_date, Line_out_depth)
  
  
taxatab_18S <- Dataset_18S %>%
  group_by(Phylum, Class, Order, Family, Genus, Taxa) %>% 
  summarise()

Dataset_18S

Taxtab_X <- taxatab_18S %>% 
  group_by(PhylumX = Phylum, ClassX = Class, OrderX = Order,
           FamilyX = Family, GenusX = Genus, TaxaX = Taxa) %>% 
  summarise()
Taxtab_Y <- taxatab_18S %>% 
  group_by(PhylumY = Phylum, ClassY = Class, OrderY = Order,
           FamilyY = Family, GenusY = Genus, TaxaY = Taxa) %>% 
  summarise()

```



# 1. Interpolation

	1.	Fill Missing Combinations with Zeros: The complete(Sample_ID, Sample_date, Taxa, Sampling_Depth, fill = list(Relative_Abundance = 0)) line ensures that any combination without observed data is filled with 0 for Relative_Abundance.
	2.	Generate Weekly Dates: complete(Sample_date = seq.Date(min(Sample_date), max(Sample_date), by = "week")) generates weekly dates for each Taxa and Sampling_Depth combination.
	3.	Interpolate Missing Values: na.approx() fills in any NAs generated by the new weekly dates in Sample_date.


### COI
```{r}

start_date <- as.Date("2017-01-01")
end_date <- as.Date("2019-02-15")


# Complete the data frame with all combinations of Sample_ID, Sample_date, Taxa, and Sampling_Depth
complete_data <- Dataset_COI %>%
  filter(Line_out_depth %in% c(0,5,30,100)) %>%
  mutate(Line_out_depth = factor(Line_out_depth, levels = c("0", "5", "30", "100"))) %>% 
  mutate(Sample_date = as.Date(Sample_date)) %>% 
  filter(Sample_date >= start_date & Sample_date <= end_date) %>% 
  complete(Sample_date, Taxa, Line_out_depth, fill = list(Abundance = 0)) %>% 
  # Ensure dates are properly ordered after completion
  arrange(Sample_date, Taxa, Line_out_depth)


strict_weekly_dates <- seq.Date(from = min(complete_data$Sample_date),
                                to = max(complete_data$Sample_date),
                                by = "week")

smoothed_data <- complete_data %>%
  # Expand to include all combinations of Taxa, Depth, and weekly dates
  group_by(Taxa, Line_out_depth) %>%
  complete(Sample_date = strict_weekly_dates) %>%
  arrange(Sample_date) %>%
  # Interpolate missing values in RA_Index within each group
  mutate(Abundance = ifelse(is.na(Abundance),
                           na.approx(Abundance, rule = 2, na.rm = FALSE),
                           Abundance)) %>%
  # Apply 2-week rolling mean for smoothing (7-day intervals)
  mutate(Smoothed_Relative_Abundance = rollmean(Abundance, k = 2, fill = NA, align = "center")) %>%
  ungroup() %>%
  # Filter to include only rows with Smoothed_Relative_Abundance calculated
  filter(!is.na(Smoothed_Relative_Abundance)) %>% 
  filter(Sample_date %in% strict_weekly_dates)

Interpolated_COI_Data <- smoothed_data %>% 
  group_by(Taxa, Sample_date) %>% 
  summarise(Abundance = sum(Smoothed_Relative_Abundance)/4)
```

### 12S

```{r}

start_date <- as.Date("2017-01-01")
end_date <- as.Date("2019-02-15")


# Complete the data frame with all combinations of Sample_ID, Sample_date, Taxa, and Sampling_Depth
complete_data <- Dataset_12S %>%
  filter(Line_out_depth %in% c(0,5,30,100)) %>%
  mutate(Line_out_depth = factor(Line_out_depth, levels = c("0", "5", "30", "100"))) %>% 
  mutate(Sample_date = as.Date(Sample_date)) %>% 
  #filter(Sample_date >= start_date & Sample_date <= end_date) %>% 
  complete(Sample_date, Taxa, Line_out_depth, fill = list(Abundance = 0)) %>% 
  # Ensure dates are properly ordered after completion
  arrange(Sample_date, Taxa, Line_out_depth)


#strict_weekly_dates <- seq.Date(from = min(complete_data$Sample_date),
#                                to = max(complete_data$Sample_date),
#                                by = "week")

smoothed_data <- complete_data %>%
  # Expand to include all combinations of Taxa, Depth, and weekly dates
  group_by(Taxa, Line_out_depth) %>%
  complete(Sample_date = strict_weekly_dates) %>%
  arrange(Sample_date) %>%
  # Interpolate missing values in RA_Index within each group
  mutate(Abundance = ifelse(is.na(Abundance),
                           na.approx(Abundance, rule = 2, na.rm = FALSE),
                           Abundance)) %>%
  # Apply 2-week rolling mean for smoothing (7-day intervals)
  mutate(Smoothed_Relative_Abundance = rollmean(Abundance, k = 2, fill = NA, align = "center")) %>%
  ungroup() %>%
  # Filter to include only rows with Smoothed_Relative_Abundance calculated
  filter(!is.na(Smoothed_Relative_Abundance)) %>% 
  filter(Sample_date %in% strict_weekly_dates)

Interpolated_12S_Data <- smoothed_data %>% 
  group_by(Taxa, Sample_date) %>% 
  summarise(Abundance = sum(Smoothed_Relative_Abundance)/4)




```


### 18S

```{r}


# Complete the data frame with all combinations of Sample_ID, Sample_date, Taxa, and Sampling_Depth
complete_data <- Dataset_18S_glom %>%
  filter(Line_out_depth %in% c(0,5,30,100)) %>%
  mutate(Line_out_depth = factor(Line_out_depth, levels = c("0", "5", "30", "100"))) %>%
  #filter(Phylum != "Metazoa") %>%
  #  filter(Phylum %in% c(
  #  "Chlorophyta",
  #  "Cryptophyta",
  #  "Dinoflagellata",
  #  "Haptophyta",
  #  "Ochrophyta"
  #)) %>% 
  #filter(is.na(Genus)==FALSE) %>% 
  mutate(Sample_date = as.Date(Sample_date)) %>% 
  complete(Sample_date, Taxa, Line_out_depth, fill = list(Abundance = 0)) %>%
  filter(Sample_date >= start_date & Sample_date <= end_date) %>%
  #filter(Sample_date <= end_date) %>%
  # Ensure dates are properly ordered after completion
  arrange(Sample_date, Taxa, Line_out_depth)


# NOTE: Commented out to reuse the series for COI
strict_weekly_dates <- seq.Date(from = min(complete_data$Sample_date),
                                to = max(complete_data$Sample_date),
                                by = "week")

smoothed_data <- complete_data %>%
  # Expand to include all combinations of Taxa, Depth, and weekly dates
  group_by(Taxa, Line_out_depth) %>%
  complete(Sample_date = strict_weekly_dates) %>%
  arrange(Sample_date) %>%
  # Interpolate missing values in RA_Index within each group
  mutate(Abundance = ifelse(is.na(Abundance),
                           na.approx(Abundance, rule = 2, na.rm = FALSE),
                           Abundance)) %>%
  # Apply 2-week rolling mean for smoothing (7-day intervals)
  mutate(Smoothed_Relative_Abundance = rollmean(Abundance, k = 4, fill = NA, align = "center")) %>%
  ungroup() %>%
  # Filter to include only rows with Smoothed_Relative_Abundance calculated
  filter(!is.na(Smoothed_Relative_Abundance)) %>% 
  filter(Sample_date %in% strict_weekly_dates)

Interpolated_18S_Data <- smoothed_data %>% 
  group_by(Taxa, Sample_date) %>% 
  summarise(Abundance = sum(Smoothed_Relative_Abundance)/4)
```

```{r}
Interpolated_COI_Data$Sample_date %>% min()
Interpolated_COI_Data$Sample_date %>% max()

Interpolated_12S_Data$Sample_date %>% min()
Interpolated_12S_Data$Sample_date %>% max()

Interpolated_18S_Data$Sample_date %>% min()
Interpolated_18S_Data$Sample_date %>% max()
```


# 2. STL - Extract seasonal data
```{r}

### COI
seasonal_data_COI <- Interpolated_COI_Data %>%
  filter(Sample_date != "2017-01-08") %>%
  
  # Extract seasonal data from STL time series, per taxa
  group_by(Taxa) %>%
  do({
    ts_data <- ts(.$Abundance, start = c(2017, 08), frequency = 52)  # Weekly frequency
    stl_decomposition <- stl(ts_data, s.window = "periodic")
    data.frame(Sample_date = .$Sample_date, Seasonal = stl_decomposition$time.series[, "seasonal"])
  }) %>%
  ungroup() %>%
  spread(key = Taxa, value = Seasonal) %>%
  drop_na()


#### 18S
seasonal_data_18S <- Interpolated_18S_Data %>%
  # Extract seasonal data from STL time series, per taxa
  group_by(Taxa) %>%
  do({
    ts_data <- ts(.$Abundance, start = c(2017, 15), frequency = 52)  # Weekly frequency
    stl_decomposition <- stl(ts_data, s.window = "periodic")
    data.frame(Sample_date = .$Sample_date, Seasonal = stl_decomposition$time.series[, "seasonal"])
  }) %>%
  ungroup() %>%
  spread(key = Taxa, value = Seasonal) %>%
  drop_na()


#### 12S
seasonal_data_12S <- Interpolated_12S_Data %>%
  filter(Sample_date != "2017-01-08") %>% 
  # Extract seasonal data from STL time series, per taxa
  group_by(Taxa) %>%
  do({
    ts_data <- ts(.$Abundance, start = c(2017, 08), frequency = 52)  # Weekly frequency
    stl_decomposition <- stl(ts_data, s.window = "periodic")
    data.frame(Sample_date = .$Sample_date, Seasonal = stl_decomposition$time.series[, "seasonal"])
  }) %>%
  ungroup() %>%
  spread(key = Taxa, value = Seasonal) %>%
  drop_na()


start_date <- as.Date("2017-01-15")
end_date <- as.Date("2019-01-13")

### Combine
seasonal_data_all <-left_join(seasonal_data_18S, seasonal_data_COI, by = "Sample_date") %>%
  left_join(seasonal_data_12S, by = "Sample_date") %>% 
  filter(Sample_date >= start_date & Sample_date <= end_date)
seasonal_data_all
  

```

# 3. Construct correlation matrixes:

### a. Without lags
```{r}

correlation_matrix <- cor(seasonal_data_all %>% select(-Sample_date), use = "pairwise.complete.obs")

correlation_matrix

# Convert the correlation matrix to a tidy data frame
correlation_data <- as.data.frame(as.table(correlation_matrix)) %>%
  rename(Taxa1 = Var1, Taxa2 = Var2, Correlation = Freq) %>%
  filter(Taxa1 != Taxa2) %>%        # Remove self-correlations
  filter(Correlation > 0.5) %>%     # Filter for strong correlations
  filter(Taxa1 %in% Interpolated_COI_Data$Taxa,
         Taxa2 %in% Interpolated_18S_Data$Taxa)



correlation_data %>% 
  ggplot() +
  geom_raster(aes(Taxa1, Taxa2, fill = Correlation)) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), name = "Correlation") +
  theme_minimal() +
  labs(x = "Taxa 2", y = "Taxa 1", title = "Correlation between Taxa") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)
```

Correlation without lag

```{r}
correlation_matrix_all <- cor(seasonal_data_all %>% select(-Sample_date), use = "pairwise.complete.obs")
correlation_matrix_all %>% colnames()
```


### b. With Lags


```{r}
species_matrix <- as.matrix(seasonal_data_all %>% select(-Sample_date))

species_matrix

# Initialize results data frame
results <- data.frame(Species1 = character(), Species2 = character(),
                      Max_Correlation = numeric(), Lag = integer(), stringsAsFactors = FALSE)
max_lag <- 3

for (i in 1:(ncol(species_matrix))) {
  for (j in 1:ncol(species_matrix)) {
    ts1 <- species_matrix[, i]
    ts2 <- species_matrix[, j]
    
    # Compute cross-correlation without plotting
    ccf_result <- ccf(ts1, ts2, lag.max = max_lag, plot = FALSE)
    
    # Check for valid cross-correlation values
    if (!is.null(ccf_result$acf) && length(ccf_result$acf) > 0 && !all(is.na(ccf_result$acf))) {
      max_corr_idx <- which.max(abs(ccf_result$acf))
      max_corr <- ccf_result$acf[max_corr_idx]
      lag_at_max_corr <- ccf_result$lag[max_corr_idx]
      
      # Store only significant correlations if they meet the threshold
      if (!is.na(max_corr) && abs(max_corr) > 0.5) {
        results <- rbind(results, data.frame(Species1 = species_names[i],
                                             Species2 = species_names[j],
                                             Max_Correlation = max_corr,
                                             Lag = lag_at_max_corr))
      }
    }
  }
}

results

# Create an adjacency matrix based on significant correlations
adj_matrix <- with(results, {
  mat <- matrix(0, ncol = length(species_names), nrow = length(species_names),
                dimnames = list(species_names, species_names))
  for (k in 1:nrow(results)) {
    mat[results$Species1[k], results$Species2[k]] <- results$Max_Correlation[k]
  }
  return(mat)
})
```


# 4. Define vectors for filtering

```{r}
# Define vectors for filtering
phototrophic_plankton <- c(
  "Chlorarachniophyceae", "Chlorodendrophyceae", "Chlorophyceae", "Mamiellophyceae", 
  "Nephroselmidophyceae", "Pedinophyceae", "Prasino-Clade-VIII", "Pyramimonadophyceae", 
  "Trebouxiophyceae", "Ulvophyceae", "Cryptophyceae", "Dinophyceae", "Noctilucophyceae", 
  "Syndiniales", "Prymnesiophyceae", "Bacillariophyta", "Bolidophyceae", "Chrysophyceae", 
  "Dictyochophyceae", "Pelagophyceae", "Phaeophyceae", "Raphidophyceae", "Perkinsida", 
  "Prasinodermophyceae", "Florideophyceae", "Embryophyceae"
)

heterotrophic_protists <- c(
  "Ellobiopsidae", "Gregarinomorphea", "Apusomonadidae_Group-1", "Filosa", 
  "Filosa-Granofilosea", "Filosa-Imbricatea", "Filosa-Thecofilosea", "Phaeodarea", 
  "Choanoflagellatea", "CONTH_7", "CONTH_8", "CONThreeP", "Cyclotrichium_like_organism", 
  "Litostomatea", "Nassophorea", "Oligohymenophorea", "Phyllopharyngea", 
  "Prostomatea_1", "Spirotrichea", "Ascomycota", "Basidiomycota", "Chytridiomycota", 
  "Mucoromycota", "Haptophyta_Clade_HAP2", "Haptophyta_Clade_HAP3", "Haptophyta_Clade_HAP4", 
  "Haptophyta_Clade_HAP5", "Planomonadida", "Katablepharidaceae", "Tubulinea", "Filasterea", 
  "Ichthyosporea", "MOCH-1", "MOCH-3", "MOCH-4", "RAD-A", "RAD-B", "RAD-C", "Rhodelphea"
)

carnivorous_zooplankton <- c(
  "Aetideus divergens", "Cyanea capillata", "Euphausia pacifica", 
  "Euphysa flammea", "Merluccius angustimanus", "Nanomia bijuga", 
  "Pandalus stenolepis", "Thysanoessa spinifera", "Clytia hemisphaerica",
  "Obelia dichotoma", "Epilabidocera amphitrites", "Pleurobrachia bachei"
  
)

# Vector for Omnivorous/Filter Feeding Zooplankton
omnivorous_filter_feeding_zooplankton <- c(
  "Acartia longiremis", "Calanus pacificus", "Centropages abdominalis", 
  "Corycaeus anglicus", "Discoconchoecia elegans",  
  "Eucalanus bungii", "Evadne nordmanni", "Metridia pacifica", 
  "Microcalanus pygmaeus", "Oithona atlantica", 
  "Paracalanus indicus", "Paracalanus parvus",  
  "Pseudocalanus acuspes", "Pseudocalanus mimus", "Pseudocalanus minutus", 
  "Pseudocalanus newmani", "Triconia borealis"
)






microbial_eukaryotes <- c(
  "Endomyxa", "Endomyxa-Phytomyxea", "Katablepharidaceae", "Hyphochytriomyceta", "Oomycota", 
  "Pirsonia_Clade", "Polycystinea", "Acantharea", "Labyrinthulomycetes", "MAST-10", 
  "MAST-12", "MAST-3", "MAST-4", "MAST-6", "MAST-7", "MAST-8", "MAST-9", "MAST-25", 
  "Stramenopiles_X-Group-2", "Stramenopiles_X-Group-6"
)

marine_viruses_symbionts <- c(
  "Picozoa_X", "Prasinodermophyceae", "Telonemia_X", "Syndiniales"
)

miscellaneous_unclassified <- c(
  "CONThreeP", "MOCH-4", "Rhodelphea", "Filasterea"
)

Dataset_COI$Taxa %>% unique()


```


# 5. PLot Heatmaps

## Zooplankton to Phytoplankton
```{r}
adj_matrix

TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()



correlation_matrix <- v[rownames(adj_matrix) %in% Dataset_COI$Taxa,
                                         colnames(adj_matrix) %in% TaxaX]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]

pdf(file = "Figure_output/Corr_Plots/Metazooplankton_Phytoplankton.pdf", width = 20, height = 20)
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Metazooplankton - Phytoplankon lagged interactions"
)
dev.off()

```
## Zooplankton to Heterotrophic 
```{r}
TaxaX <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()


correlation_matrix <- v[rownames(adj_matrix) %in% Dataset_COI$Taxa,
                                         colnames(adj_matrix) %in% TaxaX]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]


pdf(file = "Figure_output/Corr_Plots/Metazooplankton_Protozooplankton.pdf", width = 20, height = 20)
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Metazooplankton and Protozooplankton lagged interactions"
)

dev.off()
```
## Heterotrophic to phytoplankton
```{r}
TaxaY <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- adj_matrix[rownames(adj_matrix) %in% TaxaY,
                                         colnames(adj_matrix) %in% TaxaX]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]

pdf(file = "Figure_output/Corr_Plots/Protozooplankton_Phytoplankton.pdf", width = 20, height = 20)
# Create the heatmap with a custom color scale
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Protozooplankton and Phytoplankton Lagged Interactions"
)

dev.off()
```

## Carnivorous to non-carnivorous Metazooplankton
```{r}
TaxaY <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% carnivorous_zooplankton,
                                         colnames(correlation_matrix_all) %in% omnivorous_filter_feeding_zooplankton]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]

pdf(file = "Figure_output/Corr_Plots/Carnivorous_Omnivorous_Mesozooplankton.pdf", width = 10, height = 10)
# Create the heatmap with a custom color scale
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Carnivorous to Non-Canivorous Metazooplankton Lagged Interactions"
)

dev.off()
```

## Phytoplankton
```{r}
correlation_matrix_all <- cor(seasonal_data_all %>% select(-Sample_date), use = "pairwise.complete.obs")


TaxaY <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% TaxaX,
                                         colnames(correlation_matrix_all) %in% TaxaX]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]


pdf(file = "Figure_output/Corr_Plots/Phytoplankton.pdf", width = 20, height = 20)
# Create the heatmap with a custom color scale
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Phytoplankton Clusters - No Lagg"
)

dev.off()
```

## Zooplankton
```{r}
TaxaY <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% Dataset_COI$Taxa,
                                         colnames(correlation_matrix_all) %in% Dataset_COI$Taxa]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]

pdf(file = "Figure_output/Corr_Plots/Mesozooplankton.pdf", width = 20, height = 20)
# Create the heatmap with a custom color scale
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Mesozooplankton Clusters - No Lagg"
)

dev.off()
```


## Fish
```{r}

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% Dataset_12S$Taxa,
                                         colnames(correlation_matrix_all) %in% Dataset_12S$Taxa]
correlation_matrix
# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]

pdf(file = "Figure_output/Corr_Plots/Mesozooplankton.pdf", width = 20, height = 20)
# Create the heatmap with a custom color scale
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Mesozooplankton Clusters - No Lagg"
)

dev.off()
```


## Microplankton
```{r}
TaxaY <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% TaxaY,
                                         colnames(correlation_matrix_all) %in% TaxaY]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]


pdf(file = "Figure_output/Corr_Plots/Protozooplankton.pdf", width = 20, height = 20)
# Create the heatmap with a custom color scale
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Protozooplankton Clusters - No Lagg"
)

dev.off()
```


# Seasonal Clusters


## Phytoplankton

```{r}
set.seed(200)

correlation_matrix_all <- cor(seasonal_data_all %>% select(-Sample_date), use = "pairwise.complete.obs")

TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% TaxaX,
                                         colnames(correlation_matrix_all) %in% TaxaX]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]





# Sample code to perform K-means clustering on a correlation matrix

# Assuming `correlation_matrix` is your correlation matrix:
# Step 1: Convert the correlation matrix to a dissimilarity matrix
# Using 1 - correlation to represent distance
dissimilarity_matrix <- 1 - correlation_matrix

# Step 2: Choose the number of clusters (k)
# Let's assume you want 3 clusters (you can experiment with other values of k)
k <- 7

# Step 3: Perform K-means clustering
# Run K-means on the transformed dissimilarity matrix
# Use the `as.dist` function to treat it as a distance matrix
kmeans_result <- kmeans(as.dist(dissimilarity_matrix), centers = k, nstart = 25)

# Step 4: Check the results
# Cluster assignment for each variable
cluster_df <- tibble(
  Taxa = names(kmeans_result$cluster),
  Cluster = as.factor(kmeans_result$cluster)
)



library(cluster)

# Compute silhouette values for different k
silhouette_scores <- vector("numeric", length = 10)
for (i in 2:10) {  # k = 1 is not meaningful for silhouette analysis
  kmeans_result <- kmeans(correlation_matrix, centers = i)
  silhouette_scores[i] <- mean(silhouette(kmeans_result$cluster, dist(correlation_matrix))[, 3])
}

# Plot silhouette scores
plot(2:10, silhouette_scores[2:10], type = "b", main = "Silhouette Method",
     xlab = "Number of clusters", ylab = "Average Silhouette Score")





pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Phytoplankton Clusters - No Lagg"
)


plot_cluster <- function(cluster) {
  
  seasonal_data_all %>% 
  pivot_longer(!Sample_date, names_to = "Taxa", values_to = "Abundance") %>%
  left_join(cluster_df, by = "Taxa") %>% 
  filter(Cluster == cluster) %>% 
  ggplot() +
  geom_line(aes(Sample_date, Abundance)) +
  facet_wrap("Taxa", scales = "free")
  
  
}
```


```{r}
plot_cluster(1)
plot_cluster(2)
plot_cluster(3)
plot_cluster(4)
plot_cluster(5)
plot_cluster(6)
plot_cluster(7)
```


## Protozooplankton

```{r}
set.seed(200)

correlation_matrix_all <- cor(seasonal_data_all %>% select(-Sample_date), use = "pairwise.complete.obs")

TaxaX <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% TaxaX,
                                         colnames(correlation_matrix_all) %in% TaxaX]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]


# Sample code to perform K-means clustering on a correlation matrix

# Assuming `correlation_matrix` is your correlation matrix:
# Step 1: Convert the correlation matrix to a dissimilarity matrix
# Using 1 - correlation to represent distance
dissimilarity_matrix <- 1 - correlation_matrix

# Step 2: Choose the number of clusters (k)
# Let's assume you want 3 clusters (you can experiment with other values of k)
k <- 6

# Step 3: Perform K-means clustering
# Run K-means on the transformed dissimilarity matrix
# Use the `as.dist` function to treat it as a distance matrix
kmeans_result <- kmeans(as.dist(dissimilarity_matrix), centers = k, nstart = 25)

# Step 4: Check the results
# Cluster assignment for each variable
cluster_df <- tibble(
  Taxa = names(kmeans_result$cluster),
  Cluster = as.factor(kmeans_result$cluster)
)



library(cluster)

# Compute silhouette values for different k
silhouette_scores <- vector("numeric", length = 10)
for (i in 2:10) {  # k = 1 is not meaningful for silhouette analysis
  kmeans_result <- kmeans(correlation_matrix, centers = i)
  silhouette_scores[i] <- mean(silhouette(kmeans_result$cluster, dist(correlation_matrix))[, 3])
}

# Plot silhouette scores
plot(2:10, silhouette_scores[2:10], type = "b", main = "Silhouette Method",
     xlab = "Number of clusters", ylab = "Average Silhouette Score")





pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Phytoplankton Clusters - No Lagg"
)


plot_cluster <- function(cluster) {
  
  seasonal_data_all %>% 
  pivot_longer(!Sample_date, names_to = "Taxa", values_to = "Abundance") %>%
  left_join(cluster_df, by = "Taxa") %>% 
  filter(Cluster == cluster) %>% 
  ggplot() +
  geom_line(aes(Sample_date, Abundance)) +
  facet_wrap("Taxa", scales = "free")
  
  
}
```


```{r}
plot_cluster(1)
plot_cluster(2)
plot_cluster(3)
plot_cluster(4)
plot_cluster(5)
plot_cluster(6)
```

## Zooplankton

```{r}
set.seed(200)

correlation_matrix_all <- cor(seasonal_data_all %>% select(-Sample_date), use = "pairwise.complete.obs")

TaxaX <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% Dataset_COI$Taxa,
                                         colnames(correlation_matrix_all) %in% Dataset_COI$Taxa]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]


# Sample code to perform K-means clustering on a correlation matrix

# Assuming `correlation_matrix` is your correlation matrix:
# Step 1: Convert the correlation matrix to a dissimilarity matrix
# Using 1 - correlation to represent distance
dissimilarity_matrix <- 1 - correlation_matrix

# Step 2: Choose the number of clusters (k)
# Let's assume you want 3 clusters (you can experiment with other values of k)
k <- 6

# Step 3: Perform K-means clustering
# Run K-means on the transformed dissimilarity matrix
# Use the `as.dist` function to treat it as a distance matrix
kmeans_result <- kmeans(as.dist(dissimilarity_matrix), centers = k, nstart = 25)

# Step 4: Check the results
# Cluster assignment for each variable
cluster_df <- tibble(
  Taxa = names(kmeans_result$cluster),
  Cluster = as.factor(kmeans_result$cluster)
)



library(cluster)

# Compute silhouette values for different k
silhouette_scores <- vector("numeric", length = 10)
for (i in 2:10) {  # k = 1 is not meaningful for silhouette analysis
  kmeans_result <- kmeans(correlation_matrix, centers = i)
  silhouette_scores[i] <- mean(silhouette(kmeans_result$cluster, dist(correlation_matrix))[, 3])
}

# Plot silhouette scores
plot(2:10, silhouette_scores[2:10], type = "b", main = "Silhouette Method",
     xlab = "Number of clusters", ylab = "Average Silhouette Score")





pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Phytoplankton Clusters - No Lagg"
)


plot_cluster <- function(cluster) {
  
  seasonal_data_all %>% 
  pivot_longer(!Sample_date, names_to = "Taxa", values_to = "Abundance") %>%
  left_join(cluster_df, by = "Taxa") %>% 
  filter(Cluster == cluster) %>% 
  ggplot() +
  geom_line(aes(Sample_date, Abundance)) +
  facet_wrap("Taxa", scales = "free")
  
  
}
```


```{r}
plot_cluster(1)
plot_cluster(2)
plot_cluster(3)
plot_cluster(4)
plot_cluster(5)
plot_cluster(6)
```


## Fish

```{r}
set.seed(200)

correlation_matrix_all <- cor(seasonal_data_all %>% select(-Sample_date), use = "pairwise.complete.obs")



correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% Dataset_12S$Taxa,
                                         colnames(correlation_matrix_all) %in% Dataset_12S$Taxa]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]


# Sample code to perform K-means clustering on a correlation matrix

# Assuming `correlation_matrix` is your correlation matrix:
# Step 1: Convert the correlation matrix to a dissimilarity matrix
# Using 1 - correlation to represent distance
dissimilarity_matrix <- 1 - correlation_matrix

# Step 2: Choose the number of clusters (k)
# Let's assume you want 3 clusters (you can experiment with other values of k)
k <- 9

# Step 3: Perform K-means clustering
# Run K-means on the transformed dissimilarity matrix
# Use the `as.dist` function to treat it as a distance matrix
kmeans_result <- kmeans(as.dist(dissimilarity_matrix), centers = k, nstart = 25)

# Step 4: Check the results
# Cluster assignment for each variable
cluster_df <- tibble(
  Taxa = names(kmeans_result$cluster),
  Cluster = as.factor(kmeans_result$cluster)
)



library(cluster)

# Compute silhouette values for different k
silhouette_scores <- vector("numeric", length = 10)
for (i in 2:10) {  # k = 1 is not meaningful for silhouette analysis
  kmeans_result <- kmeans(correlation_matrix, centers = i)
  silhouette_scores[i] <- mean(silhouette(kmeans_result$cluster, dist(correlation_matrix))[, 3])
}

# Plot silhouette scores
plot(2:10, silhouette_scores[2:10], type = "b", main = "Silhouette Method",
     xlab = "Number of clusters", ylab = "Average Silhouette Score")





pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Phytoplankton Clusters - No Lagg"
)


plot_cluster <- function(cluster) {
  
  seasonal_data_all %>% 
  pivot_longer(!Sample_date, names_to = "Taxa", values_to = "Abundance") %>%
  left_join(cluster_df, by = "Taxa") %>% 
  filter(Cluster == cluster) %>% 
  ggplot() +
  geom_line(aes(Sample_date, Abundance)) +
  facet_wrap("Taxa", scales = "free")
  
  
}
```


```{r}
plot_cluster(1)
plot_cluster(2)
plot_cluster(3)
plot_cluster(4)
plot_cluster(5)
plot_cluster(6)
plot_cluster(7)
plot_cluster(8)
plot_cluster(9)
```


```{r}
seasonal_data_all %>% 
  pivot_longer(!Sample_date, names_to = "Taxa", values_to = "Abundance") %>%
  left_join(cluster_df, by = "Taxa") %>% 
  filter(Cluster == 1) %>%
  pull(Taxa) %>% unique()
```




# Plot Seasonality

```{r}

  
  
seasonal_data_all %>% 
  pivot_longer(!Sample_date, names_to = "Taxa", values_to = "Abundance") %>% 
  filter(Taxa %in% omnivorous_filter_feeding_zooplankton) %>% 
  ggplot() +
  geom_line(aes(Sample_date, Abundance)) +
  facet_wrap("Taxa", scales = "free")
```




# ___OLD CODE___

## Plot Heatmap in ggplot
```{r}
taxatab_18S <- Dataset_18S %>%
  group_by(Phylum, Species1 = Class) %>% 
  summarise()

corr_data <- ccf_results %>%
  dplyr::filter(Species2 %in% Interpolated_COI_Data$Taxa,
         Species1 %in% Interpolated_COI_Data$Taxa) %>% 
  filter(Max_Correlation > 0.4,
         Lag_At_Max_Correlation >= 0) %>% 
  left_join(taxatab_18S, by = c("Species1"))


corr_data %>% 
  ggplot() +
  geom_raster(aes(Species2, Species1, fill = Max_Correlation)) +
  geom_text(aes(Species2, Species1, label = Lag_At_Max_Correlation), size = 2) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
                       midpoint = 0.4, limit = c(0.4, 1), name = "Correlation") +
  theme_minimal(8) +
  labs(x = "Zooplankton", y = "Microeukaryotes", title = "Correlation between Taxa with 0-4 weeks lag") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)

ggsave("Figure_output/Correlation_plot.pdf")

```




## Correlation by sample

```{r}

all_community_mat <- bind_rows(
  filter(Dataset_18S, Hakai_ID %in% Dataset_COI$Hakai_ID),
  filter(Dataset_COI, Hakai_ID %in% Dataset_COI$Hakai_ID)) %>% 
  group_by(Hakai_ID, Taxa) %>%
  summarize(RA = mean(RA)) %>% 
  pivot_wider(names_from = Taxa, values_from = RA) %>% 
  column_to_rownames("Hakai_ID") %>%
  mutate(across(everything(), ~ replace_na(.x, 0))) %>% 
  as.matrix()

all_community_mat

all_cor_matrix <- cor(all_community_mat, method = "pearson", use = "pairwise.complete.obs")

cor_df <- as.data.frame(all_cor_matrix) %>%
  rownames_to_column(var = "Taxa1") %>%
  pivot_longer(-Taxa1, names_to = "Taxa2", values_to = "Correlation") %>%
  filter(Taxa1 %!in% Dataset_COI$Taxa,
         Taxa2 %!in% Dataset_18S$Taxa)

cor_df %>%
  filter(Correlation > 0.4) %>% 
  ggplot() +
  geom_raster(aes(Taxa2, Taxa1, fill = Correlation)) +
  #geom_text(aes(Taxa2, Taxa1, label = Lag_At_Max_Correlation), size = 2) +
  #scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
  #                     midpoint = 0.4, limit = c(0.4, 1), name = "Correlation") +
  theme_minimal(8) +
  labs(x = "Zooplankton", y = "Microeukaryotes", title = "Correlation between Taxa with 0-4 weeks lag") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)
  
cor_df
```



