---
title: "R Notebook"
---


```{r}
# Data import

library(readxl)
#library(phyloseq)


# All data format modifications and filtering and plotting
library(tidyverse)
'%!in%' <- function(x,y)!('%in%'(x,y))
source("Code/Functions.R")
library(zoo)

# Statistical tools
library(vegan)

# Specific plotting tools
library(ggpubr)
library(paletteer)
library(viridis)
library(cowplot)
#library(ggnewscale)
source("Code/scales_and_themes.R")

# Make sure tidyverse is the latest loaded package
detach("package:tidyverse", unload=TRUE)
library(tidyverse)


# Custom color scale for depths
depth_colors <- c("0" = "#add8e6",    # Light blue for 0 m
                  "5" = "#87ceeb",    # Sky blue for 5 m
                  "30" = "#4682b4",   # Steel blue for 30 m
                  "100" = "#00008b")  # Dark blue for 100 m



```


```{r}
Dataset_COI <- readRDS("Data_output/Data_merge/COI_glomm_indexed.rds")
Dataset_18S <- readRDS("Data_output/Data_merge/pr2_18S_glomm_indexed.rds")
Dataset_12S <- readRDS("Data_output/Data_merge/12S_glomm_indexed.rds")
Dataset_Bottle <- readRDS("Data_output/Data_merge/bottle_file.rds")
  
  
taxatab_18S <- Dataset_18S %>%
  group_by(Phylum, Class, Order, Family, Genus, Taxa) %>% 
  summarise()

taxatab_COI <- Dataset_COI %>%
  group_by(Phylum, Class, Order, Family, Genus, Taxa) %>% 
  summarise()

taxtab_12S <- Dataset_12S %>% 
  group_by(Class, Taxa) %>% 
  summarise() %>% 
  mutate(Phylum = "Chordata")


Taxtab_X <- taxatab_18S %>% 
  group_by(PhylumX = Phylum, ClassX = Class, OrderX = Order,
           FamilyX = Family, GenusX = Genus, TaxaX = Taxa) %>% 
  summarise()
Taxtab_Y <- taxatab_18S %>% 
  group_by(PhylumY = Phylum, ClassY = Class, OrderY = Order,
           FamilyY = Family, GenusY = Genus, TaxaY = Taxa) %>% 
  summarise()

```



# 1. Interpolation

	1.	Fill Missing Combinations with Zeros: The complete(Sample_ID, Sample_date, Taxa, Sampling_Depth, fill = list(Relative_Abundance = 0)) line ensures that any combination without observed data is filled with 0 for Relative_Abundance.
	2.	Generate Weekly Dates: complete(Sample_date = seq.Date(min(Sample_date), max(Sample_date), by = "week")) generates weekly dates for each Taxa and Sampling_Depth combination.
	3.	Interpolate Missing Values: na.approx() fills in any NAs generated by the new weekly dates in Sample_date.


### COI
```{r}

start_date <- as.Date("2017-01-01")
end_date <- as.Date("2019-02-15")


# Complete the data frame with all combinations of Sample_ID, Sample_date, Taxa, and Sampling_Depth
complete_data <- Dataset_COI %>%
  filter(Line_out_depth %in% c(0,5,30,100)) %>%
  mutate(Line_out_depth = factor(Line_out_depth, levels = c("0", "5", "30", "100"))) %>% 
  mutate(Sample_date = as.Date(Sample_date)) %>% 
  filter(Sample_date >= start_date & Sample_date <= end_date) %>% 
  complete(Sample_date, Taxa, Line_out_depth, fill = list(Abundance = 0)) %>% 
  # Ensure dates are properly ordered after completion
  arrange(Sample_date, Taxa, Line_out_depth)


strict_weekly_dates <- seq.Date(from = min(complete_data$Sample_date),
                                to = max(complete_data$Sample_date),
                                by = "week")

smoothed_data <- complete_data %>%
  # Expand to include all combinations of Taxa, Depth, and weekly dates
  group_by(Taxa, Line_out_depth) %>%
  complete(Sample_date = strict_weekly_dates) %>%
  arrange(Sample_date) %>%
  # Interpolate missing values in RA_Index within each group
  mutate(Abundance = ifelse(is.na(Abundance),
                           na.approx(Abundance, rule = 2, na.rm = FALSE),
                           Abundance)) %>%
  # Apply 2-week rolling mean for smoothing (7-day intervals)
  mutate(Smoothed_Relative_Abundance = rollmean(Abundance, k = 2, fill = NA, align = "center")) %>%
  ungroup() %>%
  # Filter to include only rows with Smoothed_Relative_Abundance calculated
  filter(!is.na(Smoothed_Relative_Abundance)) %>% 
  filter(Sample_date %in% strict_weekly_dates)

Interpolated_COI_Data <- smoothed_data %>% 
  group_by(Taxa, Sample_date) %>% 
  summarise(Abundance = sum(Smoothed_Relative_Abundance)/4)
```

### 12S

```{r}

start_date <- as.Date("2017-01-01")
end_date <- as.Date("2019-02-15")


# Complete the data frame with all combinations of Sample_ID, Sample_date, Taxa, and Sampling_Depth
complete_data <- Dataset_12S %>%
  filter(Line_out_depth %in% c(0,5,30,100)) %>%
  mutate(Line_out_depth = factor(Line_out_depth, levels = c("0", "5", "30", "100"))) %>% 
  mutate(Sample_date = as.Date(Sample_date)) %>% 
  #filter(Sample_date >= start_date & Sample_date <= end_date) %>% 
  complete(Sample_date, Taxa, Line_out_depth, fill = list(Abundance = 0)) %>% 
  # Ensure dates are properly ordered after completion
  arrange(Sample_date, Taxa, Line_out_depth)


#strict_weekly_dates <- seq.Date(from = min(complete_data$Sample_date),
#                                to = max(complete_data$Sample_date),
#                                by = "week")

smoothed_data <- complete_data %>%
  # Expand to include all combinations of Taxa, Depth, and weekly dates
  group_by(Taxa, Line_out_depth) %>%
  complete(Sample_date = strict_weekly_dates) %>%
  arrange(Sample_date) %>%
  # Interpolate missing values in RA_Index within each group
  mutate(Abundance = ifelse(is.na(Abundance),
                           na.approx(Abundance, rule = 2, na.rm = FALSE),
                           Abundance)) %>%
  # Apply 2-week rolling mean for smoothing (7-day intervals)
  mutate(Smoothed_Relative_Abundance = rollmean(Abundance, k = 2, fill = NA, align = "center")) %>%
  ungroup() %>%
  # Filter to include only rows with Smoothed_Relative_Abundance calculated
  filter(!is.na(Smoothed_Relative_Abundance)) %>% 
  filter(Sample_date %in% strict_weekly_dates)

Interpolated_12S_Data <- smoothed_data %>% 
  group_by(Taxa, Sample_date) %>% 
  summarise(Abundance = sum(Smoothed_Relative_Abundance)/4)




```


### 18S

```{r}


# Complete the data frame with all combinations of Sample_ID, Sample_date, Taxa, and Sampling_Depth
complete_data <- Dataset_18S %>%
  filter(Line_out_depth %in% c(0,5,30,100)) %>%
  mutate(Line_out_depth = factor(Line_out_depth, levels = c("0", "5", "30", "100"))) %>%
  #filter(Phylum != "Metazoa") %>%
  #  filter(Phylum %in% c(
  #  "Chlorophyta",
  #  "Cryptophyta",
  #  "Dinoflagellata",
  #  "Haptophyta",
  #  "Ochrophyta"
  #)) %>% 
  #filter(is.na(Genus)==FALSE) %>% 
  mutate(Sample_date = as.Date(Sample_date)) %>% 
  complete(Sample_date, Taxa, Line_out_depth, fill = list(Abundance = 0)) %>%
  filter(Sample_date >= start_date & Sample_date <= end_date) %>%
  #filter(Sample_date <= end_date) %>%
  # Ensure dates are properly ordered after completion
  arrange(Sample_date, Taxa, Line_out_depth)


# NOTE: Commented out to reuse the series for COI
strict_weekly_dates <- seq.Date(from = min(complete_data$Sample_date),
                                to = max(complete_data$Sample_date),
                                by = "week")

smoothed_data <- complete_data %>%
  # Expand to include all combinations of Taxa, Depth, and weekly dates
  group_by(Taxa, Line_out_depth) %>%
  complete(Sample_date = strict_weekly_dates) %>%
  arrange(Sample_date) %>%
  # Interpolate missing values in RA_Index within each group
  mutate(Abundance = ifelse(is.na(Abundance),
                           na.approx(Abundance, rule = 2, na.rm = FALSE),
                           Abundance)) %>%
  # Apply 2-week rolling mean for smoothing (7-day intervals)
  mutate(Smoothed_Relative_Abundance = rollmean(Abundance, k = 4, fill = NA, align = "center")) %>%
  ungroup() %>%
  # Filter to include only rows with Smoothed_Relative_Abundance calculated
  filter(!is.na(Smoothed_Relative_Abundance)) %>% 
  filter(Sample_date %in% strict_weekly_dates)

Interpolated_18S_Data <- smoothed_data %>% 
  group_by(Taxa, Sample_date) %>% 
  summarise(Abundance = sum(Smoothed_Relative_Abundance)/4)
```

```{r}
Interpolated_COI_Data$Sample_date %>% min()
Interpolated_COI_Data$Sample_date %>% max()

Interpolated_12S_Data$Sample_date %>% min()
Interpolated_12S_Data$Sample_date %>% max()

Interpolated_18S_Data$Sample_date %>% min()
Interpolated_18S_Data$Sample_date %>% max()
```


# 2. STL - Extract seasonal data
```{r}

### COI
seasonal_data_COI <- Interpolated_COI_Data %>%
  
  # Extract seasonal data from STL time series, per taxa
  group_by(Taxa) %>%
  do({
    ts_data <- ts(.$Abundance, start = c(2017, 08), frequency = 52)  # Weekly frequency
    stl_decomposition <- stl(ts_data, s.window = "periodic")
    data.frame(Sample_date = .$Sample_date, Seasonal = stl_decomposition$time.series[, "seasonal"])
  }) %>%
  ungroup() %>%
  spread(key = Taxa, value = Seasonal) %>%
  drop_na()


#### 18S
seasonal_data_18S <- Interpolated_18S_Data %>%
  # Extract seasonal data from STL time series, per taxa
  group_by(Taxa) %>%
  do({
    ts_data <- ts(.$Abundance, start = c(2017, 15), frequency = 52)  # Weekly frequency
    stl_decomposition <- stl(ts_data, s.window = "periodic")
    data.frame(Sample_date = .$Sample_date, Seasonal = stl_decomposition$time.series[, "seasonal"])
  }) %>%
  ungroup() %>%
  spread(key = Taxa, value = Seasonal) %>%
  drop_na()


#### 12S
seasonal_data_12S <- Interpolated_12S_Data %>%
  filter(Sample_date != "2017-01-08") %>% 
  # Extract seasonal data from STL time series, per taxa
  group_by(Taxa) %>%
  do({
    ts_data <- ts(.$Abundance, start = c(2017, 08), frequency = 52)  # Weekly frequency
    stl_decomposition <- stl(ts_data, s.window = "periodic")
    data.frame(Sample_date = .$Sample_date, Seasonal = stl_decomposition$time.series[, "seasonal"])
  }) %>%
  ungroup() %>%
  spread(key = Taxa, value = Seasonal) %>%
  drop_na()


start_date <- as.Date("2017-01-15")
end_date <- as.Date("2019-01-13")

### Combine
seasonal_data_all <-left_join(seasonal_data_18S, seasonal_data_COI, by = "Sample_date") %>%
  left_join(seasonal_data_12S, by = "Sample_date") %>% 
  filter(Sample_date >= start_date & Sample_date <= end_date)
seasonal_data_all

```

## Test matrix of just TS data

```{r}
weekly_data_all <- pivot_wider(Interpolated_12S_Data, names_from = Taxa, values_from = Abundance) %>% 
  left_join(pivot_wider(Interpolated_18S_Data, names_from = Taxa, values_from = Abundance)) %>% 
  left_join(pivot_wider(Interpolated_COI_Data, names_from = Taxa, values_from = Abundance)) %>% 
  filter(Sample_date != "2017-01-08", 
         Sample_date != "2019-01-20",
         Sample_date != "2019-01-27",
         Sample_date != "2019-02-03")
```


# 3. Define vectors for filtering

```{r}
# Define vectors for filtering
phototrophic_plankton <- c(
  "Chlorarachniophyceae", "Chlorodendrophyceae", "Chlorophyceae", "Mamiellophyceae", 
  "Nephroselmidophyceae", "Pedinophyceae", "Prasino-Clade-VIII", "Pyramimonadophyceae", 
  "Trebouxiophyceae", "Ulvophyceae", "Cryptophyceae", "Dinophyceae", "Noctilucophyceae", 
  "Syndiniales", "Prymnesiophyceae", "Bacillariophyta", "Bolidophyceae", "Chrysophyceae", 
  "Dictyochophyceae", "Pelagophyceae", "Phaeophyceae", "Raphidophyceae", "Perkinsida", 
  "Prasinodermophyceae", "Florideophyceae", "Embryophyceae"
)

heterotrophic_protists <- c(
  "Ellobiopsidae", "Gregarinomorphea", "Apusomonadidae_Group-1", "Filosa", 
  "Filosa-Granofilosea", "Filosa-Imbricatea", "Filosa-Thecofilosea", "Phaeodarea", 
  "Choanoflagellatea", "CONTH_7", "CONTH_8", "CONThreeP", "Cyclotrichium_like_organism", 
  "Litostomatea", "Nassophorea", "Oligohymenophorea", "Phyllopharyngea", 
  "Prostomatea_1", "Spirotrichea", "Ascomycota", "Basidiomycota", "Chytridiomycota", 
  "Mucoromycota", "Haptophyta_Clade_HAP2", "Haptophyta_Clade_HAP3", "Haptophyta_Clade_HAP4", 
  "Haptophyta_Clade_HAP5", "Planomonadida", "Katablepharidaceae", "Tubulinea", "Filasterea", 
  "Ichthyosporea", "MOCH-1", "MOCH-3", "MOCH-4", "RAD-A", "RAD-B", "RAD-C", "Rhodelphea"
)

carnivorous_zooplankton <- c(
  "Aetideus divergens", "Cyanea capillata", "Euphausia pacifica", 
  "Euphysa flammea", "Merluccius angustimanus", "Nanomia bijuga", 
  "Pandalus stenolepis", "Thysanoessa spinifera", "Clytia hemisphaerica",
  "Obelia dichotoma", "Epilabidocera amphitrites", "Pleurobrachia bachei"
  
)

# Vector for Omnivorous/Filter Feeding Zooplankton
omnivorous_filter_feeding_zooplankton <- c(
  "Acartia longiremis", "Calanus pacificus", "Centropages abdominalis", 
  "Corycaeus anglicus", "Discoconchoecia elegans",  
  "Eucalanus bungii", "Evadne nordmanni", "Metridia pacifica", 
  "Microcalanus pygmaeus", "Oithona atlantica", 
  "Paracalanus indicus", "Paracalanus parvus",  
  "Pseudocalanus acuspes", "Pseudocalanus mimus", "Pseudocalanus minutus", 
  "Pseudocalanus newmani", "Triconia borealis"
)




microbial_eukaryotes <- c(
  "Endomyxa", "Endomyxa-Phytomyxea", "Katablepharidaceae", "Hyphochytriomyceta", "Oomycota", 
  "Pirsonia_Clade", "Polycystinea", "Acantharea", "Labyrinthulomycetes", "MAST-10", 
  "MAST-12", "MAST-3", "MAST-4", "MAST-6", "MAST-7", "MAST-8", "MAST-9", "MAST-25", 
  "Stramenopiles_X-Group-2", "Stramenopiles_X-Group-6"
)

marine_viruses_symbionts <- c(
  "Picozoa_X", "Prasinodermophyceae", "Telonemia_X", "Syndiniales"
)

miscellaneous_unclassified <- c(
  "CONThreeP", "MOCH-4", "Rhodelphea", "Filasterea"
)

Dataset_COI$Taxa %>% unique()


pelagic_fish <- c(
  "Alosa sapidissima", "Ammodytes personatus", "Anoplopoma fimbria", 
  "Boreogadus saida", "Clupea pallasii", "Diaphus theta", 
  "Engraulis mordax", "Gasterosteus aculeatus", "Mallotus villosus", 
  "Merluccius productus", "Oncorhynchus gorbuscha", "Oncorhynchus keta", 
  "Oncorhynchus kisutch", "Oncorhynchus nerka", "Oncorhynchus tshawytscha", 
  "Osmerus mordax", "Sardinops sagax", "Stenobrachius leucopsarus", 
  "Stenobrachius nannochir", "Thaleichthys pacificus", "Zaprora silenus"
)
```


Dataframe for taxa based on above

```{r}
PZP <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>%
  ungroup() %>% 
  select(Taxa, Class, Phylum) %>% 
  mutate(FG = "Protozooplankton")

PP <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>%
  ungroup() %>% 
  select(Taxa, Class, Phylum) %>% 
  mutate(FG = "Phytoplankton")

CZP <- taxatab_COI %>% 
  filter(Taxa %in% carnivorous_zooplankton) %>% 
  ungroup() %>% 
  select(Taxa, Class, Phylum) %>% 
  mutate(FG = "Carnivorous_Zooplankton")

OZP <- taxatab_COI %>% 
  filter(Taxa %in% omnivorous_filter_feeding_zooplankton) %>% 
  ungroup() %>% 
  select(Taxa, Class, Phylum) %>% 
  mutate(FG = "Omnivorous_Zooplankton")


PF <- taxtab_12S %>% 
  ungroup() %>% 
  filter(Taxa %in% pelagic_fish) %>%
  mutate(FG = "Nekton")
  
Functional_groups <- bind_rows(PZP, PP, CZP, OZP, PF) %>% 
  select(FG, Phylum, Class, Taxa)

Functional_groups

```




# 4. Construct correlation matrixes:

## a. Without lags
```{r}

correlation_matrix_all <- cor(seasonal_data_all %>% select(-Sample_date), use = "pairwise.complete.obs")

```

Correlation without lag

```{r}
correlation_matrix_all <- cor(seasonal_data_all %>% select(-Sample_date), use = "pairwise.complete.obs")
correlation_matrix_all %>% colnames()
```


## b. With Lags

Define function for cff of specific predator and prey
```{r}
pw_ccf <- function(mat, predator, prey, max_lag = 3, min_corr = -1) {
  # Initialize results data frame
  results <- data.frame(Predator = character(), Prey = character(),
                        Max_Correlation = numeric(), Lag = integer(), stringsAsFactors = FALSE)
  
  # Loop through each predator-prey pair
  for (pred in predator) {
    for (prey_taxa in prey) {
      ts_pred <- mat[, pred]
      ts_prey <- mat[, prey_taxa]
      
      # Compute cross-correlation without plotting
      ccf_result <- ccf(ts_prey, ts_pred, lag.max = max_lag, plot = FALSE)
      
      # Check for valid cross-correlation values
      if (!is.null(ccf_result$acf) && length(ccf_result$acf) > 0 && !all(is.na(ccf_result$acf))) {
        max_corr_idx <- which.max(ccf_result$acf)
        max_corr <- ccf_result$acf[max_corr_idx]
        lag_at_max_corr <- ccf_result$lag[max_corr_idx]
        
        # Filter for significant correlations and relevant lags
        if (!is.na(max_corr) && lag_at_max_corr <= 0) {
          results <- rbind(results, data.frame(Predator = pred, Prey = prey_taxa,
                                               Max_Correlation = max_corr, Lag = lag_at_max_corr))
        }
      }
    }
  }
  
  return(results)
}


# Plotting output function


pw_ccf_plot <- function(...) {

  output_cff <- pw_ccf(...)
  
  matrix_results <- output_cff %>%
    select(Predator, Prey, Max_Correlation) %>%
    pivot_wider(names_from = Predator, values_from = Max_Correlation) %>%
    column_to_rownames(var = "Prey") %>% 
    as.matrix() %>% 
    t()
  matrix_results[is.na(matrix_results)] <- 0
  
  
  heatmap(matrix_results,
           color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
           #breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
           show_rownames = TRUE,
           show_colnames = TRUE,
           cluster_rows = TRUE,
           cluster_cols = TRUE,
           legend = TRUE
  )
}




```


```{r}
species_matrix <- as.matrix(seasonal_data_all %>% select(-Sample_date))
species_matrix_week <- as.matrix(weekly_data_all %>% select(-Sample_date))
```


### ZP - PZP
```{r}
Prey <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

pdf(file = "Figure_output/Corr_Plots_CFF/ZP - PZP.pdf", width = 20, height = 20)
pw_ccf_plot(species_matrix, unique(Dataset_COI$Taxa), Prey, min_corr = 0.5)


zp_pzp <- pw_ccf(species_matrix, unique(Dataset_COI$Taxa), Prey, min_corr = 0.1)

```

### ZP - PP
```{r}
Prey <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

pw_ccf_plot(species_matrix, unique(Dataset_COI$Taxa), Prey, min_corr = 0.1)

zp_pp <- pw_ccf(species_matrix, unique(Dataset_COI$Taxa), Prey, min_corr = 0.1)

zp_pp
```

### PZP - PP
```{r}
Predator <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

Prey <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()


pw_ccf_plot(species_matrix, Predator, Prey, min_corr = 0.1)

pzp_pp <- pw_ccf(species_matrix, Predator, Prey, min_corr = 0.1)

```

### ZP - ZP
```{r}
pw_ccf_plot(species_matrix, carnivorous_zooplankton, omnivorous_filter_feeding_zooplankton, min_corr = 0.1)

zp_zp <- pw_ccf(species_matrix, carnivorous_zooplankton, omnivorous_filter_feeding_zooplankton, min_corr = 0.1)
```
### Fish - ZP
```{r}
pw_ccf_plot(species_matrix, pelagic_fish, unique(Dataset_COI$Taxa), min_corr = 0.1)

fish_zp <- pw_ccf(species_matrix, pelagic_fish, unique(Dataset_COI$Taxa), min_corr = 0.1)
```

### Merge CCF to adj.mat

```{r}

df <- bind_rows(fish_zp, zp_zp, zp_pzp, zp_pp, pzp_pp) %>% 
  filter(Max_Correlation > 0.45) %>% 
  select(Predator, Prey)

# Get all unique taxa
all_taxa <- unique(c(df$Predator, df$Prey))

# Initialize square adjacency matrix
adj_matrix <- matrix(0, nrow = length(all_taxa), ncol = length(all_taxa),
                     dimnames = list(all_taxa, all_taxa))

# Populate matrix with interactions
for (i in seq_len(nrow(df))) {
  adj_matrix[df$Predator[i], df$Prey[i]] <- 1
}


adj_matrix


taxatab_18S %>% 
  write_csv("Data_output/taxtab_18S.csv")

```


# 5. Network analysis

```{r}
library(igraph)
```


## a. Global Seasonal Network Metrics


```{r}
# Function for constructing igraph based on date

date_graph <- function(Date) {
  
  a <- week(date(Date))
  b <- a-1
  c <- a+1
  weeks <- c(a,b,c)
  
  day_taxa <- seasonal_data_all %>%
    pivot_longer(!1, names_to = "Taxa", values_to = "Relative") %>%
    filter(Relative > 0) %>%
    mutate(Week = week(Sample_date)) %>% 
    filter(Week %in% weeks) %>%
    filter(Taxa %in% colnames(adj_matrix)) %>% 
    pull(Taxa) %>% 
    unique()


filtered_matrix <- adj_matrix[day_taxa, day_taxa, drop = FALSE] %>% 
  t()

graph <- graph_from_adjacency_matrix(t(filtered_matrix), mode = "directed", weighted = NULL)
}

```

	Global Metrics

	•	Density: Proportion of possible edges that are realized.
	•	Diameter: Longest shortest path in the network.
	•	Average Path Length: Average shortest path length between all node pairs.
	
	Other Metrics

	•	Modularity: Strength of division into clusters.
	•	Clustering Coefficient: Tendency of nodes to form triangles.

As loop:
```{r}
All_dates <- seasonal_data_all %>%
  filter(year(Sample_date) == 2017) %>% 
  pull(Sample_date) %>% unique() %>% 
  as.Date()

# Initialize an empty dataframe to store results
results_df <- data.frame(
  Date = Date(),
  Density = numeric(),
  Diameter = numeric(),
  Average_Path_Length = numeric(),
  Modularity = numeric(),
  Clustering_Coefficient = numeric(),
  stringsAsFactors = FALSE
)

# Loop through all dates
for (current_date in All_dates) {
  # Generate the graph for the current date
  graph <- date_graph(as.Date(current_date))
  
  # Calculate global metrics
  density <- edge_density(graph)
  diameter <- diameter(graph)
  avg_path_length <- mean_distance(graph)
  
  # Modularity
  clusters <- cluster_edge_betweenness(graph)
  modularity_score <- modularity(clusters)
  
  # Clustering coefficient
  clustering_coeff <- transitivity(graph, type = "average")
  
  # Number of nodes
  num_nodes <- vcount(graph)

  # Number of links (edges)
  num_links <- ecount(graph)
  
  # Append the results to the dataframe
  results_df <- rbind(results_df, data.frame(
    Date = current_date,
    Density = density,
    #Diameter = diameter,
    Average_Path_Length = avg_path_length,
    Modularity = modularity_score,
    Clustering_Coefficient = clustering_coeff,
    Number_nodes = num_nodes,
    Number_links = num_links
  ))
}

results_df %>% 
  mutate(Date = as.Date(Date)) %>%
  pivot_longer(!1, names_to = "Metrics", values_to = "Value") %>% 
  ggplot() +
  geom_line(aes(Date, Value)) +
  #geom_smooth(aes(Date, Value)) +
  facet_wrap("Metrics", scales = "free", ncol = 3) +
  theme_minimal_vgrid(font_size = 10) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        #strip.text = element_text(face = "italic"),
        panel.background = element_rect(fill = '#F5F5F5', colour = 'darkgrey'),
        #legend.position="none",,
        aspect.ratio = 1/2)

ggsave("Figure_output/Global_Metrics.pdf")

```



## b. Node-Level Metrics

	•	Degree Centrality: Number of connections (incoming or outgoing) for each node.
	•	Betweenness Centrality: How often a node lies on the shortest path between other nodes.
	•	Closeness Centrality: How close a node is to all other nodes.
	•	Eigenvector Centrality: Influence of a node in the network.
	
	
When is the peak of each node?
```{r}
Node_Data <- seasonal_data_all %>%
  filter(year(Sample_date) == 2017) %>% 
  pivot_longer(!1, names_to = "Taxa", values_to = "signal") %>% 
  group_by(Taxa) %>%
  slice_max(signal, with_ties = FALSE) %>% 
  ungroup() %>% 
  arrange(Sample_date) %>% 
  filter(Taxa %in% colnames(adj_matrix)) %>%
  select(Taxa, Date_max = Sample_date) %>% 
  right_join(Functional_groups, by = "Taxa")



```
	

### Simple Node Level Metrics

```{r}
graph <- graph_from_adjacency_matrix(t(adj_matrix), mode = "directed", weighted = NULL)
graph
# Node-Level metrics

degree <- degree(graph, mode = "all")          # Total degree (in + out)
in_degree <- degree(graph, mode = "in")        # Incoming connections
out_degree <- degree(graph, mode = "out")      # Outgoing connections
betweenness <- betweenness(graph)
closeness <- closeness(graph)
eigenvector <- eigen_centrality(graph)$vector

# Combine metrics into a data frame
node_metrics <- data.frame(
  #Taxa = V(graph)$name,
  Degree_Centrality = degree,
  In_Degree = in_degree,
  Out_Degree = out_degree,
  Betweenness_Centrality = betweenness,
  Closeness_Centrality = closeness,
  Eigenvector = eigenvector
) %>% 
  rownames_to_column("Taxa") %>% 
  left_join(Node_Data)

vcount(graph)
```




### Weighted Trophic Level
```{r}
# Calculate trophic levels iteratively
trophic_levels <- rep(NA, vcount(graph))
names(trophic_levels) <- V(graph)$name

# Assign initial trophic levels to basal nodes
trophic_levels[basal_nodes$name] <- 1

# Iteratively calculate for other nodes
for (i in 1:vcount(graph)) {
  for (node in V(graph)) {
    # Use the vertex index instead of the object directly
    prey <- neighbors(graph, node, mode = "in") # Prey are incoming neighbors
    if (length(prey) > 0) {
      trophic_levels[node] <- 1 + mean(trophic_levels[prey], na.rm = TRUE)
    }
  }
}

# Add as vertex attribute
V(graph)$trophic_level <- trophic_levels
TL <- data.frame(Trophic_level = trophic_levels) %>% 
  rownames_to_column("Taxa")
TL

node_metrics <- TL %>% 
  left_join(node_metrics, by = "Taxa")
node_metrics
```

## Paths per top predator

```{r}
# Initialize a data frame to store results
path_counts <- data.frame(
  Taxa = character(),
  Num_Paths = integer(),
  stringsAsFactors = FALSE
)

# Loop over each target node in the graph
for (target_node in V(graph)$name) {
  # Initialize an empty list for paths to this target node
  all_paths <- list()
  
  # Loop over each basal node and compute paths
  for (node in basal_nodes) {
    paths_from_node <- all_simple_paths(graph, from = node, to = V(graph)[name == target_node])
    all_paths <- c(all_paths, paths_from_node) # Append paths to the list
  }
  
  # Add the results for this target node to the data frame
  path_counts <- rbind(
    path_counts,
    data.frame(Taxa = target_node, Num_Paths = length(all_paths), stringsAsFactors = FALSE)
  )
}
node_metrics <- path_counts %>% 
  left_join(node_metrics, by = "Taxa")

node_metrics
```



### Plot

```{r}
# Functional groups
functional_groups <- c(
  "Phytoplankton",
  "Protozooplankton",
  "Omnivorous_Zooplankton",
  "Carnivorous_Zooplankton",
  "Nekton"
)

Metrics_levels <- c(
  "In_Degree",
  "Out_Degree",
  "Degree_Centrality",
  "Closeness_Centrality",
  "Betweenness_Centrality",
  "Eigenvector",
  "Trophic_level",
  "Num_Paths"
)

# Corresponding colors
colors <- c(
  "Phytoplankton" = "#66C2A5",
  "Protozooplankton" = "#FFD92F",
  "Omnivorous_Zooplankton" = "#8DA0CB",
  "Carnivorous_Zooplankton" = "#3288BD",
  "Nekton" = "#FC8D62"
)


node_metrics %>%
  pivot_longer(4:9, names_to = "Metrics", values_to = "Value") %>%
  mutate(Functional_Group = factor(FG, levels = functional_groups)) %>%
  mutate(Metrics = factor(Metrics, levels = Metrics_levels)) %>% 
  filter(is.na(FG) == FALSE) %>% 
  ggplot() +
  geom_point(aes(Date_max, Value, colour = Functional_Group), size = 0.5) +
  geom_smooth(aes(Date_max, Value, colour = Functional_Group), se = FALSE, size = 0.8 ) +
  facet_wrap("Metrics", scales = "free") +
  scale_color_manual(values = colors) +
      theme_minimal_vgrid(font_size = 10) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        #strip.text = element_text(face = "italic"),
        panel.background = element_rect(fill = '#F5F5F5', colour = 'darkgrey'),
        #legend.position="none",,
        aspect.ratio = 1/1.5)

ggsave("Figure_output/Node_Metrics.pdf")
```

Standardized metrics:
```{r}
node_metrics %>% 
  pivot_longer(2:3, names_to = "Metrics", values_to = "Value") %>%
  mutate(Functional_Group = factor(FG, levels = functional_groups)) %>%
  mutate(Metrics = factor(Metrics, levels = Metrics_levels)) %>% 
  filter(is.na(FG) == FALSE) %>% 
  group_by(Functional_Group, Metrics) %>% 
  mutate(Value = Value/mean(Value)) %>% 
  ggplot() +
  geom_point(aes(Date_max, Value, colour = Functional_Group), size = 0.5) +
  geom_smooth(aes(Date_max, Value, colour = Functional_Group), se = FALSE, size = 0.8 ) +
  facet_wrap("Metrics", scales = "free") +
  scale_color_manual(values = colors) +
      theme_minimal_vgrid(font_size = 10) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        #strip.text = element_text(face = "italic"),
        panel.background = element_rect(fill = '#F5F5F5', colour = 'darkgrey'),
        #legend.position="none",,
        aspect.ratio = 1/1.5)

ggsave("Figure_output/Trophic_Levels.pdf")
```


```{r}
node_metrics

node_metrics %>%
  filter(is.na(FG) == FALSE) %>% 
  #filter(FG %!in% c("Phytoplankton", "Protozooplankton")) %>% 
  mutate(Functional_Group = factor(FG, levels = functional_groups)) %>%
  group_by(Functional_Group) %>% 
  mutate(Num_Paths = Num_Paths/mean(Num_Paths)) %>% 
  ggplot() +
  geom_point(aes(Date_max, Num_Paths, colour = Functional_Group)) +
  geom_smooth(aes(Date_max, Num_Paths, colour = Functional_Group), se = FALSE, size = 0.8) +
  scale_color_manual(values = colors) +
      theme_minimal_vgrid(font_size = 10) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        #strip.text = element_text(face = "italic"),
        panel.background = element_rect(fill = '#F5F5F5', colour = 'darkgrey'),
        #legend.position="none",,
        aspect.ratio = 1/1.5)

ggsave("Figure_output/Taxa_Richness.pdf")
```
### Taxa richness
```{r}
Node_Data %>%
  mutate(Month_max = month(Date_max, abbr = FALSE )) %>% 
  group_by(FG, Month_max) %>% 
  summarise(n = n()) %>%
  ungroup() %>% 
  group_by(FG) %>% 
  mutate(n = n/max(n)) %>% 
  mutate(Functional_Group = factor(FG, levels = functional_groups)) %>% 
  ggplot() +
  geom_point(aes(month(Month_max), n, colour = Functional_Group)) +
  geom_smooth(aes(month(Month_max), n, colour = Functional_Group), se = FALSE) +
    scale_color_manual(values = colors) +
      theme_minimal_vgrid(font_size = 10) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        #strip.text = element_text(face = "italic"),
        panel.background = element_rect(fill = '#F5F5F5', colour = 'darkgrey'),
        #legend.position="none",,
        aspect.ratio = 1/1.5)

ggsave("Figure_output/Taxa_Richness.pdf")
```

```{r}
node_metrics %>% 
  arrange(-Eigenvector) %>% 
  filter(FG == "Protozooplankton")
```
# test plot protozoans

```{r}
seasonal_data_all %>% 
  select(Sample_date, Chromidina) %>%  plot

seasonal_data_all
```



# ___ OLD CODE ___




# 6. PLot Heatmaps

```{r}
results %>% 
  filter(Max_Correlation > 0.,
         Lag >= 0) %>%
  filter(Species2 == "Centropages abdominalis",
         Species1 %in% heterotrophic_protists)
  

```









## Zooplankton to Phytoplankton
```{r}
adj_matrix

TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()



correlation_matrix <- v[rownames(adj_matrix) %in% Dataset_COI$Taxa,
                                         colnames(adj_matrix) %in% TaxaX]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]

pdf(file = "Figure_output/Corr_Plots/Metazooplankton_Phytoplankton.pdf", width = 20, height = 20)
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Metazooplankton - Phytoplankon lagged interactions"
)
dev.off()

```
## Zooplankton to Heterotrophic 
```{r}
TaxaX <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()


correlation_matrix <- v[rownames(adj_matrix) %in% Dataset_COI$Taxa,
                                         colnames(adj_matrix) %in% TaxaX]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]


pdf(file = "Figure_output/Corr_Plots/Metazooplankton_Protozooplankton.pdf", width = 20, height = 20)
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Metazooplankton and Protozooplankton lagged interactions"
)

dev.off()
```
## Heterotrophic to phytoplankton
```{r}
TaxaY <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- adj_matrix[rownames(adj_matrix) %in% TaxaY,
                                         colnames(adj_matrix) %in% TaxaX]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]

pdf(file = "Figure_output/Corr_Plots/Protozooplankton_Phytoplankton.pdf", width = 20, height = 20)
# Create the heatmap with a custom color scale
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Protozooplankton and Phytoplankton Lagged Interactions"
)

dev.off()
```

## Carnivorous to non-carnivorous Metazooplankton
```{r}
TaxaY <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- adj_matrix[rownames(adj_matrix) %in% carnivorous_zooplankton,
                                         colnames(adj_matrix) %in% omnivorous_filter_feeding_zooplankton]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]

pdf(file = "Figure_output/Corr_Plots/Carnivorous_Omnivorous_Mesozooplankton.pdf", width = 10, height = 10)
# Create the heatmap with a custom color scale
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Carnivorous to Non-Canivorous Metazooplankton Lagged Interactions"
)

dev.off()


```



## Fish to Zooplankton
```{r}
adj_matrix

correlation_matrix <- adj_matrix[rownames(adj_matrix) %in% pelagic_fish,
                                         colnames(adj_matrix) %in% Dataset_COI$Taxa]
colnames(adj_matrix) %in% pelagic_fish
correlation_matrix
pelagic_fish

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]

pdf(file = "Figure_output/Corr_Plots/Fish_Mesozooplankton.pdf", width = 10, height = 10)
# Create the heatmap with a custom color scale
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Pelagic Fish to Metazooplankton Lagged Interactions"
)

dev.off()
```


## Phytoplankton
```{r}
correlation_matrix_all <- cor(seasonal_data_all %>% select(-Sample_date), use = "pairwise.complete.obs")


TaxaY <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% TaxaX,
                                         colnames(correlation_matrix_all) %in% TaxaX]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]


pdf(file = "Figure_output/Corr_Plots/Phytoplankton.pdf", width = 20, height = 20)
# Create the heatmap with a custom color scale
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Phytoplankton Clusters - No Lagg"
)

dev.off()
```

## Zooplankton
```{r}
TaxaY <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% Dataset_COI$Taxa,
                                         colnames(correlation_matrix_all) %in% Dataset_COI$Taxa]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]

pdf(file = "Figure_output/Corr_Plots/Mesozooplankton.pdf", width = 20, height = 20)
# Create the heatmap with a custom color scale
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Mesozooplankton Clusters - No Lagg"
)

dev.off()
```


## Fish
```{r}

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% pelagic_fish,
                                         colnames(correlation_matrix_all) %in% pelagic_fish]
correlation_matrix
# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]

pdf(file = "Figure_output/Corr_Plots/Fish.pdf", width = 20, height = 20)
# Create the heatmap with a custom color scale
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Fish Clusters - No Lagg"
)

dev.off()
```


## Microplankton
```{r}
TaxaY <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% TaxaY,
                                         colnames(correlation_matrix_all) %in% TaxaY]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]


pdf(file = "Figure_output/Corr_Plots/Protozooplankton.pdf", width = 20, height = 20)
# Create the heatmap with a custom color scale
pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Protozooplankton Clusters - No Lagg"
)

dev.off()
```


# Seasonal Clusters


## Phytoplankton

```{r}
set.seed(200)

correlation_matrix_all <- cor(seasonal_data_all %>% select(-Sample_date), use = "pairwise.complete.obs")

TaxaX <- taxatab_18S %>%
  filter(Class %in% phototrophic_plankton) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% TaxaX,
                                         colnames(correlation_matrix_all) %in% TaxaX]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]





# Sample code to perform K-means clustering on a correlation matrix

# Assuming `correlation_matrix` is your correlation matrix:
# Step 1: Convert the correlation matrix to a dissimilarity matrix
# Using 1 - correlation to represent distance
dissimilarity_matrix <- 1 - correlation_matrix

# Step 2: Choose the number of clusters (k)
# Let's assume you want 3 clusters (you can experiment with other values of k)
k <- 7

# Step 3: Perform K-means clustering
# Run K-means on the transformed dissimilarity matrix
# Use the `as.dist` function to treat it as a distance matrix
kmeans_result <- kmeans(as.dist(dissimilarity_matrix), centers = k, nstart = 25)

# Step 4: Check the results
# Cluster assignment for each variable
cluster_df <- tibble(
  Taxa = names(kmeans_result$cluster),
  Cluster = as.factor(kmeans_result$cluster)
)



library(cluster)

# Compute silhouette values for different k
silhouette_scores <- vector("numeric", length = 10)
for (i in 2:10) {  # k = 1 is not meaningful for silhouette analysis
  kmeans_result <- kmeans(correlation_matrix, centers = i)
  silhouette_scores[i] <- mean(silhouette(kmeans_result$cluster, dist(correlation_matrix))[, 3])
}

# Plot silhouette scores
plot(2:10, silhouette_scores[2:10], type = "b", main = "Silhouette Method",
     xlab = "Number of clusters", ylab = "Average Silhouette Score")





pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Phytoplankton Clusters - No Lagg"
)


plot_cluster <- function(cluster) {
  
  seasonal_data_all %>% 
  pivot_longer(!Sample_date, names_to = "Taxa", values_to = "Abundance") %>%
  left_join(cluster_df, by = "Taxa") %>% 
  filter(Cluster == cluster) %>% 
  ggplot() +
  geom_line(aes(Sample_date, Abundance)) +
  facet_wrap("Taxa", scales = "free")
  
  
}
```


```{r}
plot_cluster(1)
plot_cluster(2)
plot_cluster(3)
plot_cluster(4)
plot_cluster(5)
plot_cluster(6)
plot_cluster(7)
```


## Protozooplankton

```{r}
set.seed(200)

correlation_matrix_all <- cor(seasonal_data_all %>% select(-Sample_date), use = "pairwise.complete.obs")

TaxaX <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% TaxaX,
                                         colnames(correlation_matrix_all) %in% TaxaX]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]


# Sample code to perform K-means clustering on a correlation matrix

# Assuming `correlation_matrix` is your correlation matrix:
# Step 1: Convert the correlation matrix to a dissimilarity matrix
# Using 1 - correlation to represent distance
dissimilarity_matrix <- 1 - correlation_matrix

# Step 2: Choose the number of clusters (k)
# Let's assume you want 3 clusters (you can experiment with other values of k)
k <- 6

# Step 3: Perform K-means clustering
# Run K-means on the transformed dissimilarity matrix
# Use the `as.dist` function to treat it as a distance matrix
kmeans_result <- kmeans(as.dist(dissimilarity_matrix), centers = k, nstart = 25)

# Step 4: Check the results
# Cluster assignment for each variable
cluster_df <- tibble(
  Taxa = names(kmeans_result$cluster),
  Cluster = as.factor(kmeans_result$cluster)
)



library(cluster)

# Compute silhouette values for different k
silhouette_scores <- vector("numeric", length = 10)
for (i in 2:10) {  # k = 1 is not meaningful for silhouette analysis
  kmeans_result <- kmeans(correlation_matrix, centers = i)
  silhouette_scores[i] <- mean(silhouette(kmeans_result$cluster, dist(correlation_matrix))[, 3])
}

# Plot silhouette scores
plot(2:10, silhouette_scores[2:10], type = "b", main = "Silhouette Method",
     xlab = "Number of clusters", ylab = "Average Silhouette Score")





pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Phytoplankton Clusters - No Lagg"
)


plot_cluster <- function(cluster) {
  
  seasonal_data_all %>% 
  pivot_longer(!Sample_date, names_to = "Taxa", values_to = "Abundance") %>%
  left_join(cluster_df, by = "Taxa") %>% 
  filter(Cluster == cluster) %>% 
  ggplot() +
  geom_line(aes(Sample_date, Abundance)) +
  facet_wrap("Taxa", scales = "free")
  
  
}
```


```{r}
plot_cluster(1)
plot_cluster(2)
plot_cluster(3)
plot_cluster(4)
plot_cluster(5)
plot_cluster(6)
```

## Zooplankton

```{r}
set.seed(200)

correlation_matrix_all <- cor(seasonal_data_all %>% select(-Sample_date), use = "pairwise.complete.obs")

TaxaX <- taxatab_18S %>%
  filter(Class %in% heterotrophic_protists) %>% 
  pull(Taxa) %>% unique()

correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% Dataset_COI$Taxa,
                                         colnames(correlation_matrix_all) %in% Dataset_COI$Taxa]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]


# Sample code to perform K-means clustering on a correlation matrix

# Assuming `correlation_matrix` is your correlation matrix:
# Step 1: Convert the correlation matrix to a dissimilarity matrix
# Using 1 - correlation to represent distance
dissimilarity_matrix <- 1 - correlation_matrix

# Step 2: Choose the number of clusters (k)
# Let's assume you want 3 clusters (you can experiment with other values of k)
k <- 6

# Step 3: Perform K-means clustering
# Run K-means on the transformed dissimilarity matrix
# Use the `as.dist` function to treat it as a distance matrix
kmeans_result <- kmeans(as.dist(dissimilarity_matrix), centers = k, nstart = 25)

# Step 4: Check the results
# Cluster assignment for each variable
cluster_df <- tibble(
  Taxa = names(kmeans_result$cluster),
  Cluster = as.factor(kmeans_result$cluster)
)



library(cluster)

# Compute silhouette values for different k
silhouette_scores <- vector("numeric", length = 10)
for (i in 2:10) {  # k = 1 is not meaningful for silhouette analysis
  kmeans_result <- kmeans(correlation_matrix, centers = i)
  silhouette_scores[i] <- mean(silhouette(kmeans_result$cluster, dist(correlation_matrix))[, 3])
}

# Plot silhouette scores
plot(2:10, silhouette_scores[2:10], type = "b", main = "Silhouette Method",
     xlab = "Number of clusters", ylab = "Average Silhouette Score")





pheatmap(correlation_matrix,
         color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Phytoplankton Clusters - No Lagg"
)


plot_cluster <- function(cluster) {
  
  seasonal_data_all %>% 
  pivot_longer(!Sample_date, names_to = "Taxa", values_to = "Abundance") %>%
  left_join(cluster_df, by = "Taxa") %>% 
  filter(Cluster == cluster) %>% 
  ggplot() +
  geom_line(aes(Sample_date, Abundance)) +
  facet_wrap("Taxa", scales = "free")
  
  
}
```


```{r}
plot_cluster(1)
plot_cluster(2)
plot_cluster(3)
plot_cluster(4)
plot_cluster(5)
plot_cluster(6)
```


## Fish

```{r}
set.seed(200)

correlation_matrix_all <- cor(seasonal_data_all %>% select(-Sample_date), use = "pairwise.complete.obs")



correlation_matrix <- correlation_matrix_all[rownames(correlation_matrix_all) %in% pelagic_fish,
                                         colnames(correlation_matrix_all) %in% pelagic_fish]

# Perform hierarchical clustering on the correlation matrix
correlation_matrix[is.na(correlation_matrix) | is.nan(correlation_matrix) | is.infinite(correlation_matrix)] <- 0
taxa_order <- hclust(as.dist(1 - correlation_matrix))$order
ordered_taxa <- rownames(correlation_matrix)[taxa_order]


# Sample code to perform K-means clustering on a correlation matrix

# Assuming `correlation_matrix` is your correlation matrix:
# Step 1: Convert the correlation matrix to a dissimilarity matrix
# Using 1 - correlation to represent distance
dissimilarity_matrix <- 1 - correlation_matrix

# Step 2: Choose the number of clusters (k)
# Let's assume you want 3 clusters (you can experiment with other values of k)
k <- 6

# Step 3: Perform K-means clustering
# Run K-means on the transformed dissimilarity matrix
# Use the `as.dist` function to treat it as a distance matrix
kmeans_result <- kmeans(as.dist(dissimilarity_matrix), centers = k, nstart = 25)

# Step 4: Check the results
# Cluster assignment for each variable
cluster_df <- tibble(
  Taxa = names(kmeans_result$cluster),
  Cluster = as.factor(kmeans_result$cluster)
)



library(cluster)

# Compute silhouette values for different k
silhouette_scores <- vector("numeric", length = 10)
for (i in 2:10) {  # k = 1 is not meaningful for silhouette analysis
  kmeans_result <- kmeans(correlation_matrix, centers = i)
  silhouette_scores[i] <- mean(silhouette(kmeans_result$cluster, dist(correlation_matrix))[, 3])
}

# Plot silhouette scores
plot(2:10, silhouette_scores[2:10], type = "b", main = "Silhouette Method",
     xlab = "Number of clusters", ylab = "Average Silhouette Score")





pheatmap(correlation_matrix,
         #color = colorRampPalette(c("white", "pink"))(50),  # Custom color scale from white to blue
         #breaks = c(0, 0.5, seq(0.5001, 1, length.out = 49)),  # Ensure breaks are unique by starting slightly above 0.5
         show_rownames = TRUE,
         show_colnames = TRUE,
         #cluster_rows = TRUE,
         #cluster_cols = TRUE,
         #legend = TRUE,
         main = "Phytoplankton Clusters - No Lagg"
)


plot_cluster <- function(cluster) {
  
  seasonal_data_all %>% 
  pivot_longer(!Sample_date, names_to = "Taxa", values_to = "Abundance") %>%
  left_join(cluster_df, by = "Taxa") %>% 
  filter(Cluster == cluster) %>% 
  ggplot() +
  geom_line(aes(Sample_date, Abundance)) +
  facet_wrap("Taxa", scales = "free")
  
  
}
```


```{r}
plot_cluster(1)
plot_cluster(2)
plot_cluster(3)
plot_cluster(4)
plot_cluster(5)
plot_cluster(6)
```


```{r}
seasonal_data_all %>% 
  pivot_longer(!Sample_date, names_to = "Taxa", values_to = "Abundance") %>%
  left_join(cluster_df, by = "Taxa") %>% 
  filter(Cluster == 1) %>%
  pull(Taxa) %>% unique()
```




# Plot Seasonality

```{r}

  
  
seasonal_data_all %>% 
  pivot_longer(!Sample_date, names_to = "Taxa", values_to = "Abundance") %>% 
  filter(Taxa %in% omnivorous_filter_feeding_zooplankton) %>% 
  ggplot() +
  geom_line(aes(Sample_date, Abundance)) +
  facet_wrap("Taxa", scales = "free")
```




# ___OLD CODE___

## Plot Heatmap in ggplot
```{r}
taxatab_18S <- Dataset_18S %>%
  group_by(Phylum, Species1 = Class) %>% 
  summarise()

corr_data <- ccf_results %>%
  dplyr::filter(Species2 %in% Interpolated_COI_Data$Taxa,
         Species1 %in% Interpolated_COI_Data$Taxa) %>% 
  filter(Max_Correlation > 0.4,
         Lag_At_Max_Correlation >= 0) %>% 
  left_join(taxatab_18S, by = c("Species1"))


corr_data %>% 
  ggplot() +
  geom_raster(aes(Species2, Species1, fill = Max_Correlation)) +
  geom_text(aes(Species2, Species1, label = Lag_At_Max_Correlation), size = 2) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
                       midpoint = 0.4, limit = c(0.4, 1), name = "Correlation") +
  theme_minimal(8) +
  labs(x = "Zooplankton", y = "Microeukaryotes", title = "Correlation between Taxa with 0-4 weeks lag") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)

ggsave("Figure_output/Correlation_plot.pdf")

```




## Correlation by sample

```{r}

all_community_mat <- bind_rows(
  filter(Dataset_18S, Hakai_ID %in% Dataset_COI$Hakai_ID),
  filter(Dataset_COI, Hakai_ID %in% Dataset_COI$Hakai_ID)) %>% 
  group_by(Hakai_ID, Taxa) %>%
  summarize(RA = mean(RA)) %>% 
  pivot_wider(names_from = Taxa, values_from = RA) %>% 
  column_to_rownames("Hakai_ID") %>%
  mutate(across(everything(), ~ replace_na(.x, 0))) %>% 
  as.matrix()

all_community_mat

all_cor_matrix <- cor(all_community_mat, method = "pearson", use = "pairwise.complete.obs")

cor_df <- as.data.frame(all_cor_matrix) %>%
  rownames_to_column(var = "Taxa1") %>%
  pivot_longer(-Taxa1, names_to = "Taxa2", values_to = "Correlation") %>%
  filter(Taxa1 %!in% Dataset_COI$Taxa,
         Taxa2 %!in% Dataset_18S$Taxa)

cor_df %>%
  filter(Correlation > 0.4) %>% 
  ggplot() +
  geom_raster(aes(Taxa2, Taxa1, fill = Correlation)) +
  #geom_text(aes(Taxa2, Taxa1, label = Lag_At_Max_Correlation), size = 2) +
  #scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
  #                     midpoint = 0.4, limit = c(0.4, 1), name = "Correlation") +
  theme_minimal(8) +
  labs(x = "Zooplankton", y = "Microeukaryotes", title = "Correlation between Taxa with 0-4 weeks lag") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)
  
cor_df
```



