---
title: "R Notebook"
---


```{r}
# Data import

library(readxl)
#library(phyloseq)


# All data format modifications and filtering and plotting
library(tidyverse)
'%!in%' <- function(x,y)!('%in%'(x,y))
source("Code/Functions.R")
library(zoo)

# Statistical tools
library(vegan)

# Specific plotting tools
library(ggpubr)
library(paletteer)
library(viridis)
library(cowplot)
#library(ggnewscale)
source("Code/scales_and_themes.R")

# Make sure tidyverse is the latest loaded package
detach("package:tidyverse", unload=TRUE)
library(tidyverse)


# Custom color scale for depths
depth_colors <- c("0" = "#add8e6",    # Light blue for 0 m
                  "5" = "#87ceeb",    # Sky blue for 5 m
                  "30" = "#4682b4",   # Steel blue for 30 m
                  "100" = "#00008b")  # Dark blue for 100 m
```


```{r}
Dataset_COI <- readRDS("Data_output/Data_merge/COI_glomm_indexed.rds")
Dataset_18S <- readRDS("Data_output/Data_merge/pr2_18S_glomm_indexed.rds")
Dataset_Bottle <- readRDS("Data_output/Data_merge/bottle_file.rds")

Dataset_18S_glom <- Dataset_18S %>%
    filter(Phylum != "Metazoa") %>%
#    filter(Phylum %in% c(
#    "Chlorophyta",
#    "Cryptophyta",
#    "Dinoflagellata",
#    "Haptophyta",
#    "Ochrophyta"
#  )) %>% 
  #filter(is.na(Genus)==FALSE) %>% 
  group_by(Library_ID, Sample_date, Line_out_depth, Taxa = Class) %>%
  summarise(Abundance = sum(Abundance)) %>% 
  index_RA(Library_ID, Taxa, Abundance, Sample_date, Line_out_depth)
  
```



# 1. Interpolation

	1.	Fill Missing Combinations with Zeros: The complete(Sample_ID, Sample_date, Taxa, Sampling_Depth, fill = list(Relative_Abundance = 0)) line ensures that any combination without observed data is filled with 0 for Relative_Abundance.
	2.	Generate Weekly Dates: complete(Sample_date = seq.Date(min(Sample_date), max(Sample_date), by = "week")) generates weekly dates for each Taxa and Sampling_Depth combination.
	3.	Interpolate Missing Values: na.approx() fills in any NAs generated by the new weekly dates in Sample_date.


### COI
```{r}

start_date <- as.Date("2017-01-01")
end_date <- as.Date("2019-02-15")


# Complete the data frame with all combinations of Sample_ID, Sample_date, Taxa, and Sampling_Depth
complete_data <- Dataset_COI %>%
  filter(Line_out_depth %in% c(0,5,30,100)) %>%
  mutate(Line_out_depth = factor(Line_out_depth, levels = c("0", "5", "30", "100"))) %>% 
  mutate(Sample_date = as.Date(Sample_date)) %>% 
  filter(Sample_date >= start_date & Sample_date <= end_date) %>% 
  complete(Sample_date, Taxa, Line_out_depth, fill = list(RA_Index = 0)) %>% 
  # Ensure dates are properly ordered after completion
  arrange(Sample_date, Taxa, Line_out_depth)


strict_weekly_dates <- seq.Date(from = min(complete_data$Sample_date),
                                to = max(complete_data$Sample_date),
                                by = "week")

smoothed_data <- complete_data %>%
  # Expand to include all combinations of Taxa, Depth, and weekly dates
  group_by(Taxa, Line_out_depth) %>%
  complete(Sample_date = strict_weekly_dates) %>%
  arrange(Sample_date) %>%
  # Interpolate missing values in RA_Index within each group
  mutate(RA_Index = ifelse(is.na(RA_Index),
                           na.approx(RA_Index, rule = 2, na.rm = FALSE),
                           RA_Index)) %>%
  # Apply 2-week rolling mean for smoothing (7-day intervals)
  mutate(Smoothed_Relative_Abundance = rollmean(RA_Index, k = 2, fill = NA, align = "center")) %>%
  ungroup() %>%
  # Filter to include only rows with Smoothed_Relative_Abundance calculated
  filter(!is.na(Smoothed_Relative_Abundance)) %>% 
  filter(Sample_date %in% strict_weekly_dates)

Interpolated_COI_Data <- smoothed_data %>% 
  group_by(Taxa, Sample_date) %>% 
  summarise(Abundance = sum(Smoothed_Relative_Abundance)/4)
```

### 18S

```{r}


# Complete the data frame with all combinations of Sample_ID, Sample_date, Taxa, and Sampling_Depth
complete_data <- Dataset_18S_glom %>%
  filter(Line_out_depth %in% c(0,5,30,100)) %>%
  mutate(Line_out_depth = factor(Line_out_depth, levels = c("0", "5", "30", "100"))) %>%
  #filter(Phylum != "Metazoa") %>%
  #  filter(Phylum %in% c(
  #  "Chlorophyta",
  #  "Cryptophyta",
  #  "Dinoflagellata",
  #  "Haptophyta",
  #  "Ochrophyta"
  #)) %>% 
  #filter(is.na(Genus)==FALSE) %>% 
  mutate(Sample_date = as.Date(Sample_date)) %>% 
  complete(Sample_date, Taxa, Line_out_depth, fill = list(RA_Index = 0)) %>%
  filter(Sample_date >= start_date & Sample_date <= end_date) %>% 
  # Ensure dates are properly ordered after completion
  arrange(Sample_date, Taxa, Line_out_depth)


# NOTE: Commented out to reuse the series for COI
#strict_weekly_dates <- seq.Date(from = min(complete_data$Sample_date),
#                                to = max(complete_data$Sample_date),
#                                by = "week")

smoothed_data <- complete_data %>%
  # Expand to include all combinations of Taxa, Depth, and weekly dates
  group_by(Taxa, Line_out_depth) %>%
  complete(Sample_date = strict_weekly_dates) %>%
  arrange(Sample_date) %>%
  # Interpolate missing values in RA_Index within each group
  mutate(RA_Index = ifelse(is.na(RA_Index),
                           na.approx(RA_Index, rule = 2, na.rm = FALSE),
                           RA_Index)) %>%
  # Apply 2-week rolling mean for smoothing (7-day intervals)
  mutate(Smoothed_Relative_Abundance = rollmean(RA_Index, k = 4, fill = NA, align = "center")) %>%
  ungroup() %>%
  # Filter to include only rows with Smoothed_Relative_Abundance calculated
  filter(!is.na(Smoothed_Relative_Abundance)) %>% 
  filter(Sample_date %in% strict_weekly_dates)

Interpolated_18S_Data <- smoothed_data %>% 
  group_by(Taxa, Sample_date) %>% 
  summarise(Abundance = sum(Smoothed_Relative_Abundance)/4)
```



# 2. STL - Extract seasonal data
```{r}

### COI
seasonal_data_COI <- Interpolated_COI_Data %>%
  # Extract seasonal data from STL time series, per taxa
  group_by(Taxa) %>%
  do({
    ts_data <- ts(.$Abundance, start = c(2017, 8), frequency = 52)  # Weekly frequency
    stl_decomposition <- stl(ts_data, s.window = "periodic")
    data.frame(Sample_date = .$Sample_date, Seasonal = stl_decomposition$time.series[, "seasonal"])
  }) %>%
  ungroup() %>%
  spread(key = Taxa, value = Seasonal) %>%
  drop_na()


#### 18S
seasonal_data_18S <- Interpolated_18S_Data %>%
  # Extract seasonal data from STL time series, per taxa
  group_by(Taxa) %>%
  do({
    ts_data <- ts(.$Abundance, start = c(2017, 8), frequency = 52)  # Weekly frequency
    stl_decomposition <- stl(ts_data, s.window = "periodic")
    data.frame(Sample_date = .$Sample_date, Seasonal = stl_decomposition$time.series[, "seasonal"])
  }) %>%
  ungroup() %>%
  spread(key = Taxa, value = Seasonal) %>%
  drop_na()

### Combine
seasonal_data_all <-left_join(seasonal_data_18S, seasonal_data_COI, by = "Sample_date")
seasonal_data_all
```

# 3. Correlation
```{r}

correlation_matrix <- cor(seasonal_data_all %>% select(-Sample_date), use = "pairwise.complete.obs")

correlation_matrix

# Convert the correlation matrix to a tidy data frame
correlation_data <- as.data.frame(as.table(correlation_matrix)) %>%
  rename(Taxa1 = Var1, Taxa2 = Var2, Correlation = Freq) %>%
  filter(Taxa1 != Taxa2) %>%        # Remove self-correlations
  filter(Correlation > 0.5) %>%     # Filter for strong correlations
  filter(Taxa1 %in% Interpolated_COI_Data$Taxa,
         Taxa2 %in% Interpolated_18S_Data$Taxa)



correlation_data %>% 
  ggplot() +
  geom_raster(aes(Taxa1, Taxa2, fill = Correlation)) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), name = "Correlation") +
  theme_minimal() +
  labs(x = "Taxa 2", y = "Taxa 1", title = "Correlation between Taxa") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)
```


# 4. Correlation with lags

```{r}
# Define the matrix (excluding date column)
species_matrix <- as.matrix(seasonal_data_all %>% select(-Sample_date))

# Define maximum lag
max_lag <- 4

# Prepare an empty data frame to store results
ccf_results <- data.frame(
  Species1 = character(),
  Species2 = character(),
  Max_Correlation = numeric(),
  Lag_At_Max_Correlation = integer()
)

# Loop through each pair of species
species_names <- colnames(species_matrix)
for (i in 1:(ncol(species_matrix) - 1)) {
  for (j in (i + 1):ncol(species_matrix)) {
    
    # Get time series data for the two species
    ts1 <- species_matrix[, i]
    ts2 <- species_matrix[, j]
    
    # Compute cross-correlation
    ccf_result <- ccf(ts1, ts2, lag.max = max_lag, plot = FALSE)
    
    # Find the lag with the highest absolute cross-correlation
    max_corr_idx <- which.max(abs(ccf_result$acf))
    max_corr <- ccf_result$acf[max_corr_idx]
    lag_at_max_corr <- ccf_result$lag[max_corr_idx]
    
    # Store the result
    ccf_results <- rbind(ccf_results, data.frame(
      Species1 = species_names[i],
      Species2 = species_names[j],
      Max_Correlation = max_corr,
      Lag_At_Max_Correlation = lag_at_max_corr
    ))
  }
}
```


```{r}
taxatab_18S <- Dataset_18S %>%
  group_by(Phylum, Species1 = Class) %>% 
  summarise()

corr_data <- ccf_results %>%
  dplyr::filter(Species2 %in% Interpolated_COI_Data$Taxa,
         Species1 %in% Interpolated_18S_Data$Taxa) %>% 
  filter(Max_Correlation > 0.4,
         Lag_At_Max_Correlation >= 0) %>% 
  left_join(taxatab_18S, by = c("Species1"))


corr_data %>% 
  ggplot() +
  geom_raster(aes(Species2, Species1, fill = Max_Correlation)) +
  geom_text(aes(Species2, Species1, label = Lag_At_Max_Correlation), size = 2) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
                       midpoint = 0.4, limit = c(0.4, 1), name = "Correlation") +
  theme_minimal(8) +
  labs(x = "Zooplankton", y = "Microeukaryotes", title = "Correlation between Taxa with 0-4 weeks lag") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)

ggsave("Figure_output/Correlation_plot.pdf")

```


