---
title: "4_ZP_Community_Analysis"

---

```{r}
# Data import

library(readxl)
#library(phyloseq)


# All data format modifications and filtering and plotting
library(tidyverse)
'%!in%' <- function(x,y)!('%in%'(x,y))
source("Code/Functions.R")
library(zoo)

# Statistical tools
library(vegan)

# Specific plotting tools
library(ggpubr)
library(paletteer)
library(viridis)
library(cowplot)
#library(ggnewscale)
source("Code/scales_and_themes.R")

# Make sure tidyverse is the latest loaded package
detach("package:tidyverse", unload=TRUE)
library(tidyverse)


# Custom color scale for depths
depth_colors <- c("0" = "#add8e6",    # Light blue for 0 m
                  "5" = "#87ceeb",    # Sky blue for 5 m
                  "30" = "#4682b4",   # Steel blue for 30 m
                  "100" = "#00008b")  # Dark blue for 100 m
```


```{r}
Dataset_COI <- readRDS("Data_output/Data_merge/COI_glomm_indexed.rds")
#Dataset_18S <- readRDS("Data_output/Data_merge/pr2_18S_glomm_indexed.rds")
Dataset_Bottle <- readRDS("Data_output/Data_merge/bottle_file.rds")

  
```



# 1. Plot eDNA profiles:

```{r}
# Complete the data frame with all combinations of Sample_ID, Sample_date, Taxa, and Sampling_Depth
complete_data <- Dataset_COI %>%
  filter(Line_out_depth %in% c(0,5,30,100)) %>%
  mutate(Line_out_depth = factor(Line_out_depth, levels = c("0", "5", "30", "100"))) %>% 
  mutate(Sample_date = as.Date(Sample_date)) %>% 
  #filter(Sample_date >= start_date & Sample_date <= end_date) %>% 
  complete(Sample_date, Taxa, Line_out_depth, fill = list(RA_Index = 0)) %>% 
  # Ensure dates are properly ordered after completion
  arrange(Sample_date, Taxa, Line_out_depth)


strict_weekly_dates <- seq.Date(from = min(complete_data$Sample_date),
                                to = max(complete_data$Sample_date),
                                by = "day")

smoothed_data <- complete_data %>%
  # Expand to include all combinations of Taxa, Depth, and weekly dates
  group_by(Taxa, Line_out_depth) %>%
  complete(Sample_date = strict_weekly_dates) %>%
  arrange(Sample_date) %>%
  # Interpolate missing values in RA_Index within each group
  mutate(RA_Index = ifelse(is.na(RA_Index),
                           na.approx(RA_Index, rule = 2, na.rm = FALSE),
                           RA_Index)) %>%
  # Apply 2-week rolling mean for smoothing (7-day intervals)
  mutate(Smoothed_Relative_Abundance = rollmean(RA_Index, k = 14, fill = NA, align = "center")) %>%
  ungroup() %>%
  # Filter to include only rows with Smoothed_Relative_Abundance calculated
  filter(!is.na(Smoothed_Relative_Abundance)) %>% 
  filter(Sample_date %in% strict_weekly_dates)




```



```{r}
smoothed_data %>%
  #filter(Taxa == "Acartia longiremis") %>% 
  ggplot() +
  geom_line(aes(Sample_date, Smoothed_Relative_Abundance, colour = Line_out_depth)) +
  scale_colour_manual(values = depth_colors) +
  facet_wrap("Taxa")


smoothed_data %>%
  mutate(Line_out_depth = factor(c(100, 30, 5, 0))) %>% 
  #filter(Taxa == "Acartia longiremis") %>% 
  ggplot() +
  geom_raster(aes(Sample_date, Line_out_depth, fill = Smoothed_Relative_Abundance)) +
  #scale_colour_manual(values = depth_colors) +
  facet_wrap("Taxa")


smoothed_data %>%
  mutate(eDNA_index = Smoothed_Relative_Abundance) %>% 
  mutate(Line_out_depth = factor(Line_out_depth, levels = c(100, 30, 5, 0))) %>% 
  ggplot() +
  geom_raster(aes(Sample_date, Line_out_depth, fill = eDNA_index)) +
  geom_vline(xintercept = datebreaks, color = "#F5F5F5", size = 0.2, alpha = 0.8) +
  geom_vline(xintercept = yearbreaks, color = "grey", size = 0.2, alpha = 0.8) +
  
  #geom_text(data = trait_summary_mod,
  #          aes(x = date("2018-07-01"), y = "0",
  #              label = trophic_cluster),
  #          size = 2) +
  facet_wrap("Taxa", ncol = 4) +
  #scale_y_reverse() +
  scale_x_date(date_labels="%b %y", breaks  = datebreaks) +
  #scale_fill_viridis(option = "H") +
  scale_fill_gradient2(high = '#4daf4a',
                       low = "white") +
  theme_minimal_vgrid(font_size = 8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        strip.text = element_text(face = "italic"),
        panel.background = element_rect(fill = '#F5F5F5', colour = 'darkgrey'),
        #legend.position="none",,
        aspect.ratio = 1/4)

ggsave("Figure_output/eDNA_profiles_ZP.pdf")
```


# 2. CCA 


## Prepare data
```{r}
# Prepare comm and env datasets:

comm <- Dataset_COI %>%
  #filter(Line_out_depth %!in% c(0)) %>% 
  dplyr::group_by(Hakai_ID, Taxa) %>% 
  summarise(Index = mean(RA_Index)) %>%
  pivot_wider(names_from = "Taxa", values_from = "Index") %>%
  mutate_all(~replace(., is.na(.), 0)) %>% 
  filter(Hakai_ID != "QMIC2470") %>%          ####### WARNING - FILTERED OUTLIER SAMPLE. Verify!
  arrange(Hakai_ID) %>%
  column_to_rownames("Hakai_ID")

comm

env <- Dataset_Bottle %>% 
  arrange(hakai_id) %>%
  mutate(Depth = factor(line_out_depth),
         Month = month(date)) %>%
  filter(hakai_id %in% row.names(comm)) %>% 
  column_to_rownames("hakai_id") %>% 
  # Chl to 0
  mutate(chla_20um = replace(chla_20um, is.na(chla_20um), 0),
         chla_3um = replace(chla_3um, is.na(chla_3um), 0),
         `chla_GF/F` = replace(`chla_GF/F`, is.na(`chla_GF/F`), 0))

# Verify that rownames are identical!
tibble(comm = row.names(comm),
       env = row.names(env))
```

## Define CCA
```{r}
mod <- cca(comm ~
             Temperature + Salinity + PAR + Turbidity + O2 +
             chla_20um + chla_3um +`chla_GF/F` +
             no2_no3_um,
           data = env,
           na.action = na.exclude)
vif.cca(mod)

#mod <- cca(comm ~
#             Temperature +
#             chla_20um + chla_3um +`chla_GF/F`,
#           data = env,
#           na.action = na.exclude)
#vif.cca(mod)
```

## Extract Taxa
```{r}
# Taxa cca scores
cca_taxa <- scores(mod)$species %>% 
  as.data.frame() %>% 
  rownames_to_column("Taxa")

# Environmental cca scores
cca_variable <- scores(mod)$biplot %>%
  as.data.frame() %>% 
  rownames_to_column("Variable")


# Sample metadata...
sample_date_depth <- env %>%
  rownames_to_column("Hakai_ID") %>% 
  select(1:3)

# Sample cca scores, join with metadata
cca_sample <- scores(mod)$site %>%
  as.data.frame() %>% 
  rownames_to_column("Hakai_ID") %>% 
  left_join(sample_date_depth, by = "Hakai_ID") %>%
  mutate(line_out_depth = ifelse(line_out_depth == 250, 260, line_out_depth)) %>% 
  mutate(Depth = factor(line_out_depth,
                        levels = c(260, 100, 30, 5, 0)),
         Month = month(date, label = TRUE))
```


```{r}
library(ggrepel)

ggplot() +
  geom_point(data = cca_sample, aes(x = CCA1, y = CCA2, color = Month), size = 1, alpha = 0.8) +
  
  #geom_point(data = cca_taxa, aes(x = CCA1, y = CCA2), color = "red", size = 0.5) +
  #ggrepel::geom_text_repel(data = cca_taxa,
  #                aes(x = CCA1, y = CCA2, label = Taxa),
  #                #nudge_y = -0.05,
  #                size = 2.5, max.overlaps = 30) +
  
  geom_segment(data = cca_variable,
               aes(x = 0, y = 0, xend = CCA1*2, yend = CCA2*2),
               arrow = arrow(length = unit(0.25, "cm"))) +
  geom_text_repel(data = cca_variable,
                           aes(x = CCA1*2.5, y = CCA2*2.5, label = Variable),
                           color = "blue",
                           size = 3.5) +
  scale_color_manual(values = seasonal_colors) +
  theme_minimal(6) +
  #labs(x = "Zooplankton", y = "Microeukaryotes", title = "Correlation between Taxa with 0-4 weeks lag") +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)

ggsave("Figure_output/CCA_month.pdf", width = 4, height = 4)
```

```{r}
library(ggrepel)

ggplot() +
  geom_point(data = cca_sample, aes(x = CCA1, y = CCA2, color = Depth), size = 1, alpha = 0.8) +
  
  #geom_point(data = cca_taxa, aes(x = CCA1, y = CCA2), color = "red", size = 0.5) +
  #ggrepel::geom_text_repel(data = cca_taxa,
  #                aes(x = CCA1, y = CCA2, label = Taxa),
  #                #nudge_y = -0.05,
  #                size = 2.5, max.overlaps = 30) +
  
  geom_segment(data = cca_variable,
               aes(x = 0, y = 0, xend = CCA1*2, yend = CCA2*2),
               arrow = arrow(length = unit(0.25, "cm"))) +
  geom_text_repel(data = cca_variable,
                           aes(x = CCA1*2.5, y = CCA2*2.5, label = Variable),
                           color = "blue",
                           size = 3.5) +
  scale_color_manual(values = seasonal_colors) +
  theme_minimal(6) +
  #labs(x = "Zooplankton", y = "Microeukaryotes", title = "Correlation between Taxa with 0-4 weeks lag") +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)

ggsave("Figure_output/CCA_depth.pdf", width = 4, height = 4)
```


```{r}
```{r}
library(ggrepel)

ggplot() +
  geom_point(data = cca_sample, aes(x = CCA1, y = CCA2, color = Depth), size = 1, alpha = 0.8) +
  
  #geom_point(data = cca_taxa, aes(x = CCA1, y = CCA2), color = "red", size = 0.5) +
  ggrepel::geom_text_repel(data = cca_taxa,
                  aes(x = CCA1, y = CCA2, label = Taxa),
                  #nudge_y = -0.05,
                  size = 2.5, max.overlaps = 30) +
  
  #geom_segment(data = cca_variable,
  #             aes(x = 0, y = 0, xend = CCA1*2, yend = CCA2*2),
  #             arrow = arrow(length = unit(0.25, "cm"))) +
  #geom_text_repel(data = cca_variable,
  #                         aes(x = CCA1*2.5, y = CCA2*2.5, label = Variable),
  #                         color = "blue",
  #                         size = 3.5) +
  scale_color_manual(values = seasonal_colors) +
  theme_minimal(6) +
  #labs(x = "Zooplankton", y = "Microeukaryotes", title = "Correlation between Taxa with 0-4 weeks lag") +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)

ggsave("Figure_output/CCA_taxa.pdf", width = 4, height = 4)
```
```






```{r}
ggplot() +
  geom_point(data = cca_sample, aes(x = CCA1, y = CCA2, color = Month), size = 1, alpha = 0.8) +
  
  #geom_point(data = cca_taxa, aes(x = CCA1, y = CCA2), color = "red", size = 0.5) +
  ggrepel::geom_text_repel(data = cca_taxa,
                  aes(x = CCA1, y = CCA2, label = Taxa),
                  #nudge_y = -0.05,
                  size = 2.5, max.overlaps = 30) +
  
  geom_segment(data = cca_variable,
               aes(x = 0, y = 0, xend = CCA1, yend = CCA2),
               arrow = arrow(length = unit(0.25, "cm"))) +
  geom_text(data = cca_variable,
                           aes(x = CCA1, y = CCA2, label = Variable),
                           color = "blue",
                           size = 3.5) +
  scale_color_manual(values = seasonal_colors) +
  theme_minimal(6) +
  #labs(x = "Zooplankton", y = "Microeukaryotes", title = "Correlation between Taxa with 0-4 weeks lag") +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)

ggsave("Figure_output/CCA_month.pdf", width = 4, height = 4)






ggplot() +
  geom_point(data = cca_sample, aes(x = CCA1, y = CCA2, color = Depth), size = 1, alpha = 0.8) +
  
  #geom_point(data = cca_taxa, aes(x = CCA1, y = CCA2), color = "red", size = 0.5) +
  ggrepel::geom_text_repel(data = cca_taxa,
                  aes(x = CCA1, y = CCA2, label = Taxa),
                  #nudge_y = -0.05,
                  size = 2.5) +
  
  geom_segment(data = cca_variable,
               aes(x = 0, y = 0, xend = CCA1, yend = CCA2),
               arrow = arrow(length = unit(0.25, "cm"))) +
  ggrepel::geom_text_repel(data = cca_variable,
                           aes(x = CCA1, y = CCA2, label = Variable),
                           color = "blue",
                           size = 3.5) +
  scale_color_manual(values = seasonal_colors) +
  theme_minimal(6) +
  #labs(x = "Zooplankton", y = "Microeukaryotes", title = "Correlation between Taxa with 0-4 weeks lag") +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(aspect.ratio = 1)

ggsave("Figure_output/CCA_depth.pdf", width = 4, height = 4)
```


