---
title: "QU39 Time Series"
output: 
  flexdashboard::flex_dashboard:
    social: menu
    source_code: embed
    orientation: columns
    vertical_layout: fill
    theme: simplex
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
#Load necessary packages and import functions.

# All data format modifications and filtering and plotting
library(tidyverse)
'%!in%' <- function(x,y)!('%in%'(x,y))
source("Code/Functions.R")
library(zoo)

# Statistical tools
library(vegan)

# Specific plotting tools
library(plotly)
library(ggpubr)
library(paletteer)
library(viridis)
library(cowplot)
source("Code/scales_and_themes.R")
#library(ggnewscale)




bottle_file <- readRDS("Data_output/bottle_file.rds")
COI_species <- readRDS("Data_output/COI_species.rds")


Abundant_Species <- COI_species %>% 
  group_by(Taxa) %>%
  mutate(maxRA = median(RA)) %>%
  #filter(maxRA >= 0.01) %>% # 15 species
  #filter(max(RA) >0.14) %>% # 15 species
  filter(max(RA) >0.24) %>% # 10 Species
  pull(Taxa) %>%  unique
```





Explore Data
=======================================================================

```{r}
##### SERVER #############

# 1. Environmental Data

var_choices = c(
    "Temperature", "Salinity",
    "chlA (>20um)", "chlA (>3um)", "chlA (>0.7um)",
    "phaeo (>20um)", "phaeo (>3um)", "phaeo (>0.7um)",
    "NO2 & NO3", "PO4",
    "O2", "PAR", "Turbidity")

Bottle_File <- bottle_file %>% 
  dplyr::transmute(Hakai_ID = hakai_id,
                   Date = date,
                   Depth = line_out_depth,
                   depth = factor(line_out_depth, levels = c("260", "100",
                                                             "30", "5", "0")),
                   Temperature = Temperature,
                   Salinity = Salinity,
                   PAR = PAR,
                   Turbidity = Turbidity,
                   O2 = O2,
                   `chlA (>20um)` = chla_20um,
                   `chlA (>3um)` = chla_3um,
                   `chlA (>0.7um)` = `chla_GF/F`,
                   `phaeo (>20um)` = phaeo_20um,
                   `phaeo (>3um)` = phaeo_3um,
                   `phaeo (>0.7um)` = `phaeo_GF/F`,
                   `NO2 & NO3` =  no2_no3_um,
                   PO4 = po4
                   ) %>% 
  pivot_longer(5:15, names_to = "Variable", values_to = "value") %>% 
  dplyr::filter(is.na(value) == F) %>% 
  dplyr::mutate(Variable = factor(Variable, levels = var_choices))

plot_bottle <- function(Variable_filter, datemark) {
  
  Bottle_File %>%
  dplyr::filter(Variable %in% Variable_filter) %>% 
  ggplot() +
  geom_line(aes(Date, value, color = depth)) +
  geom_vline(xintercept = c(date("2017-01-01") + days(datemark-1),
                            date("2017-01-01") + days(datemark-1+365),
                            date("2017-01-01") + days(datemark-1+365+365),
                            color = "red", size = 0.4)) +
  facet_wrap("Variable", scales = "free", ncol = 1) +
  scale_x_date(date_labels="%b %y", breaks  = datebreaks) +
  theme_minimal_vgrid(font_size = 12) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        #strip.text = element_text(face = "italic"),
        panel.background = element_rect(fill = '#F5F5F5', colour = 'darkgrey'),
        #legend.position="none",,
        aspect.ratio = 1/4) +
  scale_color_manual(values = depth_colors)
  
}


# 2. Community Data

# Make species interpolation:
interpolate_taxa <- function(Abundant_Species, float) {
  
  COI_species %>%
  filter(Taxa %in% Abundant_Species) %>%
  group_by(Taxa) %>%
  group_modify(~interpolate_rollmean(.x, x1 = Date, x2 = Depth, y = RA_Index,
                                     # res 
                                     x1res = 1, x2res = 1, k = float,
                                     dim = 1))
}


taxa_plot <- function(interpolated_taxa, datemark, type) {
  
  if (type == "Line") {
    interpolated_taxa %>%
      mutate(Depth = factor(Depth)) %>%
      ggplot() +
      geom_line(aes(Date, RA_Index, color = Depth)) +
      geom_vline(xintercept = c(date("2017-01-01") + days(datemark-1),
                                date("2017-01-01") + days(datemark-1+365),
                                date("2017-01-01") + days(datemark-1+365+365),
                                color = "red", size = 0.4)) +
      facet_wrap("Taxa", scales = "free") +
      scale_x_date(date_labels="%b %y", breaks  = datebreaks) +
      theme_minimal_vgrid(font_size = 12) +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
            panel.background = element_rect(fill = '#F5F5F5', colour = 'darkgrey'),
            aspect.ratio = 1/4) +
      scale_color_manual(values = depth_colors)
  } else {
    
    interpolated_taxa %>%
      mutate(Depth = factor(Depth, levels = c(
        250, 100, 30, 5, 0 ))) %>% 
      ggplot() +
      geom_raster(aes(Date, Depth, fill = RA_Index)) +
      geom_vline(xintercept = datebreaks, color = "#F5F5F5", size = 0.2, alpha = 0.8) +
      geom_vline(xintercept = c(date("2017-01-01") + days(datemark-1),
                                date("2017-01-01") + days(datemark-1+365),
                                date("2017-01-01") + days(datemark-1+365+365),
                                color = "red", size = 0.4)) +
      facet_wrap("Taxa") +
      scale_x_date(date_labels="%b %y", breaks  = datebreaks) +
  scale_fill_gradient2(high = 'green',
                       low = "white") +
  theme_minimal_vgrid(font_size = 12) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        strip.text = element_text(face = "italic"),
        panel.background = element_rect(fill = '#F5F5F5', colour = 'darkgrey'),
        #legend.position="none",,
        aspect.ratio = 1/4)
    
    
    
  }

}


all_species <- COI_species %>% 
  pull(Taxa) %>% unique()




```

Inputs {.sidebar}
-----------------------------------------------------------------------

**Select 3 environmental environmental variables:**
```{r}
selectInput(
  "Filt_var1",
  NULL,
  choices = var_choices,
  selected = "Temperature"
)

selectInput(
  "Filt_var2",
  NULL,
  choices = var_choices,
  selected = "Salinity"
)

selectInput(
  "Filt_var3",
  NULL,
  choices = var_choices,
  selected = "O2"
)
```

**Select 3 taxa:**
```{r}
selectInput(
  "Filt_Taxa1",
  NULL,
  choices = all_species,
  selected = "Acartia longiremis"
)

selectInput(
  "Filt_Taxa2",
  NULL,
  choices = all_species,
  selected = "Eucalanus bungii"
)

selectInput(
  "Filt_Taxa3",
  NULL,
  choices = all_species,
  selected = "Pseudocalanus mimus"
)
```

**Day of year highlight:**
```{r}
sliderInput(
  "doy_mark",
  NULL,
  min = 1,
  max = 365,
  value = 1,
  step = 1
)
```


**Projection type (taxa):**
```{r}
# Line line or raster:
radioButtons(
  "Plottype",
  NULL,
  choices = c("Line", "Raster"),
  selected = "Line"
)
```

**Frame for floating average (taxa)**
```{r}
numericInput(
  "Floating_mean:",
  "Days:",
  14,
  min = 1,
  max = 365,
  step = 1
)
```



##### About

Built buy **Andreas Novotny** for the **Pelagic Ecosystems Lab** at the **University of British Columbia** and the **Hakai Institute**

**Contact:** a.novotny@oceans.ubc.ca


Column
-----------------------------------------------------------------------
### Environmental Variable 1
```{r}
renderPlot(plot_bottle(input$Filt_var1, input$doy_mark))
```

### Environmental Variable 2
```{r}
renderPlot(plot_bottle(input$Filt_var2, input$doy_mark))
```

### Environmental Variable 3
```{r}
renderPlot(plot_bottle(input$Filt_var3, input$doy_mark))
```

Column
-----------------------------------------------------------------------
### Taxa 1
```{r}

renderPlot({
  int <- interpolate_taxa(input$Filt_Taxa1, input$Floating_mean)
  taxa_plot(int, input$doy_mark, input$Plottype)
})

```

### Taxa 2
```{r}

renderPlot({
  int <- interpolate_taxa(input$Filt_Taxa2, input$Floating_mean)
  taxa_plot(int, input$doy_mark, input$Plottype)
})

```

### Taxa 3
```{r}

renderPlot({
  int <- interpolate_taxa(input$Filt_Taxa3, input$Floating_mean)
  taxa_plot(int, input$doy_mark, input$Plottype)
})

```



Community Analysis
=======================================================================


Inputs {.sidebar}
-----------------------------------------------------------------------

**CCA**

Constrained correspondence analysis. VIF stepwise reduction of envorinmental vaciables has been used to determine what autocorrelation between variables.

**NMDS**

NMDS of samples only (COI Libraries). Select depths to be included in NMDS:

```{r}
checkboxGroupInput(
  "Filter_depth",
  NULL,
  choices = c("0", "5", "30", "100", "250"),
  selected = c("0", "5", "30", "100", "250"),
  inline = FALSE,
  width = NULL,
  choiceNames = NULL,
  choiceValues = NULL
)
```

##### About

Built buy **Andreas Novotny** for the **Pelagic Ecosystems Lab** at the **University of British Columbia** and the **Hakai Institute**

**Contact:** a.novotny@oceans.ubc.ca


Column {.tabset}
-------------------------------------
### CCA

```{r}
plot_cca <- function(maxRA) {
  
  Abundant_Species <- COI_species %>% 
  group_by(Taxa) %>%
  mutate(maxRA = median(RA)) %>%
  filter(max(RA) >maxRA) %>% # 32 species
  pull(Taxa) %>%  unique

comm <- COI_species %>% 
  filter(Taxa %in% Abundant_Species) %>% 
  dplyr::group_by(Hakai_ID, Taxa) %>% 
  summarise(Index = mean(RA_Index)) %>%
  pivot_wider(names_from = "Taxa", values_from = "Index") %>% 
  arrange(Hakai_ID) %>% 
  column_to_rownames("Hakai_ID")

env <- bottle_file %>% 
  arrange(hakai_id) %>%
  mutate(Depth = factor(line_out_depth),
         Month = month(date)) %>%
  column_to_rownames("hakai_id")

mod <- cca(comm ~
             Temperature + Salinity + PAR + Turbidity + O2 +
             chla_20um + chla_3um +`chla_GF/F` +
             no2_no3_um,
           data = env,
           na.action = na.omit)


#extracting the data as data frame; env data
veg_1 = as.data.frame(mod$CCA$biplot) %>% 
  rownames_to_column("env")

#extracting the data; genusv
veg_2 = as.data.frame(mod$CCA$v) %>%
  rownames_to_column("Taxa")

cca_plot <-  ggplot() +
  geom_point(data = veg_2, aes(x = CCA1, y = CCA2), color = "red") +
  #geom_point(data = veg_1, aes(x = CCA1, y = CCA2), color = "blue") +
  ggrepel::geom_text_repel(data = veg_2,
                           aes(x = CCA1, y = CCA2, label = Taxa),
                           nudge_y = -0.05, size = 1.5) +

  geom_segment(data = veg_1,
               aes(x = 0, y = 0, xend = CCA1, yend = CCA2),
               arrow = arrow(length = unit(0.25, "cm"))) +
  ggrepel::geom_text_repel(data = veg_1,
                           aes(x = CCA1, y = CCA2, label = env),
                           nudge_y = -0.05,
                           color = "blue", size = 2) +
  theme_bw()

cca_plot
}


plot_cca(0.08)

```


### NMDS


```{r}
## WRAPPER
plotNMDS <- function(x) {
  

mat <- COI_species %>%
  mutate(Depth = factor(Depth)) %>% 
  dplyr::filter(Depth %in% x) %>% 
  dplyr::select(Library_ID, Taxa, RA_Index) %>% 
  pivot_wider(names_from = Taxa, values_from = RA_Index) %>% 
  column_to_rownames("Library_ID") %>% 
  as.matrix()
  
  mat
  
mod <- metaMDS(mat, trymax = 20)

# Extract NMDS data
{  Samples <- 
    scores(mod)$sites %>% 
    as.data.frame() %>%
    rownames_to_column("Library_ID") %>% 
    mutate(Layer = "Samples")
  
  Taxa <-
    scores(mod)$species %>% 
    as.data.frame() %>%
    rownames_to_column("Taxa") %>% 
    mutate(Layer = "Taxa")
  
  NMDS_res <- bind_rows(Samples, Taxa) %>% 
    pivot_wider(names_from = Layer,
              values_from = c(NMDS1, NMDS2))
}

# Merge with metadata
COI_species %>% 
  dplyr::select(Date, Depth, Library_ID) %>%
  mutate(Month = format(as.Date(Date), format="%m"),
         Depth = factor(Depth)) %>%
  left_join(NMDS_res, by = "Library_ID") %>% 
  ggplot(aes(NMDS1_Samples, NMDS2_Samples)) +
  geom_text(aes(NMDS1_Taxa, NMDS2_Taxa, label = Taxa),
            size = 3, colour = "grey", fontface = "italic") +
  geom_point(aes(color = Month, shape = Depth), size = 3) +
  scale_color_manual(values = seasonal_colors)


}


renderPlot({
  plotNMDS(input$Filter_depth)
})

```



Diagnostic
=======================================================================

Inputs {.sidebar}
-----------------------------------------------------------------------

**Relative Abundance**

Plotting all taxa identified with COI eDNA. Ordered by their relative sequence read abundance.

**Library coverage**

Successful COI libraries and their distributions over depth / date.


##### About

Built buy **Andreas Novotny** for the **Pelagic Ecosystems Lab** at the **University of British Columbia** and the **Hakai Institute**

**Contact:** a.novotny@oceans.ubc.ca

Column {.tabset}
-------------------------------------
### Relative abundance



```{r}
COI_species %>%
  filter(Abundance > 0) %>%
  group_by(Taxa) %>% 
  mutate(maxRA = median(RA)) %>%
  ungroup() %>% 
  mutate(Taxa = fct_reorder(Taxa, -maxRA)) %>% 
  ggplot() +
  geom_boxplot(aes(Taxa, RA)) +
  theme_minimal_hgrid(font_size = 8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,
                                   face = "italic"),
        panel.background = element_rect(fill = '#F5F5F5', colour = 'darkgrey')) +
  scale_y_continuous(trans='log10')
```


### Library coverage



```{r}
COI_species %>% 
  group_by(Library_ID, Depth, Date) %>% 
  summarise(Abundance = sum(Abundance)) %>% 
  mutate(Depth = factor(Depth, levels = c(
    250, 100, 30, 5, 0
  ))) %>% 
  ggplot() +
  geom_raster(aes(Date, Depth)) +
  geom_vline(xintercept = datebreaks, color = "#F5F5F5", size = 0.2, alpha = 0.8) +
  geom_vline(xintercept = yearbreaks, color = "grey", size = 0.2, alpha = 0.8) +
  scale_x_date(date_labels="%b %y", breaks  = datebreaks) +
  #scale_fill_gradient2(high = '#4daf4a') +
  theme_minimal_vgrid(font_size = 8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.background = element_rect(fill = '#F5F5F5', colour = 'darkgrey'),
        aspect.ratio = 1/4)


```


About
=======================================================================

